
### Copy number aberrations as biomarkers

A total of __`r NROW(pcg_report[['content']][['cna']][["disp"]][['tier1']]) + NROW(pcg_report[['content']][['cna']][["disp"]][['tier2']])`__ aberrations are associated with clinical evidence items in the [database for clinical interpretations of variants in cancer, CIViC](https://civic.genome.wustl.edu/#/home), with the following number of evidence items:

 *  Predictive: __`r NROW(pcg_report[['content']][['cna']][['clin_eitem']][['query_ttype']][['predictive']][['A_B']]) + NROW(pcg_report[['content']][['cna']][['clin_eitem']][['other_ttype']][['predictive']][['A_B']]) + NROW(pcg_report[['content']][['cna']][['clin_eitem']][['query_ttype']][['predictive']][['C_D_E']])`__ evidence items
 *  Prognostic: __`r NROW(pcg_report[['content']][['cna']][['clin_eitem']][['query_ttype']][['prognostic']][['A_B']]) + NROW(pcg_report[['content']][['cna']][['clin_eitem']][['other_ttype']][['prognostic']][['A_B']]) + NROW(pcg_report[['content']][['cna']][['clin_eitem']][['query_ttype']][['prognostic']][['C_D_E']])`__ evidence items
 *  Diagnostic: __`r NROW(pcg_report[['content']][['cna']][['clin_eitem']][['query_ttype']][['diagnostic']][['A_B']]) + NROW(pcg_report[['content']][['cna']][['clin_eitem']][['other_ttype']][['diagnostic']][['A_B']]) +
NROW(pcg_report[['content']][['cna']][['clin_eitem']][['query_ttype']][['diagnostic']][['C_D_E']])`__ evidence items

<br><br>

```{r prepare_cna_clin_data, echo=F, results='asis'}

show_cna_filters <- list()
missing_cna_variants <- list()
cna_evidence_items <- list()
for(sig in c('strong_significance','potential_significance')){
  show_cna_filters[[sig]] <- list()
  missing_cna_variants[[sig]] <- list()
  cna_evidence_items[[sig]] <- list()
  for(type in c('diagnostic','prognostic','predictive')){
    show_cna_filters[[sig]][[type]] <- F
    missing_cna_variants[[sig]][[type]] <- T
    cna_evidence_items[[sig]][[type]] <- data.frame()
    if(sig == 'strong_significance' & NROW(pcg_report[['content']][['cna']][['clin_eitem']][['query_ttype']][[type]][['A_B']]) > 0){
      show_cna_filters[[sig]][[type]] <- T
      missing_cna_variants[[sig]][[type]] <- F
      cna_evidence_items[[sig]][[type]] <- pcg_report[['content']][['cna']][['clin_eitem']][['query_ttype']][[type]][['A_B']]
    }
    
    if(sig == 'potential_significance' & (NROW(pcg_report[['content']][['cna']][['clin_eitem']][['query_ttype']][[type]][['C_D_E']]) > 0 | NROW(pcg_report[['content']][['cna']][['clin_eitem']][['other_ttype']][[type]][['A_B']]) > 0)){
      show_cna_filters[[sig]][[type]] <- T
      missing_cna_variants[[sig]][[type]] <- F
      cna_evidence_items[[sig]][[type]] <- 
        dplyr::bind_rows(pcg_report[['content']][['cna']][['clin_eitem']][['other_ttype']][[type]][['A_B']], pcg_report[['content']][['cna']][['clin_eitem']][['query_ttype']][[type]][['C_D_E']])
    }
  }
}

```


#### Aberrations of strong clinical significance {.tabset}

```{r active_tab_strong_predictive, echo = F, results = "asis"}
if(missing_cna_variants[['strong_significance']][['predictive']] == F | 
   (missing_cna_variants[['strong_significance']][['diagnostic']] == T & 
    missing_cna_variants[['strong_significance']][['prognostic']] == T)){
  cat("")
  cat("##### Predictive biomarkers {.active}")
  cat("")
}else{
  cat("")
  cat("##### Predictive biomarkers")
  cat("")
}
```

<br>

```{r cna_predictive, echo=F, results = 'asis', eval = show_cna_filters[['strong_significance']][['predictive']]}

cna_predictive_shared_strong <- crosstalk::SharedData$new(cna_evidence_items[['strong_significance']][['predictive']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", cna_predictive_shared_strong, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical significance", cna_predictive_shared_strong, ~CLINICAL_SIGNIFICANCE),
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", cna_predictive_shared_strong, ~EVIDENCE_LEVEL)
  ),
  list(
    crosstalk::filter_select("THERAPEUTIC_CONTEXT", "Therapeutic context", cna_predictive_shared_strong, ~THERAPEUTIC_CONTEXT),
    crosstalk::filter_slider("LogR", "Log-ratio", cna_predictive_shared_strong, ~LOG_R, ticks = F),
    crosstalk::filter_slider("RATING", "Rating", cna_predictive_shared_strong, ~RATING, min = 0, max = 5, step = 1, ticks = T)
  )
)

cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

DT::datatable(cna_predictive_shared_strong, escape=F,extensions=c("Buttons","Responsive"),
              options=list(buttons = c('csv','excel'), scrollCollapse=T, dom = 'Bfrtip')) |>
  DT::formatStyle('EVIDENCE_LEVEL', 
                  backgroundColor = DT::styleEqual(c('A: Validated','A: FDA/NCCN/ELN guidelines',
                                                     'B: Clinical evidence',
                                                     'B1: Clinical evidence: late trials',
                                                     'B2: Clinical evidence: early trials',
                                                     'C: Case study',
                                                     'D: Preclinical evidence',
                                                     'E: Indirect evidence'), 
                                                   c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))



```

```{r cna_predictive_missing, echo=F, results = 'asis', eval = missing_cna_variants[['strong_significance']][['predictive']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>


```{r active_tab_strong_prognostic, echo = F, results = "asis"}
if(missing_cna_variants[['strong_significance']][['prognostic']] == F & 
   missing_cna_variants[['strong_significance']][['predictive']] == T){
  cat("")
  cat("##### Prognostic biomarkers {.active}")
  cat("")
}else{
  cat("")
  cat("##### Prognostic biomarkers")
  cat("")
}
```

<br>

```{r cna_prognostic, echo=F, results='asis', eval = show_cna_filters[['strong_significance']][['prognostic']]}

cna_prognostic_shared_strong <- crosstalk::SharedData$new(cna_evidence_items[['strong_significance']][['prognostic']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", cna_prognostic_shared_strong, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical significance", cna_prognostic_shared_strong, ~CLINICAL_SIGNIFICANCE),
    crosstalk::filter_slider("RATING", "Rating", cna_prognostic_shared_strong, ~RATING, min = 0, max = 5, step = 1, ticks = T)
  ),
  list(
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", cna_prognostic_shared_strong, ~EVIDENCE_LEVEL),
    crosstalk::filter_slider("LOG_R", "Log-ratio", cna_prognostic_shared_strong, ~LOG_R, ticks = F)    
  )
)
cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

DT::datatable(cna_prognostic_shared_strong, escape=F,extensions=c("Buttons","Responsive"),
              options=list(buttons = c('csv','excel'), scrollCollapse=T, dom = 'Bfrtip')) |>
  DT::formatStyle('EVIDENCE_LEVEL', 
                  backgroundColor = DT::styleEqual(c('A: Validated','A: FDA/NCCN/ELN guidelines',
                                                     'B: Clinical evidence',
                                                     'B1: Clinical evidence: late trials',
                                                     'B2: Clinical evidence: early trials',
                                                     'C: Case study',
                                                     'D: Preclinical evidence',
                                                     'E: Indirect evidence'), 
                                                   c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))


```

```{r cna_prognostic_missing, echo=F, results = 'asis', eval = missing_cna_variants[['strong_significance']][['prognostic']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>



```{r active_tab_strong_diagnostic, echo = F, results = "asis"}
if(missing_cna_variants[['strong_significance']][['diagnostic']] == F & 
   missing_cna_variants[['strong_significance']][['predictive']] == T & 
   missing_cna_variants[['strong_significance']][['prognostic']] == T){
  cat("")
  cat("##### Diagnostic biomarkers {.active}")
  cat("")
}else{
  cat("")
  cat("##### Diagnostic biomarkers")
  cat("")
}
```


<br>

```{r cna_diagnostic, echo=F, results='asis', eval = show_cna_filters[['strong_significance']][['diagnostic']]}

cna_diagnostic_shared_strong <- crosstalk::SharedData$new(cna_evidence_items[['strong_significance']][['diagnostic']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", cna_diagnostic_shared_strong, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical association", cna_diagnostic_shared_strong, ~CLINICAL_SIGNIFICANCE),
    crosstalk::filter_slider("RATING", "Rating", cna_diagnostic_shared_strong, ~RATING, min = 0, max = 5, step = 1, ticks = T)
  ),
  list(
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", cna_diagnostic_shared_strong, ~EVIDENCE_LEVEL),
    crosstalk::filter_slider("LOG_R", "Log-ratio", cna_diagnostic_shared_strong, ~LOG_R, ticks = F)
  )
)
cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

DT::datatable(cna_diagnostic_shared_strong, escape=F,extensions=c("Buttons","Responsive"),
              options=list(buttons = c('csv','excel'), scrollCollapse=T, dom = 'Bfrtip')) |>
  DT::formatStyle('EVIDENCE_LEVEL', 
                  backgroundColor = DT::styleEqual(c('A: Validated','A: FDA/NCCN/ELN guidelines',
                                                     'B: Clinical evidence',
                                                     'B1: Clinical evidence: late trials',
                                                     'B2: Clinical evidence: early trials',
                                                     'C: Case study',
                                                     'D: Preclinical evidence',
                                                     'E: Indirect evidence'), 
                                                   c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))


```

```{r cna_diagnostic_missing, echo=F, results = 'asis', eval = missing_cna_variants[['strong_significance']][['diagnostic']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>

#### Aberrations of potential clinical significance {.tabset}

```{r active_tab_potential_predictive, echo = F, results = "asis"}
if(missing_cna_variants[['potential_significance']][['predictive']] == F | 
   (missing_cna_variants[['potential_significance']][['diagnostic']] == T & 
    missing_cna_variants[['potential_significance']][['prognostic']] == T)){
  cat("")
  cat("##### Predictive biomarkers {.active}")
  cat("")
}else{
  cat("")
  cat("##### Predictive biomarkers")
  cat("")
}
```


<br>

```{r cna_predictive_tier2, echo=F, results = 'asis', eval = show_cna_filters[['potential_significance']][['predictive']]}

cna_predictive_shared <- crosstalk::SharedData$new(cna_evidence_items[['potential_significance']][['predictive']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", cna_predictive_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical significance", cna_predictive_shared, ~CLINICAL_SIGNIFICANCE),
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", cna_predictive_shared, ~EVIDENCE_LEVEL)
  ),
  list(
    crosstalk::filter_select("THERAPEUTIC_CONTEXT", "Therapeutic context", cna_predictive_shared, ~THERAPEUTIC_CONTEXT),
    crosstalk::filter_slider("LOG_R", "Log-ratio", cna_predictive_shared, ~LOG_R, ticks = F),
    crosstalk::filter_slider("RATING", "Rating", cna_predictive_shared, ~RATING, min = 0, max = 5, step = 1, ticks = T)
  )
)

cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

DT::datatable(cna_predictive_shared, escape=F,extensions=c("Buttons","Responsive"),
              options=list(buttons = c('csv','excel'), scrollCollapse = T, dom = 'Bfrtip')) |>
  DT::formatStyle('EVIDENCE_LEVEL', 
                  backgroundColor = DT::styleEqual(c('A: Validated','A: FDA/NCCN/ELN guidelines',
                                                     'B: Clinical evidence',
                                                     'B1: Clinical evidence: late trials',
                                                     'B2: Clinical evidence: early trials',
                                                     'C: Case study',
                                                     'D: Preclinical evidence',
                                                     'E: Indirect evidence'), 
                                                   c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))



```

```{r cna_predictive_tier2_missing, echo=F, results = 'asis', eval = missing_cna_variants[['potential_significance']][['predictive']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>


```{r active_tab_potential_prognostic, echo = F, results = "asis"}
if(missing_cna_variants[['potential_significance']][['prognostic']] == F & 
   missing_cna_variants[['potential_significance']][['predictive']] == T){
  cat("")
  cat("##### Prognostic biomarkers {.active}")
  cat("")
}else{
  cat("")
  cat("##### Prognostic biomarkers")
  cat("")
}
```


<br>

```{r cna_prognostic_tier2, echo=F, results='asis', eval = show_cna_filters[['potential_significance']][['prognostic']]}

cna_prognostic_shared <- crosstalk::SharedData$new(cna_evidence_items[['potential_significance']][['prognostic']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", cna_prognostic_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical significance", cna_prognostic_shared, ~CLINICAL_SIGNIFICANCE),
    crosstalk::filter_slider("RATING", "Rating", cna_prognostic_shared, ~RATING, min = 0, max = 5, step = 1, ticks = T)
  ),
  list(
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", cna_prognostic_shared, ~EVIDENCE_LEVEL),
    crosstalk::filter_slider("LOG_R", "Log-ratio", cna_prognostic_shared, ~LOG_R, ticks = F)    
  )
)
cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

DT::datatable(cna_prognostic_shared, escape=F, extensions=c("Buttons","Responsive"),
              options=list(buttons = c('csv','excel'), scrollCollapse=T, dom = 'Bfrtip')) |>
  DT::formatStyle('EVIDENCE_LEVEL', 
                  backgroundColor = DT::styleEqual(
                    c('A: Validated','A: FDA/NCCN/ELN guidelines',
                      'B: Clinical evidence',
                      'B1: Clinical evidence: late trials',
                      'B2: Clinical evidence: early trials',
                      'C: Case study',
                      'D: Preclinical evidence',
                      'E: Indirect evidence'), 
                    c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))



```

```{r cna_prognostic_tier2_missing, echo=F, results = 'asis', eval = missing_cna_variants[['potential_significance']][['prognostic']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>

```{r active_tab_potential_diagnostic, echo = F, results = "asis"}
if(missing_cna_variants[['potential_significance']][['diagnostic']] == F & 
   missing_cna_variants[['potential_significance']][['predictive']] == T & 
   missing_cna_variants[['potential_significance']][['prognostic']] == T){
  cat("")
  cat("##### Diagnostic biomarkers {.active}")
  cat("")
}else{
  cat("")
  cat("##### Diagnostic biomarkers")
  cat("")
}
```

<br>

```{r cna_diagnostic_tier2, echo=F, results='asis', eval = show_cna_filters[['potential_significance']][['diagnostic']]}

cna_diagnostic_shared <- crosstalk::SharedData$new(cna_evidence_items[['potential_significance']][['diagnostic']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", cna_diagnostic_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical association", cna_diagnostic_shared, ~CLINICAL_SIGNIFICANCE),
    crosstalk::filter_slider("RATING", "Rating", cna_diagnostic_shared, ~RATING, min = 0, max = 5, step = 1, ticks = T)
  ),
  list(
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", cna_diagnostic_shared, ~EVIDENCE_LEVEL),
    crosstalk::filter_slider("LOG_R", "Log-ratio", cna_diagnostic_shared, ~LOG_R, ticks = F)
  )
)
cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

DT::datatable(cna_diagnostic_shared, escape=F,extensions=c("Buttons","Responsive"),
              options=list(buttons = c('csv','excel'),  scrollCollapse=T, dom = 'Bfrtip')) |>
  DT::formatStyle('EVIDENCE_LEVEL', 
                  backgroundColor = DT::styleEqual(
                    c('A: Validated',
                      'A: FDA/NCCN/ELN guidelines',
                      'B: Clinical evidence',
                      'B1: Clinical evidence: late trials',
                      'B2: Clinical evidence: early trials',
                      'C: Case study',
                      'D: Preclinical evidence',
                      'E: Indirect evidence'), 
                                                   c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))


```

```{r cna_diagnostic_tier2_missing, echo=F, results = 'asis', eval = missing_cna_variants[['potential_significance']][['diagnostic']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>
