### Gene expression outliers

Here, we show the protein-coding gene expression outliers identified in the RNA input sample, as compared to the selected reference cohort. We consider expression outliers as genes with expression levels (specifically log-transformed TPM values, see below) located at the extreme ends of the distributions observed in the reference cohort, specifically those with:

1) An expression level in the top or bottom percentiles (i.e. above 95 or below 5), _and_
2) An expression level that is more than $Q3 + 1.5 \times IQR$ (increased expression) or less than $Q1 - 1.5 \times IQR$ (decreased expression), _and_
3) An expression level with a Z-score above 2 or below -2

__TPM__: Transcripts per Million
__IQR__: Interquartile range
__Q1__: First quartile
__Q3__: Third quartile

Outlier genes shown here are furthermore limited to the ones that have some evidence of being cancer-associated (including, yet not limited to, tumor suppressors, oncogenes, and predicted cancer driver genes), as well as targets of approved cancer drugs.

::: {.callout-note collapse="true"}

## Note - output limitations

A maximum of __N = 300__ outlier genes are shown per category (increased/reduced expression), ranked by absolute Z-score. Additionally, all overexpressed __oncogenes__ (increased expression) and underexpressed __tumor suppressors__ (reduced expression) are always included regardless of rank. For a more complete list, please refer to the TSV/Excel output generated by PCGR.

:::

::: {.callout-note}

## Note - gene expression unit

In the table below, column __EXPR_LEVEL__ indicates the expression level of a gene in the input sample as
_log2(TPM + 0.001)_

:::

::: {.callout-warning}

## Note - gene expression outlier detection

A comparison of gene expression levels identified in the input sample to those found in reference cohorts is not a straightforward procedure. The results needs to be interpreted with considerable caution, particularly taking into account potential biases in RNA sequencing protocols, including differences with respect to total RNA concentrations [@Zhao2020-ln]. Moreover, one should note that the vast majority of samples in the current reference cohorts for comparisons (TCGA) contain expression profiles for primary tumors, with only a minor contribution for metastatic contexts. The outlier list shown here could thus potentially include false positives, and should primarily be considered as a starting point for further investigation.

:::

<br>

```{r check_expression_outliers_data}
#| eval: true
#| output: false

outliers_present <- FALSE
outliers_combined <- data.frame()

if (!is.null(pcg_report$content$expression$outliers)) {
  if (NROW(pcg_report$content$expression$outliers) > 0) {

    all_high <- pcg_report$content$expression$outliers |>
      dplyr::filter(.data$EXPR_OUTLIER == "Increased expression") |>
      dplyr::arrange(dplyr::desc(.data$Z_SCORE))

    outliers_high <- dplyr::bind_rows(
      utils::head(all_high, 300),
      dplyr::filter(all_high, .data$ONCOGENE == TRUE)) |>
      dplyr::distinct(.data$SYMBOL, .keep_all = TRUE)

    all_low <- pcg_report$content$expression$outliers |>
      dplyr::filter(.data$EXPR_OUTLIER == "Reduced expression") |>
      dplyr::arrange(dplyr::desc(abs(.data$Z_SCORE)))

    outliers_low <- dplyr::bind_rows(
      utils::head(all_low, 300),
      dplyr::filter(all_low, .data$TUMOR_SUPPRESSOR == TRUE)) |>
      dplyr::distinct(.data$SYMBOL, .keep_all = TRUE)

    outliers_combined <- dplyr::bind_rows(outliers_high, outliers_low)

    ## format expression levels with two decimal places for display
    outliers_combined <- outliers_combined |>
      dplyr::mutate(
        EXPR_LEVEL = sprintf("%.2f", EXPR_LEVEL),
        PERCENTILE = sprintf("%.1f", PERCENTILE)
      ) |>
      dplyr::mutate(
        SEARCH_INDEX = stringr::str_replace_all(
          paste(
            .data$SYMBOL,
            .data$GENENAME,
            dplyr::if_else(
              .data$TUMOR_SUPPRESSOR == TRUE,
              "tumor suppressor", ""),
            dplyr::if_else(
              .data$ONCOGENE == TRUE,
              "oncogene", ""),
            sep = " "
          ),
          "\\s+", " "
        )
      )

    if (NROW(outliers_combined) > 0) {
      outliers_present <- TRUE
    }
  }
}
```

```{r expression_outliers_table}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(outliers_present)

primary_cols <- c("SYMBOL", "EXPR_OUTLIER", "Z_SCORE", "EXPR_LEVEL", "PERCENTILE")

outliers_shared <- crosstalk::SharedData$new(outliers_combined)

crosstalk::bscols(
  widths = c(5, 7),
  crosstalk::filter_checkbox(
    "EXPR_OUTLIER", "Outlier type",
    outliers_shared, ~EXPR_OUTLIER, inline = TRUE),
  crosstalk::filter_slider(
    "Z_SCORE", "Z-score",
    outliers_shared, ~Z_SCORE)
)

high_breaks <- pcgrr::color_palette$expression_outlier_high$breaks
high_colors <- pcgrr::color_palette$expression_outlier_high$values
low_breaks  <- pcgrr::color_palette$expression_outlier_low$breaks
low_colors  <- pcgrr::color_palette$expression_outlier_low$values

expr_outlier_style <- function(data) {
  function(value, index) {
    pct          <- data$PERCENTILE[index]
    outlier_type <- data$EXPR_OUTLIER[index]
    if (is.na(pct) || is.na(outlier_type)) return(list())
    if (outlier_type == "Increased expression") {
      bin <- findInterval(pct, high_breaks) + 1
      bg  <- high_colors[min(bin, length(high_colors))]
    } else {
      bin <- findInterval(pct, low_breaks) + 1
      bg  <- low_colors[min(bin, length(low_colors))]
    }
    list(background = bg, color = "white", fontWeight = "bold", padding = "4px 8px")
  }
}

col_defs <- list(
  SYMBOL = reactable::colDef(
    name     = "Gene",
    style    = pcgrr::render_symbol_assoc_style(
      outliers_combined, rank_col = "GLOBAL_ASSOC_RANK"),
    minWidth = 100
  ),
  EXPR_OUTLIER = reactable::colDef(
    name     = "Outlier type",
    style    = expr_outlier_style(outliers_combined),
    minWidth = 130
  ),
  Z_SCORE = reactable::colDef(
    name     = "Z-score",
    minWidth = 100,
    align    = "center"
  ),
  EXPR_LEVEL = reactable::colDef(
    name     = "Expr. level",
    minWidth = 120,
    align    = "center"
  ),
  PERCENTILE = reactable::colDef(
    name     = "Percentile",
    maxWidth = 120,
    align    = "center"
  )
)

## Hide all non-primary columns from the main view;
## SEARCH_INDEX is kept for searchable = TRUE but never displayed
all_cols <- names(outliers_combined)
for (col in setdiff(all_cols, primary_cols)) {
  col_defs[[col]] <- reactable::colDef(show = FALSE)
}
col_defs[["SEARCH_INDEX"]] <- reactable::colDef(
  show = FALSE,
  searchable = TRUE
)

reactable::reactable(
  outliers_shared,
  columns        = col_defs,
  defaultColDef  = reactable::colDef(html = TRUE),
  details        = pcgrr::build_rt_row_details(
    primary_cols = primary_cols),
  striped        = FALSE,
  highlight      = TRUE,
  compact        = TRUE,
  wrap           = TRUE,
  searchable     = TRUE,
  filterable     = TRUE,
  defaultPageSize = 10,
  theme          = pcgrr::rt_theme()
)
```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(!outliers_present)
cat("<div style='padding:8px 14px; border-left:4px solid #868686; background:#f5f5f5; border-radius:0 4px 4px 0; color:#444; font-size:0.92em;'>No expression outliers identified in this sample.</div>")
```

<br><br>
