
### Copy number aberrations as biomarkers for prognosis, diagnosis, predisposition, and drug response {.tabset}

A total of __`r nrow(report_data$cna_data$cna_biomarker_segments)`__ aberrations are associated with clinical evidence items in the [database for clinical interpretations of variants in cancer, CIViC](https://civic.genome.wustl.edu/#/home), with the following number of evidence items:

 *  Predictive: __`r nrow(report_data$cna_data$cna_evidence_items_predictive)`__ evidence items linked to drug sensitivity/resistance
 *  Prognostic: __`r nrow(report_data$cna_data$cna_evidence_items_prognostic)`__ evidence items
 *  Diagnostic: __`r nrow(report_data$cna_data$cna_evidence_items_diagnostic)`__ evidence items
 *  Predisposing: __`r nrow(report_data$cna_data$cna_evidence_items_predisposing)`__ evidence items

<br><br>
```{r prepare_cna_clin_data, echo=F, results='asis'}

show_cna_diagnostic_filters <- F
missing_cna_diagnostic_variants <- T
if(nrow(report_data$cna_data$cna_evidence_items_diagnostic) > 0){
  show_cna_diagnostic_filters <- T
  missing_cna_diagnostic_variants <- F
  report_data$cna_data$cna_evidence_items_diagnostic <- dplyr::rename(report_data$cna_data$cna_evidence_items_diagnostic, Segment = SEGMENT, Length_Mb = SEGMENT_LENGTH_MB)
}

show_cna_prognostic_filters <- F
missing_cna_prognostic_variants <- T
if(nrow(report_data$cna_data$cna_evidence_items_prognostic) > 0){
  show_cna_prognostic_filters <- T
  missing_cna_prognostic_variants <- F
  report_data$cna_data$cna_evidence_items_prognostic <- dplyr::rename(report_data$cna_data$cna_evidence_items_prognostic, Segment = SEGMENT, Length_Mb = SEGMENT_LENGTH_MB)
}

show_cna_predictive_filters <- F
missing_cna_predictive_variants <- T
if(nrow(report_data$cna_data$cna_evidence_items_predictive) > 0){
  show_cna_predictive_filters <- T
  missing_cna_predictive_variants <- F
  report_data$cna_data$cna_evidence_items_predictive <- dplyr::rename(report_data$cna_data$cna_evidence_items_predictive, Segment = SEGMENT, Length_Mb = SEGMENT_LENGTH_MB)
}

show_cna_predisposing_filters <- F
missing_cna_predisposing_variants <- T
if(nrow(report_data$cna_data$cna_evidence_items_predisposing) > 0){
  show_cna_predisposing_filters <- T
  missing_cna_predisposing_variants <- F
  report_data$cna_data$cna_evidence_items_predisposing <- dplyr::rename(report_data$cna_data$cna_evidence_items_predisposing, Segment = SEGMENT, Length_Mb = SEGMENT_LENGTH_MB)
}

```


#### Predictive biomarkers

<br>

```{r cna_predictive, echo=F, results = 'asis', eval = show_cna_predictive_filters}
library(summarywidget)
library(crosstalk)

cna_predictive_shared <- crosstalk::SharedData$new(report_data$cna_data$cna_evidence_items_predictive)
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", cna_predictive_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical significance", cna_predictive_shared, ~CLINICAL_SIGNIFICANCE),
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", cna_predictive_shared, ~EVIDENCE_LEVEL)
  ),
  list(
    crosstalk::filter_select("THERAPEUTIC_CONTEXT", "Therapeutic context", cna_predictive_shared, ~THERAPEUTIC_CONTEXT),
    crosstalk::filter_slider("LogR", "Log-ratio", cna_predictive_shared, ~LogR, ticks = F)
  )
)

cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(cna_predictive_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip')) %>%
  DT::formatStyle('EVIDENCE_LEVEL', backgroundColor = DT::styleEqual(c('A: Validated','B: Clinical evidence','C: Case study','D: Preclinical evidence','E: Indirect evidence'), c("#009E73","#56B4E9", "#0072B2","#E69F00", "#F0E442")))
)

#htmltools::br()

```

```{r cna_predictive_missing, echo=F, results = 'asis', eval = missing_cna_predictive_variants}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>

#### Prognostic biomarkers

<br>

```{r cna_prognostic, echo=F, results='asis', eval = show_cna_prognostic_filters}

library(summarywidget)
library(crosstalk)

cna_prognostic_shared <- crosstalk::SharedData$new(report_data$cna_data$cna_evidence_items_prognostic)
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", cna_prognostic_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical significance", cna_prognostic_shared, ~CLINICAL_SIGNIFICANCE)
  ),
  list(
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", cna_prognostic_shared, ~EVIDENCE_LEVEL),
    crosstalk::filter_slider("LogR", "Log-ratio", cna_prognostic_shared, ~LogR, ticks = F)    
  )
)
cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(cna_prognostic_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip')) %>%
  DT::formatStyle('EVIDENCE_LEVEL', backgroundColor = DT::styleEqual(c('A: Validated','B: Clinical evidence','C: Case study','D: Preclinical evidence','E: Indirect evidence'), c("#009E73","#56B4E9", "#0072B2","#E69F00", "#F0E442")))
)
#htmltools::br()

```

```{r cna_prognostic_missing, echo=F, results = 'asis', eval = missing_cna_prognostic_variants}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>

#### Diagnostic biomarkers

<br>

```{r cna_diagnostic, echo=F, results='asis', eval = show_cna_diagnostic_filters}

library(summarywidget)
library(crosstalk)

cna_diagnostic_shared <- crosstalk::SharedData$new(report_data$cna_data$cna_evidence_items_diagnostic)
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", cna_diagnostic_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical association", cna_diagnostic_shared, ~CLINICAL_SIGNIFICANCE)
  ),
  list(
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", cna_diagnostic_shared, ~EVIDENCE_LEVEL),
    crosstalk::filter_slider("LogR", "Log-ratio", cna_diagnostic_shared, ~LogR, ticks = F)
  )
)
cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(cna_diagnostic_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip')) %>%
  DT::formatStyle('EVIDENCE_LEVEL', backgroundColor = DT::styleEqual(c('A: Validated','B: Clinical evidence','C: Case study','D: Preclinical evidence','E: Indirect evidence'), c("#009E73","#56B4E9", "#0072B2","#E69F00", "#F0E442")))
)
#htmltools::br()

```

```{r cna_diagnostic_missing, echo=F, results = 'asis', eval = missing_cna_diagnostic_variants}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>

#### Predisposition biomarkers

<br>

```{r cna_predisposing, echo=F, results='asis', eval = show_cna_predisposing_filters}

library(summarywidget)
library(crosstalk)

cna_predisposing_shared <- crosstalk::SharedData$new(report_data$cna_data$cna_evidence_items_predisposing)
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", cna_predisposing_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical significance", cna_predisposing_shared, ~CLINICAL_SIGNIFICANCE)
  ),
  list(
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", cna_predisposing_shared, ~EVIDENCE_LEVEL),
    crosstalk::filter_slider("LogR", "Log-ratio", cna_predisposing_shared, ~LogR, ticks = F)
  )
)
cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(cna_predisposing_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip')) %>%
  DT::formatStyle('EVIDENCE_LEVEL', backgroundColor = DT::styleEqual(c('A: Validated','B: Clinical evidence','C: Case study','D: Preclinical evidence','E: Indirect evidence'), c("#009E73","#56B4E9", "#0072B2","#E69F00", "#F0E442")))
)
#htmltools::br()

```

```{r cna_predisposing_missing, echo=F, results = 'asis', eval = missing_cna_predisposing_variants}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>

