## MSI status

Microsatellite instability (MSI) is the result of impaired DNA mismatch repair and constitutes a cellular phenotype of clinical significance in many cancer types, most prominently colorectal cancers, stomach cancers, endometrial cancers, and ovarian cancers ([Cortes-Ciriano et al., 2017](https://www.ncbi.nlm.nih.gov/pubmed/28585546)). We have built a statistical MSI classifier from somatic mutation profiles that separates _MSI.H_ (MSI-high) from _MSS_ (MS stable) tumors. The MSI classifier was trained using 1000 exome-sequenced TCGA tumor samples with known MSI status (i.e. assayed from mononucleotide markers), and achieved 90.2% sensitivity and 99.7% specificity on an independent test set of 427 samples. Details of the MSI classification approach can be found <a href="http://rpubs.com/sigven/msi" target="_blank">here</a>.

<br>

* _Predicted MSI status for `r report_data$sample_name`_: __`r report_data$msi_prediction_data$msi_stats$predicted_class`__

### Supporting evidence I: indel fraction among somatic calls

The plot below illustrates the fraction of indels among all calls in _`r report_data$sample_name`_ (black dashed line) along with the distribution of indel fractions for TCGA samples (colorectal, endometrial, ovarian, stomach) with known MSI status assayed from mononucleotide markers ( _MSI.H_ = high microsatellite instability, _MSS_ = microsatellite stable):

```{r msi_indelfrac_plot, echo = F, fig.width=12, fig.height=5}
htmltools::br()
p <- report_data$msi_prediction_data$msi_plot + ggplot2::geom_vline(xintercept=as.numeric(report_data$msi_prediction_data$msi_stats$fracNonRepeatIndels), size=1.1,linetype=2) 
p

htmltools::br()
```


### Supporting evidence II: coding mutations in MSI-associated genes

<br>
```{r msi_gene_variants, echo=F, results='asis'}
if(nrow(report_data$msi_prediction_data$mmr_pol_df) > 0){
  DT::datatable(report_data$msi_prediction_data$mmr_pol_df, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip'))
}else{
  cat('<i>No variants found.</i>',sep='\n')
  cat('\n')
}
```
<br><br>
