## Settings & annotation sources

The annotated variants are based on the following underlying tools and knowledge resources:

```{r list_resources, echo=F,results='asis'}
for(n in names(pcg_report$pcgr_db_release)){
  if(n != 'gwas'){
    s <- paste0('* [',pcg_report$pcgr_db_release[[n]]$name,'](',pcg_report$pcgr_db_release[[n]]$url,') - ',pcg_report$pcgr_db_release[[n]]$description, ' (',pcg_report$pcgr_db_release[[n]]$version,')')
    cat(s,sep="\n")
  }
}

```

```{r on_off_settings, echo=F}
signatures_estimation <- 'ON'
msi_prediction <- 'ON'
mutational_burden <- 'ON'
if(pcg_report$pcgr_config$msi$msi == F){
  msi_prediction <- 'OFF'
}
if(pcg_report$pcgr_config$mutational_signatures$mutsignatures == F){
  signatures_estimation <- 'OFF'
}
if(pcg_report$pcgr_config$mutational_burden$mutational_burden == F){
  mutational_burden <- 'OFF'
}
```


The report is generated with _[PCGR](https://github.com/sigven/pcgr)_ version `r pcg_report[['pcgr_version']]`, using the following key settings:

* Genome assembly: __`r pcg_report[['genome_assembly']]`__
* Tumor type:  __`r pcg_report[['tumor_class']]`__
* Sequencing mode input (VCF): __`r pcg_report[['sequencing_mode']][['mode']]`__
* Coding target size (VCF): __`r pcg_report[['pcgr_config']][['mutational_burden']][['target_size_mb']]` Mb__
* Minimum sequencing depth (DP) tumor (VCF): __`r pcg_report[['pcgr_config']][['allelic_support']][['tumor_dp_min']]`__
* Minimum allelic fraction (AF) tumor (VCF): __`r pcg_report[['pcgr_config']][['allelic_support']][['tumor_af_min']]`__
* Minimum sequencing depth (DP) normal (VCF): __`r pcg_report[['pcgr_config']][['allelic_support']][['normal_dp_min']]`__
* Maximum allelic fraction (AF) normal (VCF): __`r pcg_report[['pcgr_config']][['allelic_support']][['normal_af_max']]`__
* Tier system (VCF): __`r pcg_report[['tier_model']]`__
* MSI prediction: __`r msi_prediction`__
* Mutational burden estimation: __`r mutational_burden`__
* Mutational signatures estimation: __`r signatures_estimation`__
* Report theme (Bootstrap): __`r pcg_report[['pcgr_config']][['visual']][['report_theme']]`__


