### Tier 2 - Other cancer mutation hotspots, curated mutations, or predicted driver mutations {.tabset}

```{r num_tier2, echo=F}
n_mutations_tier2 = NROW(pcg_report[['snv_indel']][['variant_display']][['tier2']][['hotspot']]) + NROW(pcg_report[['snv_indel']][['variant_display']][['tier2']][['curated_mutation']]) + NROW(pcg_report[['snv_indel']][['variant_display']][['tier2']][['predicted_driver']])
```
- A total of __`r n_mutations_tier2`__ unique, somatic variant(s) in the tumor sample are curated as disease-causing, predicted as driver mutations, or occur in known cancer mutation hotspots.
<br><br>

#### Known mutation hotspots

<br>

```{r tier2_hotspot, echo=F, results = 'asis'}
 cat('The table below lists variants detected as mutation hotspots according to <a href="http:/cancerhotspots.org/" target="_blank">cancerhotspots.org<a/>:',sep='\n')
if(NROW(pcg_report[['snv_indel']][['variant_display']][['tier2']][['hotspot']]) > 0){
  cat('\n')
  myOptions <- list(paging = F, searching=F,caching=F,buttons = c('csv','excel'),dom = 'Bfrtip')
  if(NROW(pcg_report[['snv_indel']][['variant_display']][['tier2']][['hotspot']]) > 10){
    myOptions <- list(paging = T,pageLength=10, searching=T,caching=F,buttons = c('csv','excel'),dom = 'Bfrtip')
  }
  DT::datatable(dplyr::select(pcg_report[['snv_indel']][['variant_display']][['tier2']][['hotspot']], MUTATION_HOTSPOT,dplyr::everything()), options = myOptions, escape=F, extensions = c("Buttons","Responsive")) %>%
  DT::formatStyle('MUTATION_HOTSPOT', backgroundColor = 'black', color='white')
}else{
  cat('<i>No variants found.</i>',sep='\n')
  cat('\n')
}
```

<br><br>

#### Other disease-causing mutations

<br>

```{r tier2_disease_causing, echo=F, results='asis'}

if(NROW(pcg_report[['snv_indel']][['variant_display']][['tier2']][['curated_mutation']]) > 0){
  cat('\n')
  cat('The table below lists all other variants in Tier 2 curated as disease-causing in the <a href="http://docm.genome.wustl.edu/" target="_blank">Database of Curated Mutations<a/>:',sep='\n')
  myOptions <- list(paging = F, searching=F,caching=F,buttons = c('csv','excel'),dom = 'Bfrtip')
  if(NROW(pcg_report[['snv_indel']][['variant_display']][['tier2']][['curated_mutation']]) > 10){
    myOptions <- list(paging = T,pageLength=10, searching=T,caching=F,buttons = c('csv','excel'),dom = 'Bfrtip')
  }
  curated_mutations <- dplyr::select(pcg_report[['snv_indel']][['variant_display']][['tier2']][['curated_mutation']], -MUTATION_HOTSPOT)
  DT::datatable(dplyr::select(curated_mutations, DOCM_DISEASE, dplyr::everything()), options = myOptions, escape=F, extensions = c("Buttons","Responsive")) %>%
  DT::formatStyle('DOCM_DISEASE', backgroundColor = 'black', color = 'white')
}else{
  cat('<i>No variants found.</i>',sep='\n')
  cat('\n')
}
```

<br><br>

#### Other predicted driver mutations

<br>

```{r tier2_predicted_driver, echo=F,results='asis'}

if(NROW(pcg_report[['snv_indel']][['variant_display']][['tier2']][['predicted_driver']]) > 0){
  cat('The table below lists all other variants in Tier 2 predicted as cancer driver mutations in <a href="https://www.intogen.org/downloads" target="_blank">Intogen\'s catalog of driver mutations</a>:',sep='\n')
  
  myOptions <- list(paging = F, searching=F,caching=F,buttons = c('csv','excel'),dom = 'Bfrtip')
  if(NROW(pcg_report[['snv_indel']][['variant_display']][['tier2']][['predicted_driver']]) > 10){
    myOptions <- list(paging = T,pageLength=10, searching=T,caching=F,buttons = c('csv','excel'),dom = 'Bfrtip')
  }
  predicted_drivers <- dplyr::select(pcg_report[['snv_indel']][['variant_display']][['tier2']][['predicted_driver']], -c(INTOGEN_DRIVER_MUT, MUTATION_HOTSPOT, DOCM_DISEASE, DOCM_LITERATURE))
  DT::datatable(dplyr::select(predicted_drivers, SYMBOL, CONSEQUENCE, dplyr::everything()),options = myOptions, escape=F, extensions = c("Buttons","Responsive")) %>%
  DT::formatStyle(columns = c('SYMBOL','CONSEQUENCE'), backgroundColor = 'black',color = 'white')
}else{
  cat('<i>No variants found.</i>',sep='\n')
  cat('\n')
}
```

<br><br>
