
### Tier 1 - Genomic biomarkers for diagnosis, prognosis, and drug response {.tabset}

- A total of __`r nrow(pcg_report[['snv_indel']][['variant_display']][['tier1']])`__ unique, somatic variant(s) in the tumor sample can be mapped to genomic biomarkers in the [database for clinical interpretations of variants in cancer, CIViC](https://civic.genome.wustl.edu/#/home) or [Cancer bioMarkers database](https://www.cancergenomeinterpreter.org/biomarkers), with the following number of evidence items:
    * Tier 1 - Predictive: __`r nrow(pcg_report[['snv_indel']][['clinical_evidence_item']][['any_tumortype']][['predictive']][['any']])`__ evidence items
    * Tier 1 - Prognostic: __`r nrow(pcg_report[['snv_indel']][['clinical_evidence_item']][['any_tumortype']][['prognostic']][['any']])`__ evidence items
    * Tier 1 - Diagnostic: __`r nrow(pcg_report[['snv_indel']][['clinical_evidence_item']][['any_tumortype']][['diagnostic']][['any']])`__ evidence items

<br><br>

```{r prepare_tier1_data, echo=F, results='asis'}

show_tier1_filters <- list()
missing_tier1_items <- list()
for(type in c('diagnostic','prognostic','predictive')){
  show_tier1_filters[[type]] <- F
  missing_tier1_items[[type]] <- T
  if(nrow(pcg_report[['snv_indel']][['clinical_evidence_item']][['any_tumortype']][[type]][['any']]) > 0){
    show_tier1_filters[[type]] <- T
    missing_tier1_items[[type]] <- F
  }
}

```

#### Predictive biomarkers

<br>

```{r tier1_predictive, echo=F, results = 'asis', eval = show_tier1_filters[['predictive']]}
library(summarywidget)
library(crosstalk)

variants_tier1_predictive_shared <- crosstalk::SharedData$new(pcg_report[['snv_indel']][['clinical_evidence_item']][['any_tumortype']][['predictive']][['any']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", variants_tier1_predictive_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical significance", variants_tier1_predictive_shared, ~CLINICAL_SIGNIFICANCE),
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", variants_tier1_predictive_shared, ~EVIDENCE_LEVEL)
  ),
  list(
    crosstalk::filter_select("CONSEQUENCE", "Consequence", variants_tier1_predictive_shared, ~CONSEQUENCE),
    crosstalk::filter_select("BIOMARKER_MAPPING", "Biomarker mapping", variants_tier1_predictive_shared, ~BIOMARKER_MAPPING),
    crosstalk::filter_select("THERAPEUTIC_CONTEXT", "Therapeutic context", variants_tier1_predictive_shared, ~THERAPEUTIC_CONTEXT)
  )
)

cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(variants_tier1_predictive_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip')) %>%
  DT::formatStyle('EVIDENCE_LEVEL', backgroundColor = DT::styleEqual(c('A: Validated','A: FDA/NCCN/ELN guidelines','B: Clinical evidence','B1: Clinical evidence: late trials','B2: Clinical evidence: early trials','C: Case study','D: Preclinical evidence','E: Indirect evidence'), c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))
)

#htmltools::br()

```

```{r tier1_predictive_missing, echo=F, results = 'asis', eval = missing_tier1_items[['predictive']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>

#### Prognostic biomarkers

<br>

```{r tier1_prognostic, echo=F, results='asis', eval = show_tier1_filters[['prognostic']]}

library(summarywidget)
library(crosstalk)

variants_tier1_prognostic_shared <- crosstalk::SharedData$new(pcg_report[['snv_indel']][['clinical_evidence_item']][['any_tumortype']][['prognostic']][['any']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CONSEQUENCE", "Consequence", variants_tier1_prognostic_shared, ~CONSEQUENCE),
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", variants_tier1_prognostic_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical significance", variants_tier1_prognostic_shared, ~CLINICAL_SIGNIFICANCE)
  ),
  list(
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", variants_tier1_prognostic_shared, ~EVIDENCE_LEVEL),
    crosstalk::filter_select("BIOMARKER_MAPPING", "Biomarker mapping", variants_tier1_prognostic_shared, ~BIOMARKER_MAPPING)
  )
)
cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(variants_tier1_prognostic_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip')) %>%
  DT::formatStyle('EVIDENCE_LEVEL', backgroundColor = DT::styleEqual(c('A: Validated','A: FDA/NCCN/ELN guidelines','B: Clinical evidence','B1: Clinical evidence: late trials','B2: Clinical evidence: early trials','C: Case study','D: Preclinical evidence','E: Indirect evidence'), c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))
)
#htmltools::br()

```

```{r tier1_prognostic_missing, echo=F, results = 'asis', eval = missing_tier1_items[['prognostic']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>

#### Diagnostic biomarkers

<br>

```{r tier1_diagnostic, echo=F, results='asis', eval = show_tier1_filters[['diagnostic']]}

library(summarywidget)
library(crosstalk)

variants_tier1_diagnostic_shared <- crosstalk::SharedData$new(pcg_report[['snv_indel']][['clinical_evidence_item']][['any_tumortype']][['diagnostic']][['any']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CONSEQUENCE", "Consequence", variants_tier1_diagnostic_shared, ~CONSEQUENCE),
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", variants_tier1_diagnostic_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical association", variants_tier1_diagnostic_shared, ~CLINICAL_SIGNIFICANCE)
  ),
  list(
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", variants_tier1_diagnostic_shared, ~EVIDENCE_LEVEL),
    crosstalk::filter_select("BIOMARKER_MAPPING", "Biomarker mapping", variants_tier1_diagnostic_shared, ~BIOMARKER_MAPPING)
  )
)
cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(variants_tier1_diagnostic_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip')) %>%
  DT::formatStyle('EVIDENCE_LEVEL', backgroundColor = DT::styleEqual(c('A: Validated','A: FDA/NCCN/ELN guidelines','B: Clinical evidence','B1: Clinical evidence: late trials','B2: Clinical evidence: early trials','C: Case study','D: Preclinical evidence','E: Indirect evidence'), c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))
)
#htmltools::br()

```

```{r tier1_diagnostic_missing, echo=F, results = 'asis', eval = missing_tier1_items[['diagnostic']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>


