## Somatic CNAs

<br>

```{r plot_cna}
#| eval: !expr NROW(pcg_report$content$cna$callset$variant) > 0
#| echo: false
#| output: asis
#| fig.width: 15
#| fig.height: 10

cna_plot <- pcgrr::plot_cna_segments(
  chrom_coordinates = pcg_report$settings$chrom_coordinates,
  cna_segment = pcg_report$content$cna$callset$segment,
  cna_gene = pcg_report$content$cna$callset$variant
)

p_margin <- 
  list(l = 50, r = 0, b = 0, t = 0, pad = 0)

plotly::ggplotly(cna_plot, tooltip = c("y","z")) |> 
  plotly::layout(
    legend = list(orientation = "h", x = 0.34, y = -0.28),
    margin = p_margin)

```

<br>

```{r check_expression}
#| eval: true

# expression_present_cna <- FALSE
# if("TPM_GENE" %in% colnames(pcg_report$content$cna$callset$variant) &
#    pcg_report$settings$molecular_data$fname_expression_tsv != "None"){
#   expression_present_cna <- 
#     !any(is.na(pcg_report$content$cna$callset$variant$TPM_GENE))
# }
```

### Variant classification - actionability

Genes subject to copy number aberrations in the tumor sample have been assessed for
_clinical actionability_ through an implementation of AMP/ASCO/CAP guidelines [@Li2017-ew]. We use CIViC and CGI [@Griffith2017-do;@Tamborero2018-aj] as the underlying sources for actionable variants (versions indicated below).

The tabset below indicates which variants have been classified into two distinct _tiers of clinical significance_ according to AMP/ASCO/CAP criteria.


::: {.callout-note}
## Note - biomarker matching

The same somatic mutation may frequently match with multiple evidence items listed in biomarker databases, pending on e.g. the _disease/tumor context_, _therapeutic context_, _biomarker resolution_, or underlying _evidence source_ (i.e. publication)
:::

::: {.panel-tabset .nav-pills}

#### <b>TIER I</b> - strong clinical significance

<br>

```{r}
#| echo: false
#| output: asis
#| eval: !expr pcg_report$content$cna$vstats$n_actionable_tier1 != 0


dt_content_tier1 <- 
  pcgrr:::get_dt_tables(
    rep = pcg_report, tier = 1, variant_class = "cna")

# if(expression_present_cna == FALSE){
#   dt_content_tier1[['by_gene']]$TPM_GENE <- NULL
# }

tier1_gene_dt_cna <- DT::datatable(
  dt_content_tier1[['by_gene']],
  escape = F, 
  options = list(dom = 't'),
  extensions = c("Responsive")) |>
  DT::formatStyle(
    'SYMBOL',
    color = "white",
    `text-align` = 'center',
    fontWeight = 'bold',
    backgroundColor = pcgrr::color_palette$bg_dark
  ) |>
  DT::formatStyle(
    'BIOMARKER_EVIDENCE',
    color = "white",
    backgroundColor = pcgrr::color_palette$bg_dark
  ) |>
  DT::formatStyle(
    'VARIANT_CLASS',
    color='white',
    fontWeight = 'bold',
    backgroundColor = DT::styleEqual(
      pcgrr::color_palette$cna_variant_class$levels,
      pcgrr::color_palette$cna_variant_class$values),
    `text-align` = 'center'
  )

# if(expression_present_cna == TRUE){
#   tier1_gene_dt_cna <- tier1_gene_dt_cna |>
#     DT::formatStyle(
#      "TPM_GENE",
#      "TPM_GENE", 
#      color = "white", 
#      backgroundColor = 
#        DT::styleInterval(
#          pcgrr::color_palette$gene_expression$breaks,
#          pcgrr::color_palette$gene_expression$values
#        )
#    )
# }

tier1_gene_dt_cna


htmltools::br()
htmltools::hr()
htmltools::br()
cat(paste0(
"<font style='font-size: 110%'>Details with respect to the clinical",
" evidence items associated with <b>",
 paste(dt_content_tier1[['by_gene']]$MOLECULAR_ALTERATION, collapse=", "),
 " </b>can be explored in the data table listed below:</font>\n"))
htmltools::br()
htmltools::br()

cna_eitem_tier1_shared <- crosstalk::SharedData$new(
  dt_content_tier1[['by_eitem']] |>
    dplyr::rename(SAMPLE_ALTERATION = MOLECULAR_ALTERATION))

crosstalk::bscols(
  list(
     crosstalk::filter_select(
      "SAMPLE_ALTERATION", "Sample alteration", 
      cna_eitem_tier1_shared, ~SAMPLE_ALTERATION),
    crosstalk::filter_select(
      "BM_CANCER_TYPE", "Cancer type", 
      cna_eitem_tier1_shared, ~BM_CANCER_TYPE),
    crosstalk::filter_select(
      "BM_CLINICAL_SIGNIFICANCE", "Clinical significance", 
      cna_eitem_tier1_shared, ~BM_CLINICAL_SIGNIFICANCE),
    crosstalk::filter_slider(
      "BM_RATING", "Rating", cna_eitem_tier1_shared, 
      ~BM_RATING, min = 0, max = 5, step = 1, ticks = T)

  ),
  list(
    crosstalk::filter_select(
      "BM_EVIDENCE_TYPE", "Evidence type", 
      cna_eitem_tier1_shared, ~BM_EVIDENCE_TYPE),
    crosstalk::filter_select(
      "BM_EVIDENCE_LEVEL", "Evidence level", 
      cna_eitem_tier1_shared, ~BM_EVIDENCE_LEVEL),
    crosstalk::filter_select(
      "BM_EVIDENCE_DIRECTION", "Evidence direction", 
      cna_eitem_tier1_shared, ~BM_EVIDENCE_DIRECTION),
    crosstalk::filter_select(
      "BM_THERAPEUTIC_CONTEXT", "Therapeutic context", 
      cna_eitem_tier1_shared, ~BM_THERAPEUTIC_CONTEXT)
  )
)

htmltools::br()

DT::datatable(
  cna_eitem_tier1_shared,
  escape = F, 
  extensions = c("Buttons","Responsive"), 
  options = list(
    buttons = c('csv','excel'), 
    dom = 'Bfrtip')) |>
  DT::formatStyle(
    color = "white", 
    c("SAMPLE_ALTERATION","BM_CONTEXT"), 
    fontWeight = 'bold', 
    `text-align` = 'center',
    backgroundColor = pcgrr::color_palette$bg_dark) |>
  DT::formatStyle(
    'BM_EVIDENCE_LEVEL', 
    backgroundColor = DT::styleEqual(
      pcgrr::color_palette$clinical_evidence$levels,
      pcgrr::color_palette$clinical_evidence$values))

```


```{r }
#| output: asis
#| echo: false
#| eval: !expr pcg_report$content$cna$vstats$n_actionable_tier1 == 0

htmltools::br()
cat('\n*  <i> <font style="font-size: 110%"><b>NO</b> somatic copy-number gene aberrations of strong clinical significance were identified.</font></i>', sep = '\n')
cat('\n')
htmltools::br()

```

#### <b>TIER II</b> - potential clinical significance

<br>

```{r}
#| echo: false
#| output: asis
#| eval: !expr pcg_report$content$cna$vstats$n_actionable_tier2 != 0

dt_content_tier2 <- 
  pcgrr:::get_dt_tables(
    rep = pcg_report, tier = 2, variant_class = "cna")

# if(expression_present_cna == FALSE){
#   dt_content_tier2[['by_gene']]$TPM_GENE <- NULL
# }

tier2_gene_dt_cna <- DT::datatable(
  dt_content_tier2[['by_gene']],
  escape = F, 
  options = list(dom = 't'),
  extensions = c("Responsive")) |>
  DT::formatStyle(
    'SYMBOL',
    color = "white",
    `text-align` = 'center',
    fontWeight = 'bold',
    backgroundColor = pcgrr::color_palette$bg_dark
  ) |>
  DT::formatStyle(
    'BIOMARKER_EVIDENCE',
    color = "white",
    backgroundColor = pcgrr::color_palette$bg_dark
  ) |>
  DT::formatStyle(
    'VARIANT_CLASS',
    color='white',
    fontWeight = 'bold',
    backgroundColor = DT::styleEqual(
      pcgrr::color_palette$cna_variant_class$levels,
      pcgrr::color_palette$cna_variant_class$values),
    `text-align` = 'center'
  )

# if(expression_present_cna == TRUE){
#   tier2_gene_dt_cna <- tier2_gene_dt_cna |>
#     DT::formatStyle(
#      "TPM_GENE",
#      "TPM_GENE", 
#      color = "white", 
#      backgroundColor = 
#        DT::styleInterval(
#          pcgrr::color_palette$gene_expression$breaks,
#          pcgrr::color_palette$gene_expression$values
#        )
#    )
# }

tier2_gene_dt_cna

htmltools::br()
htmltools::hr()
htmltools::br()
cat(paste0(
"<font style='font-size: 110%'>Details with respect to the clinical",
" evidence items associated with <b>",
 paste(dt_content_tier2[['by_gene']]$MOLECULAR_ALTERATION, collapse=", "),
 " </b>can be explored in the data table listed below:</font>\n"))
htmltools::br()

cna_eitem_tier2_shared <- crosstalk::SharedData$new(
  dt_content_tier2[['by_eitem']] |>
    dplyr::rename(SAMPLE_ALTERATION = MOLECULAR_ALTERATION))

crosstalk::bscols(
  list(
     crosstalk::filter_select(
      "SAMPLE_ALTERATION", "Sample alteration", 
      cna_eitem_tier2_shared, ~SAMPLE_ALTERATION),
    crosstalk::filter_select(
      "BM_CANCER_TYPE", "Cancer type", 
      cna_eitem_tier2_shared, ~BM_CANCER_TYPE),
    crosstalk::filter_select(
      "BM_CLINICAL_SIGNIFICANCE", "Clinical significance", 
      cna_eitem_tier2_shared, ~BM_CLINICAL_SIGNIFICANCE),
    crosstalk::filter_slider(
      "BM_RATING", "Rating", cna_eitem_tier2_shared, 
      ~BM_RATING, min = 0, max = 5, step = 1, ticks = T)

  ),
  list(
    crosstalk::filter_select(
      "BM_EVIDENCE_TYPE", "Evidence type", 
      cna_eitem_tier2_shared, ~BM_EVIDENCE_TYPE),
    crosstalk::filter_select(
      "BM_EVIDENCE_LEVEL", "Evidence level", 
      cna_eitem_tier2_shared, ~BM_EVIDENCE_LEVEL),
    crosstalk::filter_select(
      "BM_EVIDENCE_DIRECTION", "Evidence direction", 
      cna_eitem_tier2_shared, ~BM_EVIDENCE_DIRECTION),
    crosstalk::filter_select(
      "BM_THERAPEUTIC_CONTEXT", "Therapeutic context", 
      cna_eitem_tier2_shared, ~BM_THERAPEUTIC_CONTEXT)
  )
)

htmltools::br()

DT::datatable(
  cna_eitem_tier2_shared,
  escape = F, 
  extensions = c("Buttons","Responsive"), 
  options = list(
    buttons = c('csv','excel'), 
    dom = 'Bfrtip')) |>
  DT::formatStyle(
    color = "white", 
    c("SAMPLE_ALTERATION","BM_CONTEXT"), 
    fontWeight = 'bold', 
    `text-align` = 'center',
    backgroundColor = pcgrr::color_palette$bg_dark) |>
  DT::formatStyle(
    'BM_EVIDENCE_LEVEL', 
    backgroundColor = DT::styleEqual(
      pcgrr::color_palette$clinical_evidence$levels,
      pcgrr::color_palette$clinical_evidence$values))

```


```{r }
#| output: asis
#| echo: false
#| eval: !expr pcg_report$content$cna$vstats$n_actionable_tier2 == 0

htmltools::br()
cat('\n*  <i> <font style="font-size: 110%"><b>NO</b> somatic copy-number gene aberrations of potential clinical significance were identified.</font></i>', sep = '\n')
cat('\n')
htmltools::br()

```

:::

<br>
<hr>
<br>

### Other oncogenic events

Here we show other genes subject to copy number alteration that may potentially have oncogenic effects. Specifically, we list

 * [Proto-oncogenes](#documentation) subject to copy-number gain (total copy number >= __`r pcg_report$settings$conf$somatic_cna$n_copy_gain`__ (gain threshold, user-configurable))
 * [Tumor suppressor](#documentation) genes subject to homozygous deletions (total copy number = __0__)

Gene symbols are color-coded according to their strength of association to cancer (here, specifically __`r pcg_report$settings$conf$sample_properties$site`__ cancers), from <span style="background-color:`r pcgrr::color_palette$cancer_assoc$values[1]`; padding:2x; color:white"><b>&nbsp; weak associations &nbsp; </b></span>&nbsp; to <span style="background-color:`r pcgrr::color_palette$cancer_assoc$values[length(pcgrr::color_palette$cancer_assoc$values)]`; padding:2x; color:white"><b>&nbsp; strong associations &nbsp; </b></span> [@Ochoa2021-jc;@Nakken2023-ol].


<br>

```{r }


dt_other_oncogenic <- pcgrr::get_oncogenic_cna_events(
  cna_df_display = pcg_report[["content"]][["cna"]][["callset"]][["variant_display"]])

dt_other_oncongenic_cna <- DT::datatable(
  dt_other_oncogenic,
  escape = F, 
  extensions = c("Buttons","Responsive"), 
  options = list(
    buttons = c('csv','excel'), 
    dom = 'Bfrtip')) |>
   DT::formatStyle(
    'VARIANT_CLASS',
    color='white',
    backgroundColor = DT::styleEqual(
      pcgrr::color_palette$cna_variant_class$levels,
      pcgrr::color_palette$cna_variant_class$values),
    `text-align` = 'center'
  ) |>
   DT::formatStyle(
     "SYMBOL",
     "TISSUE_ASSOC_RANK", 
     color = "white", 
     backgroundColor = 
       DT::styleInterval(
         pcgrr::color_palette$cancer_assoc$breaks,
         pcgrr::color_palette$cancer_assoc$values
       )
   )

if(pcg_report$settings$conf$sample_properties$site == "Any"){
  dt_other_oncongenic_cna <- dt_other_oncongenic_cna |>
    DT::formatStyle(
    "SYMBOL",
    "GLOBAL_ASSOC_RANK", 
    color = "white", 
    backgroundColor = 
      DT::styleInterval(
        pcgrr::color_palette$cancer_assoc$breaks,
        pcgrr::color_palette$cancer_assoc$values
      )
  )
}


dt_other_oncongenic_cna

```



