
Overview {data-navmenu="sCNA"}
================================================================== 

Row
-------------------------------------

### Overview

```{r cna_vbox_overview}
flexdashboard::valueBox("sCNA", caption = NULL, color = "#000000", icon = NULL)
```

### Copy number gains

```{r cna_gains, echo = F}
flexdashboard::valueBox(
  pcg_report$content$cna$vstats$n_segments_gain,
  color = ifelse(
    pcg_report$content$cna$vstats$n_segments_gain > 0,
    pcg_report$settings$conf$report_color,
    pcgrr::color_palette$none),
  icon = "fa-dna")
```

### Copy number losses

```{r cna_losses, echo = F}
flexdashboard::valueBox(
  pcg_report$content$cna$vstats$n_segments_loss,
  color = ifelse(
    pcg_report$content$cna$vstats$n_segments_loss > 0,
    pcg_report$settings$conf$report_color,
    pcgrr::color_palette$none),
  icon = "fa-dna")
```

### TIER 1 biomarkers

```{r cna_tier1, echo = F}
flexdashboard::valueBox(
  pcg_report$content$cna$vstats$n_tier1, 
  color = ifelse(
    pcg_report$content$cna$vstats$n_tier1 > 0,
    pcgrr::color_palette$success,
    pcgrr::color_palette$none), 
  href = "#tier-1-1", icon = "fa-file-prescription")
```

### TIER 2 biomarkers

```{r cna_tier2, echo = F}
flexdashboard::valueBox(
  pcg_report$content$cna$vstats$n_tier2, 
  color = ifelse(
    pcg_report$content$cna$vstats$n_tier2 > 0,
    pcgrr::color_palette$success,
    pcgrr::color_palette$none), 
  href = "#tier-2-1", 
  icon = "fa-file-prescription")

```



Row {data-height=700} 
-------------------------------------

### Copy number aberration plot {data-padding=15}

```{r cna_segment_plot, echo=F, fig.width=20, results="asis"}
plotly::ggplotly(CNAqc::plot_segments(
  pcg_report$content$cna$cnaqc, cex=2
))

```

Row 
-------------------------------------

### Key findings {data-width=300}

<br><br>

<font style="color:white; font-size: 115%; font-weight: bold">

  **  [ - Proto-oncogenes subject to amplifications: `r pcg_report$content$cna$vstats$n_oncogene_gain`](#scna-oncogene)**

  **  [ - Tumor suppressor genes subject to homozygous deletions: `r pcg_report$content$cna$vstats$n_tsgene_loss`](#scna-tsgene)**

  **  [ - Other drug targets subject to amplification: `r pcg_report$content$cna$vstats$n_other_drugtarget_gain`](#scna-other)**

</font>

### Documentation {data-padding=20}

Somatic copy number aberrations identified in the tumor sample are classified into **two main tiers**:

- [**TIER 1: Aberrations of strong clinical significance**](#tier-1-1) - constitutes amplified/lost genes linked to predictive, prognostic, or diagnostic biomarkers in the [CIViC database](http://civic.genome.wustl.edu) and the [Cancer Biomarkers Database](https://www.cancergenomeinterpreter.org/biomarkers) that are
	 - Found within the same tumor type/class as specified by the user, **AND**
	 - Of strong clinical evidence (i.e. part of guidelines, validated or discovered in late clinical trials ([CIViC evidence levels A/B](https://civic.readthedocs.io/en/latest/model/evidence/level.html)))
	 	 
- [**TIER 2: Aberrations of potential clinical significance**](#tier-2-1) - constitutes amplified/lost genes linked to predictive, prognostic, or diagnostic biomarkers in the [CIViC database](http://civic.genome.wustl.edu) and the [Cancer Biomarkers Database](https://www.cancergenomeinterpreter.org/biomarkers) that are either
	 - Of strong clinical evidence in other tumor types/classes than the one specified by the user, **OR**
	 - Of weak clinical evidence (early trials, case reports etc. ([CIViC evidence levels C/D/E](https://civic.readthedocs.io/en/latest/model/evidence/level.html)))) in the same tumor type/class as specified by the user
	 
Included in the report is also a complete list of [**all oncogenes subject to amplifications**](#scna-oncogene), [**tumor suppressor genes subject to homozygous deletions**](#scna-tsgene), and [**other drug targets subject to amplifications**](#scna-other)

- Genes are annotated as oncogenes and/or tumor suppressors genes if they
       i) have an associated record in curated databases ([Network of Cancer Genes (NCG)](http://ncg.kcl.ac.uk/)/[Cancer Gene Census](https://www.sanger.ac.uk/data/cancer-gene-census/)), **OR**,
       ii) have evidence from high-throughput literature mining (at least 20 publications, as collected from the [CancerMine text-mining resource](http://bionlp.bcgsc.ca/cancermine/)) 
     
