Clinical trials
==================================================================

Row
-------------------------------------

### Molecularly targeted trials

```{r}
flexdashboard::valueBox("Clinical trials (Beta)", caption = NULL, color = '#000000', icon = NULL)
n_not_recruiting <- pcg_report[['content']][['clinicaltrials']][['trials']] %>% 
  dplyr::filter(OVERALL_STATUS == "Not yet recruiting") %>% NROW()

n_recruiting <- pcg_report[['content']][['clinicaltrials']][['trials']] %>% 
  dplyr::filter(OVERALL_STATUS == "Recruiting") %>% NROW()

n_enrolling <- pcg_report[['content']][['clinicaltrials']][['trials']] %>% 
  dplyr::filter(OVERALL_STATUS == "Enrolling by invitation") %>% NROW()

n_active <- pcg_report[['content']][['clinicaltrials']][['trials']] %>% 
  dplyr::filter(OVERALL_STATUS == "Active, not recruiting") %>% NROW()

n_unknown <- pcg_report[['content']][['clinicaltrials']][['trials']] %>% 
  dplyr::filter(OVERALL_STATUS == "Unknown status") %>% NROW()
``` 

### Not yet recruiting

```{r}


flexdashboard::valueBox(n_not_recruiting, color = dplyr::if_else(
  n_not_recruiting > 0,
  pcg_report[['metadata']][['color_value_box']],
  pcg_report[['metadata']][['color_none']]))


```

### Recruiting

```{r}

flexdashboard::valueBox(n_recruiting, color = dplyr::if_else(
  n_recruiting > 0,
  pcg_report[['metadata']][['color_value_box']],
  pcg_report[['metadata']][['color_none']]))

```


### Enrolling by invitation

```{r}

flexdashboard::valueBox(n_enrolling, color = dplyr::if_else(
  n_enrolling > 0,
  pcg_report[['metadata']][['color_value_box']],
  pcg_report[['metadata']][['color_none']]))
```

### Active, not recruiting

```{r}
flexdashboard::valueBox(n_active, color = dplyr::if_else(
  n_active > 0,
  pcg_report[['metadata']][['color_value_box']],
  pcg_report[['metadata']][['color_none']]))

```

### Unknown status

```{r}
flexdashboard::valueBox(n_unknown, color = dplyr::if_else(
  n_unknown > 0,
  pcg_report[['metadata']][['color_value_box']],
  pcg_report[['metadata']][['color_none']]))
```


Row
-----------------------------------------------------

### Molecularly targeted trials - filters {data-padding=15 data-width=300}

* Ongoing or planned clinical trials in the relevant tumor type have been retrieved from [clinicaltrials.gov](https://clinicaltrials.gov), focusing on the subset with molecularly targeted therapies
* Key information entities (interventions/drugs, conditions) in trial records have been mapped to established thesauri ([ChEMBL](https://www.ebi.ac.uk/chembl/), [NCI Thesaurus](https://ncithesaurus.nci.nih.gov/ncitbrowser/), [UMLS/MedGen](https://www.ncbi.nlm.nih.gov/medgen/))
* Results from a text-mining procedure on unstructured trial text (e.g. <i>inclusion/exclusion criteria</i>) attempts to highlight the presence of established molecular biomarkers in cancer and relevant therapeutic contexts. 

<br>

```{r table_browse_trials, echo=F, results = "asis", eval = !pcg_report[['content']][['clinicaltrials']][['missing_data']]}


if(NROW(pcg_report[['content']][['clinicaltrials']][['trials']]) > 2000){
  cat('<b>NOTE - only considering top 2000 trials (ranked according to status and start date) </b><br>',sep="\n")
  cat('<br>')
  pcg_report[['content']][['clinicaltrials']][['trials']] <- 
    head(pcg_report[['content']][['clinicaltrials']][['trials']], 2000)
}

trials_ttype <- crosstalk::SharedData$new(pcg_report[['content']][['clinicaltrials']][['trials']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CONDITION_RAW", "Condition (cancer subtype)", trials_ttype, ~CONDITION_RAW),
    crosstalk::filter_select("OVERALL_STATUS", "Status", trials_ttype, ~OVERALL_STATUS),
    crosstalk::filter_select("INTERVENTION_RAW", "Drug(s)", trials_ttype, ~INTERVENTION_RAW),
    crosstalk::filter_select("INTERVENTION_TARGET", "Drug target(s)", trials_ttype, ~INTERVENTION_TARGET),
    crosstalk::filter_select("KEYWORD", "Therapeutic context mentions (text-mined)", trials_ttype, ~KEYWORD),
    crosstalk::filter_select("BIOMARKER_INDEX", "Biomarker mentions (text-mined)", trials_ttype, ~BIOMARKER_INDEX)

  ),
  list(
    crosstalk::filter_select("PHASE","Phase",trials_ttype,~PHASE),
    crosstalk::filter_checkbox("GENDER", "Gender", trials_ttype, ~GENDER),
    crosstalk::filter_slider("MINIMUM_AGE", "Minimum age", trials_ttype, ~MINIMUM_AGE),
    crosstalk::filter_slider("MAXIMUM_AGE", "Maximum age", trials_ttype, ~MAXIMUM_AGE),
    crosstalk::filter_select("METASTASES_INDEX", "Metastases mentions (text-mined)", trials_ttype, ~METASTASES_INDEX)
  )
)


```


```{r trials_missing_filters, echo=F, results = 'asis', eval = pcg_report[['content']][['clinicaltrials']][['missing_data']]}
cat('\n*  <i> <font style="font-size: 110%"><b>No</b> molecularly targeted trials retrieved for the tumor type in question.</i></font>', sep='\n')
cat('\n')
```


### Molecularly targeted trials {data-padding=15}

<br>

```{r trials_table_all, eval = !pcg_report[['content']][['clinicaltrials']][['missing_data']]}

if(NROW(trials_ttype) > 2000){
  cat('<b>NOTE - only considering top 2000 trials (ranked according to status and start date) </b><br>',sep="\n")
  cat('<br>')
}


trials_ttype %>%
  DT::datatable(escape = F, 
                width = "100%",
                extensions = c("Buttons","Responsive"), 
                options = list(pageLength = 8, 
                               scrollY = scrollY_flexdb_container,
                               buttons = c('csv','excel'), 
                                fixedColumns = T,
                 fixedHeader = T,
                 #scrollX = F,
                  columnDefs = list(list(width = '110px', targets = c(1,3,5)),
                                    list(width = '200px', targets = c(2,4))),
                               dom = 'Bfrtip')) %>%
  DT::formatStyle("OVERALL_STATUS",color="white", `text-align` = 'center', fontWeight = "bold",
                  backgroundColor = DT::styleEqual(c('Recruiting',
                                                     'Not yet recruiting',
                                                     'Active, not recruiting',
                                                     'Enrolling by invitation',
                                                     'Completed',
                                                     'Suspended',
                                                     'Withdrawn',
                                                     'Unknown status'), c("#00a65a","#00a65a","#CD534C","#CD534C","#CD534C","#CD534C","#CD534C", "#8F7700")))

  # DT::formatStyle(color="white", "SYMBOL", "BIOMARKER_MAPPING", fontWeight = 'bold', `text-align` = 'center',
  #                 backgroundColor = DT::styleEqual(c('exact','codon','exon','gene'), 
  #                                                  c('#000','#000',pcg_report[['metadata']][['color_palette']][['warning']][['values']][1], pcg_report[['metadata']][['color_palette']][['warning']][['values']][1])))

```


```{r trials_missing_data, echo=F, results = 'asis', eval = pcg_report[['content']][['clinicaltrials']][['missing_data']]}
cat('\n*  <i> <font style="font-size: 110%"><b>No</b> molecularly targeted trials retrieved for the tumor type in question .</i></font>',sep='\n')
cat('\n')
```

