
Overview {data-navmenu="sCNA"}
================================================================== 

Row
-------------------------------------

### Overview

```{r cna_vbox_overview}
flexdashboard::valueBox("sCNA", caption = NULL, color = "#000000", icon = NULL)
```

### Copy number gains

```{r cna_gains, echo = F}
flexdashboard::valueBox(pcg_report$content$cna$v_stat$n_cna_gain,
                        color = ifelse(pcg_report$content$cna$v_stat$n_cna_gain > 0,
                                       pcg_report$metadata$color_value_box,
                                       pcg_report$metadata$color_none),
                        icon = "fa-dna")
```

### Copy number losses

```{r cna_losses, echo = F}
flexdashboard::valueBox(pcg_report$content$cna$v_stat$n_cna_loss,
                        color = ifelse(pcg_report$content$cna$v_stat$n_cna_loss > 0,
                                       pcg_report$metadata$color_value_box,
                                       pcg_report$metadata$color_none),
                        icon = "fa-dna")
```

### TIER 1 biomarkers

```{r cna_tier1, echo = F}
num_bms1 <- NROW(pcg_report[["content"]][["cna"]][["disp"]][["tier1"]])
flexdashboard::valueBox(num_bms1, color = ifelse(num_bms1 > 0,
                        "#00a65a",
                        pcg_report$metadata$color_none), href = "#tier-1-1", icon = "fa-file-prescription")
```

### TIER 2 biomarkers

```{r cna_tier2, echo = F}
num_bms2 <- NROW(pcg_report[["content"]][["cna"]][["disp"]][["tier2"]])
flexdashboard::valueBox(num_bms2, color = ifelse(num_bms2 > 0,
                        "#00a65a",
                        pcg_report$metadata$color_none), href = "#tier-2-1", icon = "fa-file-prescription")
```



Row {data-height=700} 
-------------------------------------

### Copy number segments - filters {data-width=300 data-padding=15}

The following user-defined thresholds determine copy number aberrations shown here: 

* <span style="background-color:#00a65a; color:white"><b>&nbsp; Copy number amplifications </b></span>&nbsp; : Log(2) ratio >= **`r pcg_report[["metadata"]][["config"]][["cna"]][["log_r_gain"]]`**

* <span style="background-color:#CD534CFF; color:white"><b>&nbsp; Homozygous deletions </b></span>&nbsp; : Log(2) ratio <= **`r pcg_report[["metadata"]][["config"]][["cna"]][["log_r_homdel"]]`**

A total of __`r NROW(pcg_report[["content"]][["cna"]][["disp"]][["segment"]])`__ unfiltered aberration segments satisfied the above criteria.



```{r cna_segments_setup, echo=F}
show_segment_filter_table <- F
if(NROW(pcg_report[["content"]][["cna"]][["disp"]][["segment"]]) > 0){
  show_segment_filter_table <- T
}

```


```{r cna_segments_sd_setup, echo = F, results = "asis"}

cna_segments_sd <- NULL
cna_segments <- data.frame()

if(NROW(pcg_report[["content"]][["cna"]][["disp"]][["segment"]]) > 0){
  cna_segments <- pcg_report[["content"]][["cna"]][["disp"]][["segment"]]
  cna_segments_sd <- crosstalk::SharedData$new(cna_segments)
}else{
  cna_segments_sd <- crosstalk::SharedData$new(cna_segments)
}

```


```{r cna_segments_ampl_del, echo=F, results = "asis", eval = show_segment_filter_table}

crosstalk::bscols(
  list(
    crosstalk::filter_slider("LogR", "Log-ratio", cna_segments_sd, ~LOG_R, ticks = F),
    crosstalk::filter_select("Cytoband", "Cytoband", cna_segments_sd, ~CYTOBAND),
    crosstalk::filter_checkbox("Event", "Event type", cna_segments_sd, ~EVENT_TYPE)
  )
)
```

### Copy number segments {data-padding=20}

```{r cna_segments_show, echo = F, eval = show_segment_filter_table}
DT::datatable(cna_segments_sd, extensions=c("Buttons","Responsive"),
              escape = F,
              width = "100%",
              options = list(pageLength = 10, 
                             scrollCollapse = T, 
                             buttons = c("csv","excel"),
                             #autoWidth = T, 
                             #fixedColumns = T,
                             #fixedHeader = T,
                             #scrollX = T,
                             columnDefs = list(list(width = '20%', targets = c(1,2,3,4,5))),
                             dom = "Bfrtip")) %>%
DT::formatStyle("LOG_R",color="white", `text-align` = 'center', fontWeight = "bold", backgroundColor = DT::styleInterval(pcg_report[["metadata"]][["config"]][["cna"]][["log_r_homdel"]], c("#CD534CFF", "#00a65a")))


```


```{r scna_missing_global, echo=F, results = 'asis', eval = NROW(pcg_report$content$cna$disp$segment) == 0}
cat('\n*  <i> <font style="font-size: 110%"><b>No</b> somatic copy-number aberrations were detected.</font></i>', sep = '\n')
cat('\n')
```


Row 
-------------------------------------

### Key findings {data-width=300}

<br><br>

<font style="color:white; font-size: 115%; font-weight: bold">

  **  [ - Proto-oncogenes subject to amplifications: `r NROW(pcg_report[["content"]][["cna"]][["disp"]][["oncogene_gain"]])`](#scna-oncogene)**

  **  [ - Tumor suppressor genes subject to homozygous deletions: `r NROW(pcg_report[["content"]][["cna"]][["disp"]][["tsgene_loss"]])`](#scna-tsgene)**

  **  [ - Other drug targets subject to amplification: `r NROW(pcg_report[["content"]][["cna"]][["disp"]][["other_target"]])`](#scna-other)**

</font>

### Documentation {data-padding=20}

Somatic copy number aberrations identified in the tumor sample are classified into **two main tiers**:

- [**TIER 1: Aberrations of strong clinical significance**](#tier-1-1) - constitutes amplified/lost genes linked to predictive, prognostic, or diagnostic biomarkers in the [CIViC database](http://civic.genome.wustl.edu) and the [Cancer Biomarkers Database](https://www.cancergenomeinterpreter.org/biomarkers) that are
	 - Found within the same tumor type/class as specified by the user, **AND**
	 - Of strong clinical evidence (i.e. part of guidelines, validated or discovered in late clinical trials ([CIViC evidence levels A/B](https://civic.readthedocs.io/en/latest/model/evidence/level.html)))
	 	 
- [**TIER 2: Aberrations of potential clinical significance**](#tier-2-1) - constitutes amplified/lost genes linked to predictive, prognostic, or diagnostic biomarkers in the [CIViC database](http://civic.genome.wustl.edu) and the [Cancer Biomarkers Database](https://www.cancergenomeinterpreter.org/biomarkers) that are either
	 - Of strong clinical evidence in other tumor types/classes than the one specified by the user, **OR**
	 - Of weak clinical evidence (early trials, case reports etc. ([CIViC evidence levels C/D/E](https://civic.readthedocs.io/en/latest/model/evidence/level.html)))) in the same tumor type/class as specified by the user
	 
Included in the report is also a complete list of [**all oncogenes subject to amplifications**](#scna-oncogene), [**tumor suppressor genes subject to homozygous deletions**](#scna-tsgene), and [**other drug targets subject to amplifications**](#scna-other)

- Status as oncogenes and/or tumor suppressors genes are done according to the following scheme in PCGR:
    - Five or more publications in the biomedical literature that suggests an oncogenic/tumor suppressor role for a given gene (as collected from the [CancerMine text-mining resource](http://bionlp.bcgsc.ca/cancermine/)), **OR**
    - At least two publications from CancerMine that suggests an oncogenic/tumor suppressor role for a given gene **AND** an existing record for the same gene as a tumor suppressor/oncogene in the [Network of Cancer Genes (NCG)](http://ncg.kcl.ac.uk/)
    - Status as oncogene is ignored if a given gene also has literature evidence support for a role as a tumor suppressor gene which is three times as large (and vice versa)
    - Oncogenes/tumor suppressor candidates from CancerMine/NCG that are found in the [curated list of false positive cancer drivers compiled by Bailey et al. (Cell, 2018)](https://www.ncbi.nlm.nih.gov/pubmed/30096302) have been excluded
     
