### Signature similarity

Here, we perform a comparison of the sample's raw mutational spectrum to each of the known signatures (SBS) in COSMIC. Input samples with less than 30 SNVs are omitted from this analysis. The cosine similarity is calculated between the mutational spectrum of the sample (i.e. frequency of DNA trinucleotide contexts) and each of the COSMIC signatures. The cosine similarity (*SIMILARITY* column below) ranges from 0 to 1, where 1 indicates a perfect match.

```{r highlight_signatures}
#| echo: false
#| eval: !expr as.logical(pcg_report$settings$conf$somatic_snv$mutational_signatures$all_reference_signatures) == FALSE & as.logical(pcg_report$settings$conf$somatic_snv$mutational_signatures$no_prevalence_data) == FALSE
#| output: asis

cat('\n::: {.callout-note}\n## Site-specific signatures\n\n',
    'Signatures that are previously attributed to ',
    conf$sample_properties$site,
    ' cancers (if any available, with prevalence >= ', 
    as.character(msig_conf$prevalence_reference_signatures),
    '%) are highlighted. If you see non site-attributed signatures ',
    'with very high similarity to the mutational spectrum of the input sample, a re-run of the fitting procedure using all ',
    'reference signatures is warranted. \n\n:::\n\n', sep='')
```

<br>

```{r signature_similarity}
#| eval: !expr is.null(msig_content$result$signature_similarity) == FALSE
#| output: asis

sim_data    <- msig_content$result$signature_similarity

## format similarity to 2 decimal places
sim_data <- sim_data |>
  dplyr::mutate(SIMILARITY = sprintf("%.2f", SIMILARITY))

report_color <- pcg_report$settings$conf$report_color
none_color   <- pcgrr::color_palette$none

site_specific_style <- function(data) {
  function(value, index) {
    ss <- data$SITE_SPECIFIC[index]
    bg <- if (isTRUE(ss == "YES")) report_color else none_color
    list(background = bg, color = "white", fontWeight = "bold", padding = "4px 8px")
  }
}

col_defs <- list(
  SIGNATURE_ID = reactable::colDef(
    name     = "Signature",
    style    = site_specific_style(sim_data),
    minWidth = 110
  ),
  SIMILARITY = reactable::colDef(
    name     = "Similarity",
    minWidth = 100,
    align    = "center",
    style    = list(fontWeight = "bold")
  ),
  AETIOLOGY_KEYWORD = reactable::colDef(
    name     = "Aetiology",
    style    = site_specific_style(sim_data),
    minWidth = 150
  ),
  SITE_SPECIFIC = reactable::colDef(
    name     = "Site-specific signature")
)

reactable::reactable(
  sim_data,
  columns         = col_defs,
  filterable      = FALSE,
  searchable      = TRUE,
  highlight       = TRUE,
  compact         = TRUE,
  wrap            = TRUE,
  defaultPageSize = 10,
  theme           = pcgrr::rt_theme())
# 
# bslib::card(
#   height = "700px",
#   bslib::card_header(
#     class = "bg-dark",
#     paste0(
#       "Similarity of ",
#       pcg_report$settings$sample_id,
#       " to COSMIC signatures (SBS)")
#   ),
#   bslib::card_body(
#     similarity_table
#   )
# )



```

<br>
