#### Signature similarity

Here, we perform a comparison of the sample's raw mutational spectrum to each of the known signatures (SBS) in COSMIC. Input samples with less than 30 SNVs are omitted from this analysis. The cosine similarity is calculated between the mutational spectrum of the sample (i.e. frequency of DNA trinucleotide contexts) and each of the COSMIC signatures. The cosine similarity (*SIMILARITY* column below) ranges from 0 to 1, where 1 indicates a perfect match.

```{r highlight_signatures}
#| echo: false
#| eval: !expr as.logical(pcg_report$settings$conf$somatic_snv$mutational_signatures$all_reference_signatures) == FALSE & as.logical(pcg_report$settings$conf$somatic_snv$mutational_signatures$no_prevalence_data) == FALSE
#| output: asis

cat('\n::: {.callout-note}\n## Site-specific signatures\n\n',
    'Signatures that are previously attributed to ',
    conf$sample_properties$site,
    ' cancers (if any available, with prevalence >= ', 
    as.character(msig_conf$prevalence_reference_signatures),
    '%) are highlighted. If you see non site-attributed signatures ',
    'with very high similarity to the mutational spectrum of the input sample, a re-run of the fitting procedure using all ',
    'reference signatures is warranted. \n\n:::\n\n', sep='')
```

<br>

```{r signature_similarity}
#| eval: !expr is.null(msig_content$result$signature_similarity) == FALSE
#| output: asis

similarity_table <- 
  DT::datatable(
    msig_content$result$signature_similarity,
    extensions = "Responsive",
    options = list(
      pageLength = 13,
      dom = 'tp'),
    escape = F) |>
  DT::formatStyle(
    'SIMILARITY',
    fontWeight = 'bold') |>
  DT::formatStyle(
    c('SIGNATURE_ID',
      'AETIOLOGY_KEYWORD'),
    'SITE_SPECIFIC',
    fontWeight = 'bold',
    color = "white",
    backgroundColor = DT::styleEqual(
      c('YES','NO','NOT_DEFINED'), 
      c(pcg_report$settings$conf$report_color,
        pcgrr::color_palette$none,
        pcgrr::color_palette$none)
    )
  )

#similarity_table$x$data$SITE_SPECIFIC <- NULL

bslib::card(
  height = "700px",
  bslib::card_header(
    class = "bg-dark",
    paste0(
      "Similarity of ",
      pcg_report$settings$sample_id,
      " to COSMIC signatures (SBS)")
  ),
  bslib::card_body(
    similarity_table
  )
)



```

<br>
