### Similarity to other tumor samples

Here we show how the gene expression profile of the input RNA sample correlates with samples harvested from other sources. The similarity is calculated using the Spearman's correlation (protein-coding genes only). The tables below show the correlation coefficient for each (external) sample, along with the corresponding diagnosis, tissue/site, sample type, age category etc. The tables are sorted by the correlation coefficient in descending order, and only the top 200 samples are shown per source.

::: {.panel-tabset}

```{r similarity_table_tcga}
#| output: asis
#| eval: !expr as.logical(TRUE %in% stringr::str_detect(names(pcg_report[['content']]$expression[['similarity_analysis']]), 'tcga')) == TRUE

source_name <- 'Primary tumors - TCGA'
cat('\n\n#### ',source_name,'\n\n<br>')

dt_similarity_tcga <- data.frame()

for(cohort in names(pcg_report[['content']]$expression[['similarity_analysis']])){
  if(grepl('tcga', cohort)){
    if(NROW(pcg_report[['content']]$expression[['similarity_analysis']][[cohort]]) > 0){
      tcga_cohort <- 
        pcg_report[['content']]$expression[['similarity_analysis']][[cohort]] |>
        dplyr::mutate(CORR = round(CORR, digits = 5)) |>
        dplyr::select(-c("EXT_SAMPLE_ID2","PROTEIN_CODING_ONLY")) |>
        dplyr::mutate(
          EXT_SAMPLE_NAME = 
            glue::glue(
              "<a href='{EXT_LINK}' target='_blank'>{EXT_SAMPLE_NAME}</a>")
        ) |>
        dplyr::select(-c("EXT_LINK")) |>
        dplyr::select(
          EXT_SAMPLE_ID,
          EXT_SAMPLE_NAME,
          CORR,
          EXT_PRIMARY_DIAGNOSIS,
          EXT_PRIMARY_SITE,
          EXT_SAMPLE_TYPE,
          EXT_AGE_CATEGORY,
          EXT_GENDER,
          dplyr::everything()
        )
      
      dt_similarity_tcga <- dt_similarity_tcga |>
        dplyr::bind_rows(tcga_cohort)
    }
  }
}

dt_similarity_tcga <- dt_similarity_tcga |>
  dplyr::arrange(desc(CORR)) |>
  head(200)

highlighted_site <- pcg_report$settings$conf$sample_properties$site
all_sites <- unique(dt_similarity_tcga$EXT_PRIMARY_SITE)
other_sites <- all_sites[ !all_sites == highlighted_site]

DT::datatable(
  dt_similarity_tcga,
  escape = F,
  options = list(pageLength = 10),
  extensions=c("Buttons","Responsive")) |>
  DT::formatStyle(
    c('EXT_PRIMARY_SITE'),
    fontWeight = 'bold',
    color = "white",
    backgroundColor = DT::styleEqual(
      c(highlighted_site, other_sites), 
      c(pcg_report$settings$conf$report_color,
        rep(pcgrr::color_palette$none, length(other_sites)))
    )
  )

```


```{r similarity_table_depmap}
#| output: asis
#| eval: !expr as.logical("depmap" %in% names(pcg_report[['content']]$expression[['similarity_analysis']])) == TRUE

source_name <- 'Cell lines - DepMap'
cat('\n\n#### ',source_name,'\n\n<br>')

dt_similarity_depmap <- 
  pcg_report[['content']]$expression[['similarity_analysis']][['depmap']] |>
  dplyr::mutate(CORR = round(CORR, digits = 5)) |>
  dplyr::select(-c("EXT_SAMPLE_ID2","PROTEIN_CODING_ONLY")) |>
  dplyr::mutate(
    EXT_SAMPLE_NAME = 
      glue::glue(
        "<a href='{EXT_LINK}' target='_blank'>{EXT_SAMPLE_NAME}</a>")
  ) |>
  dplyr::select(-c("EXT_LINK")) |>
  dplyr::select(
    EXT_SAMPLE_ID,
    EXT_SAMPLE_NAME,
    CORR,
    EXT_PRIMARY_DIAGNOSIS,
    EXT_PRIMARY_SITE,
    EXT_SAMPLE_TYPE,
    EXT_AGE_CATEGORY,
    EXT_GENDER,
    dplyr::everything()
  ) |>
  dplyr::arrange(desc(CORR)) |>
  head(200)

highlighted_site <- pcg_report$settings$conf$sample_properties$site
all_sites <- unique(dt_similarity_depmap$EXT_PRIMARY_SITE)
other_sites <- all_sites[ !all_sites == highlighted_site]

DT::datatable(
  dt_similarity_depmap,
  escape = F,
  options = list(pageLength = 10),
  extensions=c("Buttons","Responsive")) |>
  DT::formatStyle(
    c('EXT_PRIMARY_SITE'),
    fontWeight = 'bold',
    color = "white",
    backgroundColor = DT::styleEqual(
      c(highlighted_site, other_sites), 
      c(pcg_report$settings$conf$report_color,
        rep(pcgrr::color_palette$none, length(other_sites)))
    )
  )


```

```{r similarity_table_treehouse}
#| output: asis
#| eval: !expr as.logical("treehouse" %in% names(pcg_report[['content']]$expression[['similarity_analysis']])) == TRUE

source_name <- 'Treehouse Childhood Cancer Data Initiative'
cat('\n\n#### ',source_name,'\n\n<br>')

dt_similarity_treehouse <- 
  pcg_report[['content']]$expression[['similarity_analysis']][['treehouse']] |>
  dplyr::mutate(CORR = round(CORR, digits = 5)) |>
  dplyr::select(-c("EXT_SAMPLE_ID2","PROTEIN_CODING_ONLY")) |>
  dplyr::select(
    EXT_SAMPLE_ID,
    EXT_SAMPLE_NAME,
    CORR,
    EXT_PRIMARY_DIAGNOSIS,
    EXT_PRIMARY_SITE,
    EXT_SAMPLE_TYPE,
    EXT_AGE_CATEGORY,
    EXT_GENDER,
    dplyr::everything()
  ) |>
  dplyr::arrange(desc(CORR)) |>
  head(200)

highlighted_site <- pcg_report$settings$conf$sample_properties$site
all_sites <- unique(dt_similarity_treehouse$EXT_PRIMARY_SITE)
other_sites <- all_sites[ !all_sites == highlighted_site]

DT::datatable(
  dt_similarity_treehouse,
  escape = F,
  options = list(pageLength = 10),
  extensions=c("Buttons","Responsive")) |>
  DT::formatStyle(
    c('EXT_PRIMARY_SITE'),
    fontWeight = 'bold',
    color = "white",
    backgroundColor = DT::styleEqual(
      c(highlighted_site, other_sites), 
      c(pcg_report$settings$conf$report_color,
        rep(pcgrr::color_palette$none, length(other_sites)))
    )
  )

```

:::

<br><br>



