### Tier 2 - Other cancer mutation hotspots, curated mutations or predicted driver mutations {.tabset}

- A total of __`r nrow(report_data$variants_tier2_display)`__ unique, somatic variant(s) in the tumor sample are curated as disease-causing, predicted as driver mutations, or occur in previously detected mutation hotspots.
<br><br>

#### Known mutation hotspots

<br>

```{r tier2A, echo=F, results = 'asis'}
#colnames(report_data$variants_tier2_display) <- paste0("<span style=\"color:red\">",colnames(report_data$variants_tier2_display),"</span>")
 cat('The table below lists variants detected as mutation hotspots according to <a href="http:/cancerhotspots.org/" target="_blank">cancerhotspots.org<a/>:',sep='\n')
if(nrow(report_data$variants_tier2_hotspots) > 0){
  cat('\n')
  myOptions <- list(paging = F, searching=F,caching=F)
  if(nrow(report_data$variants_tier2_hotspots) > 10){
    myOptions <- list(paging = T,pageLength=10, searching=T,caching=F)
  }
  report_data$variants_tier2_hotspots <- dplyr::rename(report_data$variants_tier2_hotspots, DOCM_DISEASE = OTHER_DISEASE_DOCM, DOCM_LITERATURE = OTHER_LITERATURE_DOCM)
  if(nrow(report_data$variants_tier2_hotspots[!is.na(report_data$variants_tier2_hotspots$INTOGEN_DRIVER_MUT),]) > 0){
    report_data$variants_tier2_hotspots[!is.na(report_data$variants_tier2_hotspots$INTOGEN_DRIVER_MUT),]$INTOGEN_DRIVER_MUT <- TRUE
  }
  DT::datatable(dplyr::select(report_data$variants_tier2_hotspots, CANCER_MUTATION_HOTSPOT,dplyr::everything()), options = myOptions, escape=F, extensions = "Responsive") %>%
  DT::formatStyle('CANCER_MUTATION_HOTSPOT', backgroundColor = 'black', color='white')
}else{
  cat('<i>No variants found.</i>',sep='\n')
  cat('\n')
}
```

<br><br>

#### Other disease-causing mutations

<br>

```{r tier2B, echo=F, results='asis'}

if(nrow(report_data$variants_tier2_curated_mutations) > 0){
  cat('\n')
  cat('The table below lists all other variants in Tier 2 curated as disease-causing in the <a href="http://docm.genome.wustl.edu/" target="_blank">Database of Curated Mutations<a/>:',sep='\n')
  myOptions <- list(paging = F, searching=F,caching=F)
  if(nrow(report_data$variants_tier2_curated_mutations) > 10){
    myOptions <- list(paging = T,pageLength=10, searching=T,caching=F)
  }

  report_data$variants_tier2_curated_mutations <- dplyr::rename(report_data$variants_tier2_curated_mutations, DOCM_DISEASE = OTHER_DISEASE_DOCM, DOCM_LITERATURE = OTHER_LITERATURE_DOCM)
  if(nrow(report_data$variants_tier2_curated_mutations[!is.na(report_data$variants_tier2_curated_mutations$INTOGEN_DRIVER_MUT),]) > 0){
    report_data$variants_tier2_curated_mutations[!is.na(report_data$variants_tier2_curated_mutations$INTOGEN_DRIVER_MUT),]$INTOGEN_DRIVER_MUT <- TRUE
  }
  report_data$variants_tier2_curated_mutations <- dplyr::select(report_data$variants_tier2_curated_mutations, -CANCER_MUTATION_HOTSPOT)
  DT::datatable(dplyr::select(report_data$variants_tier2_curated_mutations, DOCM_DISEASE, dplyr::everything()), options = myOptions, escape=F, extensions = "Responsive") %>%
  DT::formatStyle('DOCM_DISEASE', backgroundColor = 'black', color = 'white')
}else{
  cat('<i>No variants found.</i>',sep='\n')
  cat('\n')
}
```

<br><br>

#### Other predicted driver mutations

<br>

```{r tier2C, echo=F,results='asis'}

if(nrow(report_data$variants_tier2_predicted_drivers) > 0){
  cat('The table below lists all other variants in Tier 2 predicted as cancer driver mutations in <a href="https://www.intogen.org/downloads" target="_blank">Intogen\'s catalog of driver mutations</a>:',sep='\n')
  
  myOptions <- list(paging = F, searching=F,caching=F)
  if(nrow(report_data$variants_tier2_predicted_drivers) > 10){
    myOptions <- list(paging = T,pageLength=10, searching=T,caching=F)
  }
  #myOptions <- list(paging = T,pageLength=10, searching=T,caching=F)
  report_data$variants_tier2_predicted_drivers <- dplyr::rename(report_data$variants_tier2_predicted_drivers, DOCM_DISEASE = OTHER_DISEASE_DOCM, DOCM_LITERATURE = OTHER_LITERATURE_DOCM)
  report_data$variants_tier2_predicted_drivers <- dplyr::select(report_data$variants_tier2_predicted_drivers, -c(INTOGEN_DRIVER_MUT, CANCER_MUTATION_HOTSPOT, DOCM_DISEASE, DOCM_LITERATURE))
  DT::datatable(dplyr::select(report_data$variants_tier2_predicted_drivers, SYMBOL, CONSEQUENCE, dplyr::everything()),options = myOptions, escape=F, extensions = "Responsive") %>%
  DT::formatStyle(columns = c('SYMBOL','CONSEQUENCE'), backgroundColor = 'black',color = 'white')
}else{
  cat('<i>No variants found.</i>',sep='\n')
  cat('\n')
}
```

<br><br>
