## Somatic CNAs

```{r setup_cna}
#| echo: false
#| eval: true

vstats <- pcg_report$content$cna$vstats
tumor_site <- pcg_report$settings$conf$sample_properties$site
cs <- pcgrr::color_palette$bm_callout

```


<br>

::: {.panel-tabset}

#### Absolute copy number profile

```{r plot_cna}
#| eval: !expr NROW(pcg_report$content$cna$callset$variant) > 0
#| echo: false
#| output: asis
#| fig.width: 15
#| fig.height: 10

tumor_ploidy_plot <- pcg_report$settings$conf$sample_properties$tumor_ploidy
if (is.na(tumor_ploidy_plot) || tumor_ploidy_plot == "NA") {
  tumor_ploidy_plot <- 2
}
tumor_ploidy_plot <- as.numeric(tumor_ploidy_plot)

amp_threshold_effective_plot <-
  pcg_report$settings$conf$somatic_cna$amp_threshold_effective
if (is.null(amp_threshold_effective_plot) ||
    is.na(amp_threshold_effective_plot)) {
  amp_threshold_effective_plot <-
    pcg_report$settings$conf$somatic_cna$amp_threshold_absolute
}
amp_threshold_effective_plot <- as.numeric(amp_threshold_effective_plot)
threshold_mode_plot <- as.character(
  pcg_report$settings$conf$somatic_cna$threshold_mode)
gain_threshold_effective_plot <- if (threshold_mode_plot == "relative") {
  tumor_ploidy_plot * as.numeric(
    pcg_report$settings$conf$somatic_cna$gain_threshold_relative)
} else {
  as.numeric(pcg_report$settings$conf$somatic_cna$gain_threshold_absolute)
}
del_threshold_effective_plot <- if (threshold_mode_plot == "relative") {
  tumor_ploidy_plot * as.numeric(
    pcg_report$settings$conf$somatic_cna$del_threshold_relative)
} else {
  as.numeric(pcg_report$settings$conf$somatic_cna$del_threshold_absolute)
}

cna_plot <- pcgrr::plot_cna_segments_absolute(
  chrom_coordinates = pcg_report$settings$chrom_coordinates,
  cna_segment = pcg_report$content$cna$callset$segment,
  cna_gene = pcg_report$content$cna$callset$variant,
  tumor_ploidy = tumor_ploidy_plot,
  amp_threshold_effective = amp_threshold_effective_plot,
  threshold_mode = threshold_mode_plot,
  gain_threshold_effective = gain_threshold_effective_plot,
  del_threshold_effective = del_threshold_effective_plot
)

p_margin <-
  list(l = 50, r = 0, b = 0, t = 0, pad = 0)

plotly::ggplotly(cna_plot, tooltip = c("y","z")) |>
  plotly::layout(
    legend = list(
      orientation = "h", x = 0.5, xanchor = "center", y = -0.28,
      font = list(size = 12)),
    margin = p_margin)

```

#### Relative copy number profile

```{r plot_cna_relative}
#| eval: !expr NROW(pcg_report$content$cna$callset$variant) > 0
#| echo: false
#| output: asis
#| fig.width: 15
#| fig.height: 10

tumor_ploidy_plot <- pcg_report$settings$conf$sample_properties$tumor_ploidy
if (is.na(tumor_ploidy_plot) || tumor_ploidy_plot == "NA") {
  tumor_ploidy_plot <- 2
}
tumor_ploidy_plot <- as.numeric(tumor_ploidy_plot)

amp_threshold_effective_plot <-
  pcg_report$settings$conf$somatic_cna$amp_threshold_effective
if (is.null(amp_threshold_effective_plot) ||
    is.na(amp_threshold_effective_plot)) {
  amp_threshold_effective_plot <-
    pcg_report$settings$conf$somatic_cna$amp_threshold_absolute
}
amp_threshold_effective_plot <- as.numeric(amp_threshold_effective_plot)
threshold_mode_plot <- as.character(
  pcg_report$settings$conf$somatic_cna$threshold_mode)
gain_threshold_effective_plot <- if (threshold_mode_plot == "relative") {
  tumor_ploidy_plot * as.numeric(
    pcg_report$settings$conf$somatic_cna$gain_threshold_relative)
} else {
  as.numeric(pcg_report$settings$conf$somatic_cna$gain_threshold_absolute)
}
del_threshold_effective_plot <- if (threshold_mode_plot == "relative") {
  tumor_ploidy_plot * as.numeric(
    pcg_report$settings$conf$somatic_cna$del_threshold_relative)
} else {
  as.numeric(pcg_report$settings$conf$somatic_cna$del_threshold_absolute)
}

cna_plot_relative <- pcgrr::plot_cna_segments_relative(
  chrom_coordinates = pcg_report$settings$chrom_coordinates,
  cna_segment = pcg_report$content$cna$callset$segment,
  cna_gene = pcg_report$content$cna$callset$variant,
  tumor_ploidy = tumor_ploidy_plot,
  amp_threshold_effective = amp_threshold_effective_plot,
  threshold_mode = threshold_mode_plot,
  gain_threshold_effective = gain_threshold_effective_plot,
  del_threshold_effective = del_threshold_effective_plot
)

p_margin <-
  list(l = 50, r = 0, b = 0, t = 0, pad = 0)

plotly::ggplotly(cna_plot_relative, tooltip = c("y", "z")) |>
  plotly::layout(
    legend = list(
      orientation = "h", x = 0.5, xanchor = "center", y = -0.28,
      font = list(size = 12)),
    margin = p_margin)

```

:::



<br>

### Potentially oncogenic events

Here we show all protein-coding genes subject to copy number alteration that may potentially have oncogenic effects. Specifically, we list

 * [Proto-oncogenes](#documentation) subject to copy-number __amplification__ (total copy number >= __`r pcg_report$settings$conf$somatic_cna$amp_threshold_effective`__ (effective amplification threshold, user-configurable))
 * [Tumor suppressor](#documentation) genes subject to __deep deletions__ (homozygous/hemizygous deletions; total copy number = __0__), and __shallow deletions__ (heterozygous deletions; total copy number less than baseline ploidy (__`r pcg_report$settings$conf$sample_properties$tumor_ploidy`__))
 * Other clinically actionable genes subject to amplification

Gene symbols are color-coded according to their strength of association to cancer (here, specifically __`r pcg_report$settings$conf$sample_properties$site`__ cancers), from <span style="background-color:`r pcgrr::color_palette$cancer_assoc$values[1]`; padding:2x; color:white"><b>&nbsp; weak associations &nbsp; </b></span>&nbsp; to <span style="background-color:`r pcgrr::color_palette$cancer_assoc$values[length(pcgrr::color_palette$cancer_assoc$values)]`; padding:2x; color:white"><b>&nbsp; strong associations &nbsp; </b></span> [@Ochoa2021-jc;@Nakken2023-ol].


<br>

```{r cna_oncogenic_data}
#| echo: false

oncogenic_events <- pcgrr::get_oncogenic_cna_events(
  cna_df_display =
    pcg_report[["content"]][["cna"]][["callset"]][["variant_display"]])

```

```{r cna_oncogenic_display}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(NROW(oncogenic_events) > 0)

symbol_rank_col <- if (
  pcg_report$settings$conf$sample_properties$site == "Any"
) "GLOBAL_ASSOC_RANK" else "TISSUE_ASSOC_RANK"

primary_cols <- c(
  "SYMBOL", 
  "GENENAME", 
  "VARIANT_CLASS_DISPLAY", 
  "EVENT_TYPE",
  "CN_TOTAL", 
  "CYTOBAND", 
  "SEGMENT_LENGTH_MB"
)

vc_levels <- 
  pcgrr::color_palette$cna_variant_class$levels_display
vc_colors <- 
  pcgrr::color_palette$cna_variant_class$values

col_defs <- list(
  SYMBOL = reactable::colDef(
    name = "Gene",
    style = pcgrr::render_symbol_assoc_style(
      oncogenic_events, 
      rank_col = symbol_rank_col),
    minWidth = 90
  ),
  GENENAME = reactable::colDef(
    name = "Genename", minWidth = 150),
  VARIANT_CLASS_DISPLAY = reactable::colDef(
    name = "Variant Class",
    align = "center",
    style = function(value) {
      bg <- vc_colors[match(value, vc_levels)]
      if (is.na(bg)) bg <- "transparent"
      list(background = bg, 
           color = "white", 
           fontWeight = "bold",
           padding = "4px 8px")
    },
    minWidth = 110
  ),
  EVENT_TYPE = reactable::colDef(
    name = "Event Type", align = "center", maxWidth = 100),
  CN_TOTAL = reactable::colDef(
    name = "Total CN", align = "center", maxWidth = 85),
  CYTOBAND = reactable::colDef(
    name = "Cytoband", maxWidth = 100),
  SEGMENT_LENGTH_MB = reactable::colDef(
    name = "Segment (Mb)", maxWidth = 110),
  GLOBAL_ASSOC_RANK = reactable::colDef(show = FALSE),
  TISSUE_ASSOC_RANK = reactable::colDef(show = FALSE)
)

## Hide all remaining non-primary columns from main view (available in row details)
for (col in colnames(oncogenic_events)) {
  if (!col %in% names(col_defs)) {
    col_defs[[col]] <- reactable::colDef(show = FALSE)
  }
}

oncogenic_events <- oncogenic_events |>
  dplyr::mutate(
    dplyr::across(where(is.numeric), \(x) round(x, 3)))

reactable::reactable(
  oncogenic_events,
  columns = col_defs,
  defaultColDef = reactable::colDef(html = TRUE),
  details = pcgrr::build_rt_row_details(
    primary_cols = primary_cols),
  searchable = TRUE,
  filterable = TRUE,
  striped = TRUE,
  highlight = TRUE,
  compact = TRUE,
  wrap = TRUE,
  defaultPageSize = 10,
  theme = pcgrr::rt_theme())

```

```{r cna_other_oncogenic_empty}
#| echo: false
#| output: asis
#| eval: !expr NROW(oncogenic_events) == 0

htmltools::br()
cat('\n*  <i> <font style="font-size: 110%"><b>NO</b> potentially oncogenic copy-number events were identified.</font></i>', sep = '\n')
cat('\n')
htmltools::br()

```

<br>

### Potential two-hit events

For tumor suppressor genes (TSGs) subject to loss-of-heterozygosity (LOH) – either a deletion or copy-neutral LOH event – we show any co-occurring somatic or germline loss-of-function SNV/InDels that may constitute a _two-hit_ inactivation of the gene. 

<br>

::: {.callout-note}

## Note - loss-of-function annotation

If non-truncating, missense variants are listed among the loss-of-function variants, these are likely annotated by other means than PCGR's main loss-of-function
annotation strategy. Specifically, evidence from curated sources (e.g. OncoKB) may indicate loss-of-function also for missense variation. In such cases, the underlying evidence should be carefully consulted for further interpretation (cross-check with the detailed table view of SNVs/InDels elsewhere in the report).

:::

::: {.callout-note collapse="true"}

## Note - VAF consistency assessment

For somatic candidates, the **VAF consistency** column indicates whether the observed tumor VAF is consistent with the variant residing on the _retained_ (major) allele of the LOH segment. The expected VAF of a clonal mutation is derived from the purity-adjusted mixture model ([Carter et al., Nat Biotechnol 2012](https://doi.org/10.1038/nbt.2203)):

$$\text{VAF}_{\text{expected}} = \frac{p \times n_{\text{mut}}}{p \times n_{\text{total}} + 2 \times (1 - p)}$$

where $p$ is tumor purity, $n_{\text{mut}}$ is the number of mutant allele copies on the retained allele (1 for deletion LOH; lower-bound of 1 for copy-neutral LOH), and $n_{\text{total}}$ is the total copy number of the segment. A variant is considered consistent if the observed tumor VAF falls within a tolerance of ±0.15 of the expected VAF.

- **VAF_CONSISTENT**: observed VAF ≥ expected − tolerance → variant likely on the retained allele
- **VAF_LOW**: observed VAF below expectation → variant may reside on the _deleted_ allele, or be subclonal relative to the LOH event; interpret with caution
- **VAF_UNKNOWN**: tumor purity not provided, or VAF not available for the variant; VAF consistency cannot be assessed


:::


```{r twohit_purity_warn}
#| echo: false
#| output: asis

twohit_purity_missing <- is.null(pcg_report$settings$conf$sample_properties$tumor_purity) ||
  is.na(pcg_report$settings$conf$sample_properties$tumor_purity) ||
  pcg_report$settings$conf$sample_properties$tumor_purity == "NA"

if (isTRUE(twohit_purity_missing)) {
  cat("::: {.callout-warning}\n\n")
  cat("## Tumor purity not provided\n\n")
  cat("VAF consistency assessment for somatic two-hit candidates requires tumor purity. ")
  cat("The **VAF consistency** column will show `VAF_UNKNOWN` for all somatic candidates. ")
  cat("Consider re-running PCGR with `--tumor_purity` to enable purity-adjusted VAF interpretation.\n\n")
  cat(":::\n\n")
}
```

<br>

```{r twohit_data}
#| echo: false
#| eval: true

snv_somatic_twohit <- pcg_report$content$snv_indel$callset$variant
snv_germline_twohit <- NULL
if (!is.null(pcg_report$content$germline_classified)) {
  snv_germline_twohit <- pcg_report$content$germline_classified$callset$variant
}

twohit <- pcgrr::build_twohit_display_data(
  cna_variant  = pcg_report$content$cna$callset$variant,
  snv_somatic  = snv_somatic_twohit,
  snv_germline = snv_germline_twohit,
  settings     = pcg_report$settings
)

twohit_present <- NROW(twohit$main) > 0

```

```{r twohit_table}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(twohit_present)

vc_levels <- pcgrr::color_palette$cna_variant_class$levels_display
vc_colors <- pcgrr::color_palette$cna_variant_class$values

reactable::reactable(
  twohit$main,
  columns = list(
    .row_id      = reactable::colDef(show = FALSE),
    NESTED_HTML  = reactable::colDef(show = FALSE),
    SYMBOL       = reactable::colDef(name = "Gene", minWidth = 90),
    GENENAME     = reactable::colDef(name = "Gene name", minWidth = 150),
    VARIANT_CLASS_DISPLAY = reactable::colDef(
      name = "Variant Class",
      align = "center",
      style = function(value) {
        bg <- vc_colors[match(value, vc_levels)]
        if (is.na(bg)) bg <- "transparent"
        list(background = bg, color = "white",
             fontWeight = "bold", padding = "4px 8px")
      },
      minWidth = 110),
    CN_TOTAL = reactable::colDef(
      name = "Total CN", align = "center", maxWidth = 85),
    LOH = reactable::colDef(
      name = "LOH type", align = "center", maxWidth = 120),
    VAF_FLAG_SUMMARY = reactable::colDef(
      name = "VAF consistency",
      align = "center",
      maxWidth = 130,
      cell = function(value) {
        if (is.na(value)) return(htmltools::span("–", style = "color:#bdbdbd;"))
        col <- switch(value,
          VAF_CONSISTENT = "#2e7d32",
          VAF_LOW        = "#b71c1c",
          VAF_UNKNOWN    = "#757575",
          "#757575")
        htmltools::span(
          value,
          style = paste0(
            "background:", col, ";color:#fff;font-size:0.82em;",
            "font-weight:600;padding:2px 7px;border-radius:10px;",
            "white-space:nowrap;"))
      })
  ),
  defaultColDef = reactable::colDef(html = TRUE),
  details = reactable::JS(
    "function(rowInfo) {",
    "  var html = rowInfo.row['NESTED_HTML'];",
    "  if (!html) return null;",
    "  return html;",
    "}"
  ),
  searchable = FALSE,
  striped = TRUE,
  highlight = TRUE,
  compact = TRUE,
  defaultPageSize = 5,
  theme = pcgrr::rt_theme()
)

```

```{r twohit_empty}
#| echo: false
#| output: asis
#| eval: !expr isFALSE(twohit_present)

htmltools::div(
  style = paste0(
    "border-left:4px solid ", cs$therapeutic_resistance$border, ";",
    "background:", cs$therapeutic_resistance$bg, ";",
    "color:", cs$therapeutic_resistance$text, ";",
    "padding:10px 16px;border-radius:4px;margin:6px 0;",
    "font-size:0.95em;"),
  htmltools::strong("No potential two-hit events identified."),
  " No tumor suppressor genes with LOH were found to carry co-occurring",
  " somatic or germline loss-of-function variants."
)

```

<br>

### Clinical actionability classification

Genes subject to copy number aberrations in the tumor sample have been assessed for
_clinical actionability_ through an implementation of AMP/ASCO/CAP guidelines [@Li2017-ew]. We use `r bm_db_label` `r bm_db_citation` as the underlying sources for actionable variants (versions indicated below).

The table below indicates which variants have been classified into _tiers of clinical significance_ according to AMP/ASCO/CAP criteria.


```{r}
#| echo: false
#| output: asis
pcgrr::callout_biomarker_scope("copy number aberrations")
```


<br>


::: {.panel-tabset}


#### Therapeutic biomarkers

```{r tier1_2_3_sensitivity_variants_cna}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_sensitivity$tier1$n_gene > 0) || isTRUE(vstats$therapeutic_sensitivity$tier2$n_gene > 0) || isTRUE(vstats$therapeutic_sensitivity$tier3$n_gene > 0)

rctbl_recs <- pcgrr::prep_actble_display_tbl(
  rep = pcg_report, 
  tier = c(1,2,3),
  etype_for_tiering = c("predictive"),
  variant_category = 'cna')

result_table <- 
  pcgrr::render_actble_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'cna',
    clnsig = "therapeutic_sensitivity"
  )

result_table

```

<br>

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_sensitivity$tier1$n_gene == 0) && isTRUE(vstats$therapeutic_sensitivity$tier2$n_gene == 0) && isTRUE(vstats$therapeutic_sensitivity$tier3$n_gene == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$therapeutic_sensitivity$border, "; background:", cs$therapeutic_sensitivity$bg, "; border-radius:0 4px 4px 0; color:", cs$therapeutic_sensitivity$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no therapeutic sensitivity biomarkers (Tier 1&ndash;3) were identified among CNAs in this tumor sample.</div>"))
```

<br>

```{r tier1_2_3_resistance_variants_cna}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_resistance$tier1$n_gene > 0) || isTRUE(vstats$therapeutic_resistance$tier2$n_gene > 0) || isTRUE(vstats$therapeutic_resistance$tier3$n_gene > 0)

rctbl_recs <- pcgrr::prep_actble_display_tbl(
  rep = pcg_report, 
  tier = c(1,2,3),
  clnsig = "therapeutic_resistance",
  variant_category = 'cna')

result_table <-
  pcgrr::render_actble_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'cna',
    clnsig = "therapeutic_resistance"
  )

if (is.null(result_table)) {
  cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$therapeutic_resistance$border, "; background:", cs$therapeutic_resistance$bg, "; border-radius:0 4px 4px 0; color:", cs$therapeutic_resistance$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no therapeutic resistance biomarkers for ", tumor_site, " cancers were identified among CNAs in this tumor sample.</div>"))
} else {
  result_table
}

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_resistance$tier1$n_gene == 0) && isTRUE(vstats$therapeutic_resistance$tier2$n_gene == 0) && isTRUE(vstats$therapeutic_resistance$tier3$n_gene == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$therapeutic_resistance$border, "; background:", cs$therapeutic_resistance$bg, "; border-radius:0 4px 4px 0; color:", cs$therapeutic_resistance$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no therapeutic resistance biomarkers for ", tumor_site, " cancers were identified among CNAs in this tumor sample.</div>"))
```

#### Prognostic biomarkers

```{r tier1_2_poor_prognosis_variants_cna}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_poor$tier1$n_gene > 0) || isTRUE(vstats$prognostic_poor$tier2$n_gene > 0)

rctbl_recs <- pcgrr::prep_progn_display_tbl(
  rep = pcg_report,
  etype_for_tiering = c("prognostic"),
  variant_category = 'cna',
  clnsig = "prognostic_poor")

result_table <-
  pcgrr::render_progn_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'cna',
    clnsig = "prognostic_poor"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_poor$tier1$n_gene == 0) && isTRUE(vstats$prognostic_poor$tier2$n_gene == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$prognostic_poor$border, "; background:", cs$prognostic_poor$bg, "; border-radius:0 4px 4px 0; color:", cs$prognostic_poor$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no poor outcome biomarkers for ", tumor_site, " cancers were identified among CNAs in this tumor sample.</div>"))
```

<br>

```{r tier1_2_better_prognosis_variants_cna}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_better$tier1$n_gene > 0) || isTRUE(vstats$prognostic_better$tier2$n_gene > 0)

rctbl_recs <- pcgrr::prep_progn_display_tbl(
  rep = pcg_report,
  etype_for_tiering = c("prognostic"),
  variant_category = 'cna',
  clnsig = "prognostic_better")

result_table <-
  pcgrr::render_progn_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'cna',
    clnsig = "prognostic_better"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_better$tier1$n_gene == 0) && isTRUE(vstats$prognostic_better$tier2$n_gene == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$prognostic_better$border, "; background:", cs$prognostic_better$bg, "; border-radius:0 4px 4px 0; color:", cs$prognostic_better$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no better outcome biomarkers for ", tumor_site, " cancers were identified among CNAs in this tumor sample.</div>"))
```

#### Diagnostic biomarkers

```{r tier1_2_diagnostic_variants_cna}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$diagnostic_positive$tier1$n_gene > 0) || isTRUE(vstats$diagnostic_positive$tier2$n_gene > 0)

rctbl_recs <- pcgrr::prep_diagn_display_tbl(
  rep = pcg_report,
  etype_for_tiering = c("diagnostic"),
  variant_category = 'cna',
  clnsig = "diagnostic_positive")

result_table <-
  pcgrr::render_diagn_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'cna',
    clnsig = "diagnostic_positive"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$diagnostic_positive$tier1$n_gene == 0) && isTRUE(vstats$diagnostic_positive$tier2$n_gene == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$diagnostic_positive$border, "; background:", cs$diagnostic_positive$bg, "; border-radius:0 4px 4px 0; color:", cs$diagnostic_positive$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no diagnostic biomarkers for ", tumor_site, " cancers were identified among CNAs in this tumor sample.</div>"))
```




:::


<br>
<br>
