## Germline findings

```{r prepare_panel_url}
#| echo: false
#| results: asis

panel_link <- pcg_report[['content']][['germline_classified']][['panel_info']][['description']]
if (pcg_report[['content']][['germline_classified']][['panel_info']][['panel_id']] != "-1" &
   !stringr::str_detect(pcg_report[['content']][['germline_classified']][['panel_info']][['url']], ",")) {
  description <- pcg_report[['content']][['germline_classified']][['panel_info']][['description']]
  description_trait <- 
    pcg_report[['content']][['germline_classified']][['panel_info']][['description_trait']]
  url_raw <- pcg_report[['content']][['germline_classified']][['panel_info']][['url']]
  description_full <- paste0(description,': ', description_trait)
  if (pcg_report[['content']][['germline_classified']][['panel_info']][['panel_id']] == "0") {
    description_full <- description
  }
  panel_link <- paste0("<a href='", url_raw, "' target='_blank'>", 
                       description_full, 
                       "</a>")
}

```


* Based on a germline variant analysis of the query case using the [Cancer Predisposition Sequencing Reporter (CPSR)](https://github.com/sigven/cpsr), we here list the variants of clinical significance in cancer predisposition genes, both novel (not recorded in ClinVar), and those with existing classifications in ClinVar.
    * Virtual panel of cancer predisposition genes screened: `r panel_link`
    * All variants classified as Pathogenic or Likely Pathogenic are listed here, regardless of coding status
    * Variants of uncertain significance (VUS) are `r ifelse(!pcg_report$settings$conf$germline$ignore_vus, "__also shown__, but limited to a focused subset of protein-coding VUS (missense, nonsense, frameshift, essential splice site etc.) - __not__ the complete set of VUS identified by CPSR", "__not included__ in this list")`


```{r cpsr_clinvar_findings}
#| echo: false
#| eval: true
#| output: asis

cat('\n')
htmltools::br()

germline_calls <- pcg_report[['content']][['germline_classified']][['callset']]

if (NROW(germline_calls$variant_display) > 100) {
  cat('<b>NOTE - only considering top 100 variants (due to limitations with client-side tables) </b><br>',sep="\n")
  cat('<br>')
  germline_calls$variant_display <-
    head(germline_calls$variant_display, 100)
}

## DP_CONTROL is set to -1 when CPSR is run without a matched control sample
## (i.e. depth is unknown for all variants in that case) - only flag variants
## as low-depth when a real (non-sentinel) depth value is present and low
low_depth_threshold <- 10
has_control_depth <- "DP_CONTROL" %in% colnames(germline_calls$variant_display) &&
  any(germline_calls$variant_display$DP_CONTROL >= 0, na.rm = TRUE)

if (has_control_depth) {
  cat(paste0(
    '<b>NOTE</b> - variants with a low sequencing depth (DP < ',
    low_depth_threshold,
    'X) are flagged with an orange marker on the genotype - interpret their classification with caution<br>'), sep = "\n")
  cat('<br>')
}

germline_dt <- DT::datatable(
  germline_calls$variant_display,
  escape = FALSE,
  extensions = c("Buttons","Responsive"),
  options = list(
    pageLength = 10,
    scrollCollapse = TRUE,
    buttons = c('csv','excel'),
    dom = 'Bfrtip'
  )) |>
  DT::formatStyle(
    columns = c("SYMBOL","ALTERATION"),
    valueColumns = c("CLINICAL_SIGNIFICANCE"),
    color = "white",
    backgroundColor =
      DT::styleEqual(
        c("Pathogenic", "Likely Pathogenic","VUS"),
        c("#9E0142","#D53E4F","#000000")
      )
  )

if (has_control_depth) {
  germline_dt <- germline_dt |>
    DT::formatStyle(
      columns = "GENOTYPE",
      valueColumns = "DP_CONTROL",
      `border-left` = DT::styleInterval(
        c(0, low_depth_threshold), c("none", "4px solid #E69F00", "none")),
      fontWeight = DT::styleInterval(
        c(0, low_depth_threshold), c("normal", "bold", "normal"))
    )
}

bslib::page_fillable(
  bslib::card(
    bslib::card_header(
      class = "bg-dark",
      paste0("Germline variants - ",
               pcg_report[['content']][['germline_classified']]$sample_id)
    ),
    bslib::card_body(
      height = min(500, 150 + NROW(germline_calls$variant_display) * 80),
      if (NROW(germline_calls$variant_display) > 0) {
        germline_dt
      }else{
        "NO cancer-predisposing variants of clinical significance were found in the query case (CPSR report)."
      }
    )
  )
)


```

<br>

