## Germline findings

```{r prepare_panel_url}
#| echo: false
#| results: asis

panel_link <- pcg_report[['content']][['germline_classified']][['panel_info']][['description']]
if (pcg_report[['content']][['germline_classified']][['panel_info']][['panel_id']] != "-1" &
   !stringr::str_detect(pcg_report[['content']][['germline_classified']][['panel_info']][['url']], ",")) {
  description <- pcg_report[['content']][['germline_classified']][['panel_info']][['description']]
  description_trait <- 
    pcg_report[['content']][['germline_classified']][['panel_info']][['description_trait']]
  url_raw <- pcg_report[['content']][['germline_classified']][['panel_info']][['url']]
  description_full <- paste0(description,': ', description_trait)
  if (pcg_report[['content']][['germline_classified']][['panel_info']][['panel_id']] == "0") {
    description_full <- description
  }
  panel_link <- paste0("<a href='", url_raw, "' target='_blank'>", 
                       description_full, 
                       "</a>")
}

```


* Based on a germline variant analysis of the query case using the [Cancer Predisposition Sequencing Reporter (CPSR)](https://github.com/sigven/cpsr), we here list the variants of clinical significance in cancer predisposition genes, both novel (not recorded in ClinVar), and those with existing classifications in ClinVar.
    * Virtual panel of cancer predisposition genes screened: `r panel_link`
    * All variants classified as Pathogenic or Likely Pathogenic are listed here, regardless of coding status
    * Variants of uncertain significance (VUS) are `r ifelse(!pcg_report$settings$conf$germline$ignore_vus, "__also shown__, but limited to a focused subset of protein-coding VUS (missense, nonsense, frameshift, essential splice site etc.) - __not__ the complete set of VUS identified by CPSR", "__not included__ in this list")`


```{r cpsr_clinvar_findings}
#| echo: false
#| eval: true
#| output: asis

cat('\n')
htmltools::br()

germline_calls <- pcg_report[['content']][['germline_classified']][['callset']]

if (NROW(germline_calls$variant_display) > 100) {
  cat('<b>NOTE - only considering top 100 variants (due to limitations with client-side tables) </b><br>',sep="\n")
  cat('<br>')
  germline_calls$variant_display <-
    head(germline_calls$variant_display, 100)
}

## DP_CONTROL is set to -1 when CPSR is run without a matched control sample
## (i.e. depth is unknown for all variants in that case) - only flag variants
## as low-depth when a real (non-sentinel) depth value is present and low
low_depth_threshold <- 10
has_control_depth <- "DP_CONTROL" %in% colnames(germline_calls$variant_display) &&
  any(germline_calls$variant_display$DP_CONTROL >= 0, na.rm = TRUE)

if (has_control_depth) {
  cat(paste0(
    '<b>NOTE</b> - variants with a low sequencing depth (DP < ',
    low_depth_threshold,
    'X) are flagged with an orange marker on the genotype - interpret their classification with caution<br>'), sep = "\n")
  cat('<br>')
} else if ("DP_CONTROL" %in% colnames(germline_calls$variant_display)) {
  ## DP_CONTROL carries no information without a matched control sample -
  ## drop it so a sentinel value (-1) does not surface in the row details
  germline_calls$variant_display$DP_CONTROL <- NULL
}

## Some cancer predisposition genes (e.g. the mismatch repair genes,
## BRCA1) are linked to more than one inheritance mode, as monoallelic vs.
## biallelic pathogenic variants cause distinct conditions (e.g. Lynch
## syndrome vs. CMMRD for PMS2/MSH6) - flag this explicitly, and only when
## relevant for the variants at hand
has_dual_moi <- "CPG_MOI" %in% colnames(germline_calls$variant_display) &&
  any(stringr::str_detect(
    germline_calls$variant_display$CPG_MOI, "/"), na.rm = TRUE)

if (has_dual_moi) {
  cat(paste0(
    '<b>NOTE</b> - for a subset of genes, <i>Inheritance</i> combines two ',
    'modes (e.g. "AD/AR"): a heterozygous finding (see <i>Genotype</i>) ',
    'typically reflects the dominant condition, while a homozygous/',
    'biallelic finding may reflect a distinct, recessive condition<br>'),
    sep = "\n")
  cat('<br>')
}

germline_primary_cols <- c(
  "SYMBOL", "ALTERATION", "CONSEQUENCE", "GENOTYPE",
  "CLINICAL_SIGNIFICANCE", "ACMG_CODE", "CPG_MOI"
)

## reactable renders visible columns in the underlying data frame's
## column order (not the order of the colDef list) - relocate the
## primary columns to the front, in the intended display order
germline_calls$variant_display <- germline_calls$variant_display |>
  dplyr::relocate(dplyr::any_of(germline_primary_cols))

if (NROW(germline_calls$variant_display) > 0) {
  reactable::reactable(
    germline_calls$variant_display,
    columns = pcgrr::build_germline_col_defs(
      data = germline_calls$variant_display,
      primary_cols = germline_primary_cols,
      has_control_depth = has_control_depth),
    defaultColDef = reactable::colDef(html = TRUE),
    details = pcgrr::build_rt_row_details(
      germline_primary_cols,
      styling_cols = character(0)),
    searchable = TRUE,
    striped = TRUE,
    highlight = TRUE,
    compact = TRUE,
    wrap = TRUE,
    defaultPageSize = 10,
    theme = pcgrr::rt_theme())
} else {
  cat("NO cancer-predisposing variants of clinical significance were found in the query case (CPSR report).")
}


```

<br>

