## Germline findings

```{r prepare_panel_url}
#| echo: false
#| results: asis

panel_link <- pcg_report[['content']][['germline_classified']][['panel_info']][['description']]
if(pcg_report[['content']][['germline_classified']][['panel_info']][['panel_id']] != "-1" &
   !stringr::str_detect(pcg_report[['content']][['germline_classified']][['panel_info']][['url']], ",")){
  description <- pcg_report[['content']][['germline_classified']][['panel_info']][['description']]
  description_trait <- 
    pcg_report[['content']][['germline_classified']][['panel_info']][['description_trait']]
  url_raw <- pcg_report[['content']][['germline_classified']][['panel_info']][['url']]
  description_full <- paste0(description,': ', description_trait)
  if(pcg_report[['content']][['germline_classified']][['panel_info']][['panel_id']] == "0"){
    description_full <- description
  }
  panel_link <- paste0("<a href='", url_raw, "' target='_blank'>", 
                       description_full, 
                       "</a>")
}

```


* Based on a germline variant analysis of the query case using the [Cancer Predisposition Sequencing Reporter (CPSR)](https://github.com/sigven/cpsr), we here list the variants of clinical significance in cancer predisposition genes, both novel (not recorded in ClinVar), and those with existing classifications in ClinVar.
    * Virtual panel of cancer predisposition genes screened: `r panel_link`
    * Variants of uncertain significance shown: __`r !pcg_report$settings$conf$germline$ignore_vus`__


```{r cpsr_clinvar_findings}
#| echo: false
#| eval: true
#| output: asis

cat('\n')
htmltools::br()

germline_calls <- pcg_report[['content']][['germline_classified']][['callset']]

if(NROW(germline_calls$variant_display) > 100){
  cat('<b>NOTE - only considering top 100 variants (due to limitations with client-side tables) </b><br>',sep="\n")
  cat('<br>')
  germline_calls$variant_display <- 
    head(germline_calls$variant_display, 100)
}

germline_dt <- DT::datatable(
  germline_calls$variant_display, 
  escape = F,
  extensions = c("Buttons","Responsive"),
  options = list(
    pageLength = 10,
    scrollCollapse = T,
    buttons = c('csv','excel'),
    dom = 'Bfrtip'
  )) |>
  DT::formatStyle(
    columns = c("SYMBOL","ALTERATION"),
    valueColumns = c("CLINICAL_SIGNIFICANCE"), 
    color = "white", 
    backgroundColor = 
      DT::styleEqual(
        c("Pathogenic", "Likely_Pathogenic","VUS"), 
        c("#9E0142","#D53E4F","#000000")
      )
  )

bslib::page_fillable(
  bslib::card(
    bslib::card_header(
      class = "bg-dark",
      paste0("Germline variants - ",
               pcg_report[['content']][['germline_classified']]$sample_id)
    ),
    bslib::card_body(
      height = min(500, 150 + NROW(germline_calls$variant_display) * 80),
      if(NROW(germline_calls$variant_display) > 0){
        germline_dt
      }else{
        "NO cancer-predisposing variants of clinical significance were found in the query case (CPSR report)."
      }
    )
  )
)


```

<br>

