## Somatic RNA fusions

```{r setup_fusion}
#| output: asis
#| echo: false
#| eval: true

vstats <- pcg_report$content$fusion$vstats
callset <- pcg_report$content$fusion$callset
tumor_site <- pcg_report$settings$conf$sample_properties$site
cs <- pcgrr::color_palette$bm_callout

```


The table below lists somatic RNA fusion events detected in the input sample. 
Fusion events are annotated with respect to known cancer gene partners and 
presence in public fusion databases (i.e. Mitelman Database).

::: {.callout-important title="Quality control of fusion events" collapse="true"}

The user need to perform quality control of fusion events upstream of PCGR analysis - the 
focus here is strictly on functional annotation and potential clinical impact. 
We recommend to carefully check e.g. the number of supporting reads for each 
fusion event, and to visually inspect fusion events in a genome browser 
(e.g. IGV) or other third-party tools (e.g. [FusViz](https://github.com/senzhaocode/FuSViz)) 
to conduct frame inference, investigate domain retention, and rule out potential false positives.
:::

::: {.callout-note title="Recurrence of fusion events in Mitelman Database" collapse="true"}

Fusion events reported here are annotated with respect to recurrence in the 
[Mitelman Database of chromosome aberrations and gene fusions in cancer](https://mitelmandatabase.isb-cgc.org/). 
This annotation is notably done at the level of fusion partners, and _not_ 
at the level of exact genomic breakpoints. The absence of a fusion 
event in this database does not rule out biological relevance, 
particularly if it involves known oncogenic driver genes.
:::

<br>

```{r all_fusion_events}
#| eval: !expr isTRUE(NROW(callset$variant_display) > 0)
#| output: asis
#| echo: false

fusion_events <- callset$variant_display |>
  dplyr::mutate(
    PRIMARY_SITE_MATCH = dplyr::if_else(
      stringr::str_detect(
        .data$MITDB_EVIDENCE, 
        glue::glue("<b>{pcg_report$settings$conf$sample_properties$site}</b>")),
      TRUE,FALSE,FALSE
    )
  ) |>
  dplyr::arrange(
    dplyr::desc(PRIMARY_SITE_MATCH),
    dplyr::desc(ONCOGENE_5P),
    dplyr::desc(ONCOGENE_3P),
    dplyr::desc(MITDB_NUM_EVIDENCE)) |>
  dplyr::select(-dplyr::any_of(
     c("VAR_ID","ENTREZGENE","VARIANT_CLASS",
       "SAMPLE_ALTERATION",
       "GLOBAL_ASSOC_RANK","ACTIONABILITY_TIER",
       "ACTIONABILITY",
       "PRIMARY_SITE_MATCH")))

fusion_events_shared <- crosstalk::SharedData$new(
  fusion_events)

crosstalk::bscols(
  list(
    crosstalk::filter_checkbox(
      id = "ONCOGENE_5P", inline = TRUE, label = "Oncogenic fusion partner - 5'", 
      sharedData = fusion_events_shared, group = ~ONCOGENE_5P),
     crosstalk::filter_checkbox(
      id = "ONCOGENE_3P", inline = TRUE, label = "Oncogenic fusion partner - 3'", 
      sharedData = fusion_events_shared, group = ~ONCOGENE_3P)
  ),
  list(
    crosstalk::filter_slider(
      "MITDB_NUM_EVIDENCE","Recurrence in MitelmanDB", 
      fusion_events_shared, ~MITDB_NUM_EVIDENCE)
  )
)

# htmltools::br()
# 
# DT::datatable(
#   fusion_events_shared,
#   escape = F, 
#   extensions = c("Buttons","Responsive"), 
#   options = list(
#     buttons = c('csv','excel'), 
#     dom = 'Bfrtip')) |>
#   DT::formatStyle(
#     color = "white", 
#     c("FUSION_GENE"), 
#     fontWeight = 'bold', 
#     `text-align` = 'center',
#     backgroundColor = pcgrr::color_palette$bg_dark)
# 
# 

primary_cols <- c(
  "FUSION_GENE2", "SPLIT_READS","FUSION_GENE_5P",            
 "FUSION_GENE_3P","BREAKPOINT_5P","BREAKPOINT_3P")

#vc_levels <- pcgrr::color_palette$cna_variant_class$levels
#vc_colors <- pcgrr::color_palette$cna_variant_class$values

col_defs <- list(
  FUSION_GENE2 = reactable::colDef(
    name = "Fusion gene",
    #style = pcgrr::render_symbol_assoc_style(
    #  dt_other_oncogenic, rank_col = symbol_rank_col),
    minWidth = 130
  ),
  SPLIT_READS = reactable::colDef(name = "Split Reads", minWidth = 90),
  FUSION_GENE_3P = reactable::colDef(name = "3' partner", minWidth = 120),
  FUSION_GENE_5P = reactable::colDef(name = "5' partner", minWidth = 120),
  BREAKPOINT_5P = reactable::colDef(name = "5' breakpoint", minWidth = 120),
  BREAKPOINT_3P = reactable::colDef(name = "3' breakpoint", minWidth = 120)
)

## Hide all remaining non-primary columns from main view (available in row details)
for (col in colnames(fusion_events)) {
  if (!col %in% names(col_defs)) {
    col_defs[[col]] <- reactable::colDef(show = FALSE)
  }
}

#dt_other_oncogenic <- dt_other_oncogenic |>
#  dplyr::mutate(dplyr::across(where(is.numeric), \(x) round(x, 3)))

reactable::reactable(
  fusion_events_shared,
  columns = col_defs,
  defaultColDef = reactable::colDef(html = TRUE),
  details = pcgrr::build_rt_row_details(
    primary_cols = primary_cols),
  searchable = TRUE,
  filterable = TRUE,
  striped = TRUE,
  highlight = TRUE,
  compact = TRUE,
  wrap = TRUE,
  defaultPageSize = 10,
  theme = pcgrr::rt_theme())


```

<br>

### Clinical actionability classification

RNA fusions reported in the tumor sample have been assessed for
_clinical actionability_ through an implementation of AMP/ASCO/CAP guidelines [@Li2017-ew]. We use `r bm_db_label` `r bm_db_citation` as the underlying sources for actionable variants (versions indicated below).

The table below indicates which variants have been classified into _tiers of clinical significance_ according to AMP/ASCO/CAP criteria.


::: {.panel-tabset}


#### Therapeutic biomarkers

```{r tier1_2_3_sensitivity_variants_fusion}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_sensitivity$tier1$n_fusion > 0) || isTRUE(vstats$therapeutic_sensitivity$tier2$n_fusion > 0) || isTRUE(vstats$therapeutic_sensitivity$tier3$n_fusion > 0)

rctbl_recs <- pcgrr::prep_actble_display_tbl(
  rep = pcg_report, 
  tier = c(1,2,3),
  etype_for_tiering = c("predictive"),
  variant_category = 'fusion')

result_table <- 
  pcgrr::render_actble_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'fusion',
    clnsig = "therapeutic_sensitivity"
  )

result_table

```


```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_sensitivity$tier1$n_fusion == 0) && isTRUE(vstats$therapeutic_sensitivity$tier2$n_fusion == 0) && isTRUE(vstats$therapeutic_sensitivity$tier3$n_fusion == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$therapeutic_sensitivity$border, "; background:", cs$therapeutic_sensitivity$bg, "; border-radius:0 4px 4px 0; color:", cs$therapeutic_sensitivity$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no therapeutic sensitivity biomarkers (Tier 1&ndash;3) were identified among RNA fusions in this tumor sample.</div>"))
```

<br>

```{r tier1_2_3_resistance_variants_fusion}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_resistance$tier1$n_fusion > 0) || isTRUE(vstats$therapeutic_resistance$tier2$n_fusion > 0) || isTRUE(vstats$therapeutic_resistance$tier3$n_fusion > 0)

rctbl_recs <- pcgrr::prep_actble_display_tbl(
  rep = pcg_report, 
  tier = c(1,2,3),
  clnsig = "therapeutic_resistance",
  variant_category = 'fusion')

result_table <- 
  pcgrr::render_actble_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'fusion',
    clnsig = "therapeutic_resistance"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_resistance$tier1$n_fusion == 0) && isTRUE(vstats$therapeutic_resistance$tier2$n_fusion == 0) && isTRUE(vstats$therapeutic_resistance$tier3$n_fusion == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$therapeutic_resistance$border, "; background:", cs$therapeutic_resistance$bg, "; border-radius:0 4px 4px 0; color:", cs$therapeutic_resistance$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no therapeutic resistance biomarkers (Tier 1&ndash;3) were identified among RNA fusions in this tumor sample.</div>"))
```

#### Prognostic biomarkers

```{r tier1_2_poor_prognosis_variants_fusion}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_poor$tier1$n_fusion > 0) || isTRUE(vstats$prognostic_poor$tier2$n_fusion > 0)

rctbl_recs <- pcgrr::prep_progn_display_tbl(
  rep = pcg_report,
  etype_for_tiering = c("prognostic"),
  variant_category = 'fusion',
  clnsig = "prognostic_poor")

result_table <-
  pcgrr::render_progn_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'fusion',
    clnsig = "prognostic_poor"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_poor$tier1$n_fusion == 0) && isTRUE(vstats$prognostic_poor$tier2$n_fusion == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$prognostic_poor$border, "; background:", cs$prognostic_poor$bg, "; border-radius:0 4px 4px 0; color:", cs$prognostic_poor$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no poor outcome biomarkers for ", tumor_site, " cancers were identified among RNA fusions in this tumor sample.</div>"))
```

<br>

```{r tier1_2_better_prognosis_variants_fusion}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_better$tier1$n_fusion > 0) || isTRUE(vstats$prognostic_better$tier2$n_fusion > 0)

rctbl_recs <- pcgrr::prep_progn_display_tbl(
  rep = pcg_report,
  etype_for_tiering = c("prognostic"),
  variant_category = 'fusion',
  clnsig = "prognostic_better")

result_table <-
  pcgrr::render_progn_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'fusion',
    clnsig = "prognostic_better"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_better$tier1$n_fusion == 0) && isTRUE(vstats$prognostic_better$tier2$n_fusion == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$prognostic_better$border, "; background:", cs$prognostic_better$bg, "; border-radius:0 4px 4px 0; color:", cs$prognostic_better$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no better outcome biomarkers for ", tumor_site, " cancers were identified among RNA fusions in this tumor sample.</div>"))
```

#### Diagnostic biomarkers

```{r tier1_2_diagnostic_fusions}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$diagnostic_positive$tier1$n_fusion > 0) || isTRUE(vstats$diagnostic_positive$tier2$n_fusion > 0)

rctbl_recs <- pcgrr::prep_diagn_display_tbl(
  rep = pcg_report,
  etype_for_tiering = c("diagnostic"),
  variant_category = "fusion",
  clnsig = "diagnostic_positive")

result_table <-
  pcgrr::render_diagn_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'fusion',
    clnsig = "diagnostic_positive"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$diagnostic_positive$tier1$n_fusion == 0) && isTRUE(vstats$diagnostic_positive$tier2$n_fusion == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$diagnostic_positive$border, "; background:", cs$diagnostic_positive$bg, "; border-radius:0 4px 4px 0; color:", cs$diagnostic_positive$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no positive diagnostic biomarkers for ", tumor_site, " cancers were identified among RNA fusions in this tumor sample.</div>"))
```

:::

