## Settings

```{r on_off_settings}
#| echo: false
#| eval: true

signatures_estimation <- 'ON'
msi_prediction <- 'ON'
mutational_burden <- 'ON'

conf <- pcg_report$settings$conf

oncokb_active <- !is.null(conf$oncokb$api_token) &&
  !is.na(conf$oncokb$api_token) &&
  conf$oncokb$api_token != "None"
bm_db_label <- if (oncokb_active && isTRUE(as.logical(conf$oncokb$exclusive))) {
  "OncoKB"
} else if (oncokb_active) {
  "CIViC, CGI, and OncoKB"
} else {
  "CIViC/CGI"
}
bm_db_citation <- if (oncokb_active && isTRUE(as.logical(conf$oncokb$exclusive))) {
  "[@Chakravarty2017-rd]"
} else if (oncokb_active) {
  "[@Griffith2017-do;@Tamborero2018-aj;@Chakravarty2017-rd]"
} else {
  "[@Griffith2017-do;@Tamborero2018-aj]"
}

sample_name <- pcg_report[['settings']][['sample_id']]
assay_type <- conf$assay_properties$type
tumor_site <- conf$sample_properties$site
sex <- conf$sample_properties$sex
assay_mode <- conf$assay_properties$mode
genome_assembly <- pcg_report[['settings']][['genome_assembly']]
tumor_ploidy <- "N/A"
if (!is.na(conf$sample_properties$tumor_ploidy) &
   conf$sample_properties$tumor_ploidy != "NA") {
  tumor_ploidy <- conf$sample_properties$tumor_ploidy
}
tumor_purity <- "N/A"
if (!is.na(conf$sample_properties$tumor_purity) &
   conf$sample_properties$tumor_purity != "NA") {
  tumor_purity <- conf$sample_properties$tumor_purity
}

target_size <- conf$assay_properties$effective_target_size_mb
report_color <- pcgrr::color_palette$report_color$values[1]
if (as.logical(conf$assay_properties$vcf_tumor_only) == TRUE) {
  report_color <- pcgrr::color_palette$report_color$values[2]
}

success_color <- pcgrr::color_palette$success
none_color <- pcgrr::color_palette$none

rna_expression_similarity <- 'OFF'
if (as.logical(conf$expression$similarity_analysis) == TRUE) {
  rna_expression_similarity <- 'ON'
}

include_germline_findings <- 'OFF'
if (pcg_report$content$germline_classified$eval == TRUE) {
  include_germline_findings <- 'ON'
}
if (pcg_report$content$msi$eval == FALSE) {
  msi_prediction <- 'OFF'
}
if (pcg_report$content$mutational_signatures$eval == FALSE) {
  signatures_estimation <- 'OFF'
}
if (pcg_report$content$tmb$eval == FALSE) {
  mutational_burden <- 'OFF'
}

snv_indel_eval <- pcg_report$content$snv_indel$eval
cna_eval <- pcg_report$content$cna$eval
fusion_eval <- pcg_report$content$fusion$eval
expression_eval <- pcg_report$content$expression$eval

tumor_dp_tag <- conf[['somatic_snv']][['allelic_support']][['tumor_dp_tag']]
tumor_af_tag <- conf[['somatic_snv']][['allelic_support']][['tumor_af_tag']]
control_dp_tag <- conf[['somatic_snv']][['allelic_support']][['control_dp_tag']]
control_af_tag <- conf[['somatic_snv']][['allelic_support']][['control_af_tag']]
tumor_dp_min <- conf[['somatic_snv']][['allelic_support']][['tumor_dp_min']]
tumor_af_min <- conf[['somatic_snv']][['allelic_support']][['tumor_af_min']]
tumor_ad_min <- conf[['somatic_snv']][['allelic_support']][['tumor_ad_min']]
control_dp_min <- conf[['somatic_snv']][['allelic_support']][['control_dp_min']]
control_af_max <- conf[['somatic_snv']][['allelic_support']][['control_af_max']]
control_ad_max <- conf[['somatic_snv']][['allelic_support']][['control_ad_max']]

gencode_metadata <-
  pcg_report$settings$reference_data$source_metadata |>
  dplyr::filter(source_abbreviation == "gencode")

transcript_set_gencode <- paste0("GENCODE - basic set (v", gencode_metadata$source_version, ")")
if (as.logical(pcg_report[["settings"]][["conf"]][["vep"]][["vep_gencode_basic"]]) == FALSE) {
  transcript_set_gencode <- paste0("GENCODE - all transcripts (v", gencode_metadata$source_version, ")")
}
```

<br><br>

```{r input_data_indicators}
#| output: asis
#| echo: false
#| eval: true

bslib::layout_column_wrap(
  width = 1/4,
  height = 90,
  bslib::value_box(
    title = "Data input",
    value = "Somatic SNVs/InDels",
    showcase = NULL,
    #showcase = bsicons::bs_icon(
    #  if (snv_indel_eval) "check-square" else "x-square", size = "0.4em"),
    #showcase_layout = "top right",
    theme = bslib::value_box_theme(
      bg = if (snv_indel_eval) success_color else none_color, fg = "#fff")
  ),
  bslib::value_box(
    title = "Data input",
    value = "Somatic CNAs",
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = if (cna_eval) success_color else none_color, fg = "#fff")
  ),
  bslib::value_box(
    title = "Data input",
    value = "RNA fusions",
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = if (fusion_eval) success_color else none_color, fg = "#fff")
  ),
  bslib::value_box(
    title = "Data input",
    value = "Gene expression",
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = if (expression_eval) success_color else none_color, fg = "#fff")
  )
)
```


```{r sample_properties}
#| output: asis
#| echo: false
#| eval: true

bslib::layout_column_wrap(
  width = 1/4,
  height = 90,
  bslib::value_box(
    title = "Tumor site",
    value = tumor_site,
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = report_color, fg = "#fff")
  ),
  bslib::value_box(
    title = "Sex",
    value = sex,
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = report_color, fg = "#fff")
  ),
  bslib::value_box(
    title = "Tumor purity",
    value = tumor_purity,
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = report_color, fg = "#fff")
  ),
  bslib::value_box(
    title = "Tumor ploidy",
    value = tumor_ploidy,
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = report_color, fg = "#fff")
  )
)
```


::: {.callout-note collapse="true"}

## Run configuration and report settings

The report is generated with _[pcgr](https://github.com/sigven/pcgr)_ version `r pcg_report[['settings']][['software']][['pcgr_version']]`, using the following key settings:

:::: {.columns}

::: {.column width="47.5%"}

**General**

* Genome assembly: __`r genome_assembly`__
* Variant actionability guidelines: __AMP/ASCO/CAP__
* Variant oncogenicity guidelines: __ClinGen/CGC/VICC__
* Biomarker sources: __`r if (as.logical(conf$oncokb$exclusive)) 'OncoKB only' else 'OncoKB + CIViC + CGI'`__

```{r snv_indel_settings}
#| output: asis
#| echo: false
#| eval: !expr pcg_report$content$snv_indel$eval == TRUE
cat("\n**SNV/InDel settings**\n\n")
cat(glue::glue("* Sequencing assay mode: __{assay_mode}__"), "\n")
cat(glue::glue("* Sequencing assay type: __{assay_type}__"), "\n")
cat(glue::glue("* Coding target size: __{target_size}Mb__"), "\n")
cat(glue::glue("* Show noncoding variants: __{conf$other$show_noncoding}__"), "\n")
cat(glue::glue("* MSI prediction: __{msi_prediction}__"), "\n")
cat(glue::glue("* Include germline findings: __{include_germline_findings}__"), "\n")
cat(glue::glue("* Variant Effect Predictor (VEP) settings"), "\n")
cat(glue::glue("  * Transcript set: __{transcript_set_gencode}__"), "\n")
cat(glue::glue("  * Transcript pick order: __{stringr::str_replace_all(conf[['vep']][['vep_pick_order']], ',', ', ')}__"), "\n")
cat(glue::glue("  * Regulatory regions annotation: __{as.logical(conf[['vep']][['vep_regulatory']])}__"), "\n")
cat(glue::glue("  * Buffer size: __{conf$vep$vep_buffer_size}__"), "\n")
cat(glue::glue("  * Number of forks: __{conf$vep$vep_n_forks}__"), "\n")
allelic_filter_items <- character(0)
if (tumor_dp_tag != "_NA_" & !is.null(tumor_dp_min))
  allelic_filter_items <- c(allelic_filter_items,
    glue::glue("  * Minimum sequencing depth (DP) tumor (SNV + InDels): __{tumor_dp_min}__"))
if (tumor_af_tag != "_NA_" & !is.null(tumor_af_min))
  allelic_filter_items <- c(allelic_filter_items,
    glue::glue("  * Minimum allelic fraction (AF) tumor (SNV + InDels): __{tumor_af_min}__"))
if (tumor_dp_tag != "_NA_" & tumor_af_tag != "_NA_" & !is.null(tumor_ad_min))
  allelic_filter_items <- c(allelic_filter_items,
    glue::glue("  * Minimum allelic depth (AD) tumor (SNV + InDels): __{tumor_ad_min}__"))
if (control_dp_tag != "_NA_" & !is.null(control_dp_min))
  allelic_filter_items <- c(allelic_filter_items,
    glue::glue("  * Minimum sequencing depth (DP) control (SNV + InDels): __{control_dp_min}__"))
if (control_af_tag != "_NA_" & !is.null(control_af_max))
  allelic_filter_items <- c(allelic_filter_items,
    glue::glue("  * Maximum allelic fraction (AF) control (SNV + InDels): __{control_af_max}__"))
if (control_dp_tag != "_NA_" & control_af_tag != "_NA_" & !is.null(control_ad_max))
  allelic_filter_items <- c(allelic_filter_items,
    glue::glue("  * Maximum allelic depth (AD) control (SNV + InDels): __{control_ad_max}__"))
if (length(allelic_filter_items) > 0) {
  cat("* Allelic support filtering settings", "\n")
  for(item in allelic_filter_items) cat(item, "\n")
}
```

:::

::: {.column width="5%"}
<!-- empty column to create gap -->
:::

::: {.column width="47.5%"}

```{r snv_indel_settings_right}
#| output: asis
#| echo: false
#| eval: !expr pcg_report$content$snv_indel$eval == TRUE
cat("**SNV/InDel settings (continued)**\n\n")
cat(glue::glue("* Mutational burden estimation: __{mutational_burden}__"), "\n")
cat(glue::glue("  * TMB algorithm: __TMB_{conf$somatic_snv[['tmb']][['tmb_display']]}__"), "\n")
cat(glue::glue("* Mutational signatures estimation: __{signatures_estimation}__"), "\n")
cat(glue::glue("  * Minimum number of mutations required: __{conf$somatic_snv$mutational_signatures[['mutation_limit']]}__"), "\n")
cat(glue::glue("  * All reference signatures: __{as.logical(conf$somatic_snv$mutational_signatures[['all_reference_signatures']])}__"), "\n")
cat(glue::glue("  * Inclusion of artefact signatures: __{as.logical(conf$somatic_snv$mutational_signatures[['include_artefact_signatures']])}__"), "\n")
cat(glue::glue("  * Minimum tumor-type prevalence (percent) of reference signatures used for refitting: __{conf$somatic_snv$mutational_signatures[['prevalence_reference_signatures']]}__"), "\n")
```

```{r cna_settings}
#| output: asis
#| echo: false
#| eval: !expr pcg_report$content$cna$eval == TRUE
cat("\n**CNA settings**\n\n")
cat(glue::glue("* Amplification threshold (absolute): __{conf$somatic_cna$amp_threshold_absolute}__"), "\n")
cat(glue::glue("* Amplification threshold (relative): __{conf$somatic_cna$amp_threshold_relative}__"), "\n")
cat(glue::glue("* CNA thresholding mode: __{conf$somatic_cna$threshold_mode}__"), "\n")
cat(glue::glue("* Effective amplification threshold: __{conf$somatic_cna$amp_threshold_effective}__"), "\n")
cat(glue::glue("* Gain threshold (absolute): __{conf$somatic_cna$gain_threshold_absolute}__"), "\n")
cat(glue::glue("* Gain threshold (relative): __{conf$somatic_cna$gain_threshold_relative}__"), "\n")
cat(glue::glue("* Heterozygous deletion threshold (absolute): __{conf$somatic_cna$del_threshold_absolute}__"), "\n")
cat(glue::glue("* Heterozygous deletion threshold (relative): __{conf$somatic_cna$del_threshold_relative}__"), "\n")
```

```{r expression_settings}
#| output: asis
#| echo: false
#| eval: !expr as.logical(pcg_report$content$expression$eval) == TRUE
cat("\n**Expression settings**\n\n")
cat(glue::glue("* RNA expression similarity analysis: __{rna_expression_similarity}__"), "\n")
```

:::

::::

:::

<br>
