## Settings

```{r on_off_settings}
#| echo: false
#| eval: true

signatures_estimation <- 'ON'
msi_prediction <- 'ON'
mutational_burden <- 'ON'

conf <- pcg_report$settings$conf

sample_name <- pcg_report[['settings']][['sample_id']]
assay_type <- conf$assay_properties$type
tumor_site <- conf$sample_properties$site
assay_mode <- conf$assay_properties$mode
tumor_ploidy <- "N/A"
if(!is.na(conf$sample_properties$ploidy) &
   conf$sample_properties$ploidy != "NA"){
  tumor_ploidy <- conf$sample_properties$ploidy
}
tumor_purity <- "N/A"
if(!is.na(conf$sample_properties$purity) &
   conf$sample_properties$purity != "NA"){
  tumor_purity <- conf$sample_properties$purity
}

target_size <- conf$assay_properties$effective_target_size_mb
report_color <- pcgrr::color_palette$report_color$values[1]
if(as.logical(conf$assay_properties$vcf_tumor_only) == TRUE){
  report_color <- pcgrr::color_palette$report_color$values[2]
}

success_color <- pcgrr::color_palette$success

if(pcg_report$content$msi$eval == F){
  msi_prediction <- 'OFF'
}
if(pcg_report$content$mutational_signatures$eval == F){
  signatures_estimation <- 'OFF'
}
if(pcg_report$content$tmb$eval == F){
  mutational_burden <- 'OFF'
}

gencode_metadata <- 
  pcg_report$settings$reference_data$source_metadata |>
  dplyr::filter(source_abbreviation == "gencode")

transcript_set_gencode <- paste0("GENCODE - basic set (v", gencode_metadata$source_version, ")")
if (as.logical(pcg_report[["settings"]][["conf"]][["vep"]][["vep_gencode_basic"]]) == F) {
  transcript_set_gencode <- paste0("GENCODE - all transcripts (v", gencode_metadata$source_version, ")")
}
```

The report is generated with _[pcgr](https://github.com/sigven/pcgr)_ version `r pcg_report[['settings']][['software']][['pcgr_version']]`, using the following key settings:

:::: {.columns}

::: {.column width="47.5%"}
* Genome assembly: __`r pcg_report[['settings']][['genome_assembly']]`__
* Minimum sequencing depth (DP) tumor (SNV + InDels): __`r conf[['somatic_snv']][['allelic_support']][['tumor_dp_min']]`__
* Minimum allelic fraction (AF) tumor (SNV + InDels): __`r conf[['somatic_snv']][['allelic_support']][['tumor_af_min']]`__
* Minimum sequencing depth (DP) control (SNV + InDels): __`r conf[['somatic_snv']][['allelic_support']][['control_dp_min']]`__
* Maximum allelic fraction (AF) control (SNV + InDels): __`r conf[['somatic_snv']][['allelic_support']][['control_af_max']]`__
* Actionability framework: __ACMG/AMP__
* Oncogenicity framework: __VICC/ClinGen__
* Show noncoding variants: __`r conf$other$show_noncoding `__
* MSI prediction: __`r msi_prediction`__
* Mutational burden estimation: __`r mutational_burden`__
  * TMB algorithm: __`r paste0('TMB_',conf$somatic_snv[['tmb']][['tmb_display']])`__ 
:::

::: {.column width="5%"}
<!-- empty column to create gap -->
:::

::: {.column width="47.5%"}
* Mutational signatures estimation: __`r signatures_estimation`__
  * Minimum number of mutations required: __`r conf$somatic_snv$mutational_signatures[['mutation_limit']]`__
  * All reference signatures: __`r as.logical(conf$somatic_snv$mutational_signatures[['all_reference_signatures']])`__
  * Inclusion of artefact signatures: __`r as.logical(conf$somatic_snv$mutational_signatures[['include_artefact_signatures']])`__
  * Minimum tumor-type prevalence (percent) of reference signatures used for refitting: __`r conf$somatic_snv$mutational_signatures[['prevalence_reference_signatures']]`__
* Variant Effect Predictor (VEP) settings:
  * Transcript set: __`r transcript_set_gencode`__
  * Transcript pick order: __`r stringr::str_replace_all(conf[['vep']][['vep_pick_order']],",", ", ")`__
  * Regulatory regions annotation: __`r as.logical(conf[['vep']][['vep_regulatory']])`__
  * Buffer size: __`r conf$vep$vep_buffer_size`__
  * Number of forks: __`r conf$vep$vep_n_forks`__

:::

::::

<br>


```{r sample_assay_properties}
#| output: asis
#| echo: false
#| eval: true

bslib::layout_column_wrap(
  width = 1/3,
  height = 85,
  bslib::value_box(
    title = "Tumor site",
    value = tumor_site,
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = report_color, fg = "#fff")
  ),
  bslib::value_box(
    title = "Tumor purity",
    value = tumor_purity,
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = report_color, fg = "#fff")
  ),
  bslib::value_box(
    title = "Tumor ploidy",
    value = tumor_ploidy,
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = report_color, fg = "#fff")
  )
)

bslib::layout_column_wrap(
  width = 1/3,
  height = 85,
  bslib::value_box(
    title = "Sequencing assay mode",
    value = assay_mode,
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = report_color, fg = "#fff")
  ),
  bslib::value_box(
    title = "Sequencing assay type",
    value = assay_type,
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = report_color, fg = "#fff")
  ),
  bslib::value_box(
    title = "Coding target size",
    value = paste0(target_size, "Mb"),
    showcase = NULL,
    theme = bslib::value_box_theme(
      bg = report_color, fg = "#fff")
  )
)

```

```{r input_data}
#| output: asis
#| echo: false
#| eval: false

bslib::layout_column_wrap(
  width = 1/4,
  height = 80,
  bslib::value_box(
    title = "",
    value = "DNA - SNVs/InDels",
    showcase = bsicons::bs_icon("check-square", size = "0.6em"),
    showcase_layout = "top right",
    theme = bslib::value_box_theme(
      bg = pcgrr::color_palette$success, fg = "#fff")
  ),
  bslib::value_box(
    title = "",
    value = "DNA - CNAs",
    showcase = bsicons::bs_icon("check-square", size = "0.6em"),
    showcase_layout = "top right",
    theme = bslib::value_box_theme(
      bg = pcgrr::color_palette$success, fg = "#fff")
  ),
  bslib::value_box(
    title = "",
    value = "RNA - gene fusions",
    showcase = bsicons::bs_icon("check-square", size = "0.6em"),
    showcase_layout = "top right",
    theme = bslib::value_box_theme(
      bg = pcgrr::color_palette$success, fg = "#fff")
  ),
  bslib::value_box(
    title = "",
    value = "RNA - gene expression",
    showcase = bsicons::bs_icon("check-square", size = "0.6em"),
    showcase_layout = "top right",
    theme = bslib::value_box_theme(
      bg = pcgrr::color_palette$success, fg = "#fff")
  )
)
```



