### Clinical actionability classification

```{r actionability_setup}
#| output: asis
#| echo: false
#| eval: true

vstats <- pcg_report$content$snv_indel$vstats
tumor_site <- pcg_report$settings$conf$sample_properties$site
cs <- pcgrr::color_palette$bm_callout
```

Acquired SNVs and InDels in the tumor sample have been classified into tiers of _clinical actionability_ - adopting the AMP/ASCO/CAP guidelines [@Li2017-ew]. We use `r bm_db_label` `r bm_db_citation` as the underlying sources for actionable variants (versions indicated [below](#documentation)).


::: {.callout-note collapse="false"}
## Note — biomarker types and report scope

Tier classification is driven by **therapeutic sensitivity** biomarkers (i.e. those related to treatment response). Evidence items where variants match **resistance**, **prognostic**, or **diagnostic** biomarkers are also shown — all restricted to biomarkers matching the tumor type of the query sample.

Note that the _complete_ set of **Tier III** variants (uncertain/unknown clinical significance) are only available in the TSV and Excel workbook outputs.

:::



```{r}
#| echo: false
#| output: asis
#| include: !expr isTRUE(vstats$therapeutic_sensitivity$tier1$n_var > 0) || isTRUE(vstats$therapeutic_sensitivity$tier2$n_var > 0) || isTRUE(vstats$therapeutic_sensitivity$tier3$n_var > 0) || isTRUE(vstats$therapeutic_resistance$tier1$n_var > 0) || isTRUE(vstats$therapeutic_resistance$tier2$n_var > 0) || isTRUE(vstats$therapeutic_resistance$tier3$n_var > 0)

cat(paste0(
  "\n::: {.callout-important collapse='false'}\n## Note - biomarker resolution and matching against query variants\n\n",
  "A somatic aberration in the tumor sample may frequently match with multiple evidence items listed in biomarker databases (CIViC/CGI/OncoKB), pending on e.g. the _disease/tumor context_, _therapeutic context_, or underlying _evidence source_ (i.e. publication). Biomarkers are furthermore recorded at different _resolutions_.",
  " The accuracy of a match between a variant ",
  "in the tumor sample and a clinical evidence item (biomarker) will vary accordingly, ",
  "as indicated by the styling of the <b>Alteration</b> column below:\n\n",
  "<div style='margin-bottom:6px;'>",
  "<span style='display:inline-block; width:9px; height:9px; border-radius:50%; ",
  "background-color:#37303A; margin-right:8px; vertical-align:middle;'></span>",
  "<span style='font-weight:600; color:#1a1a1a; vertical-align:middle;'>Bold text + filled circle</span>",
  " &mdash; high-resolution match (genomic position / amino acid / codon level)",
  "</div>",
  "<div style='margin-bottom:6px;'>",
  "<span style='display:inline-block; width:9px; height:9px; border-radius:50%; ",
  "background-color:transparent; border:2px solid #999999; margin-right:8px; vertical-align:middle;'></span>",
  "<span style='font-weight:400; color:#666666; vertical-align:middle;'>Normal text + hollow circle</span>",
  " &mdash; medium-resolution match (exon / gene level)",
  "</div>",
  "<p style='margin-top:8px; font-size:0.99em; color:#555;'>",
  "<em>Note:</em> Variants matched at exon/gene level (medium resolution) are limited to those ",
  "with an <b>oncogenic</b> or <b>likely oncogenic</b> classification, or overlapping with a ",
  "<b>mutation hotspot</b>.",
  "</p>\n\n:::"
))
```

<br>

::: {.panel-tabset}

#### Therapeutic biomarkers

```{r tier_legend}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_sensitivity$tier1$n_var > 0) || isTRUE(vstats$therapeutic_sensitivity$tier2$n_var > 0)

cat(paste0(
  "\n<div style='margin-top:10px; padding:10px 14px; border:1px solid #e0e0e0; ",
  "border-radius:6px; background:#fafafa; font-size:0.97em;'>",
  "<b>Tier color legend - AMP/ASCO/CAP</b>",
  "<div style='margin-top:6px;'>",
  "<div style='margin-bottom:4px;'><span style='display:inline-block; padding:2px 10px; border-radius:4px; font-size:0.90em; font-weight:600; background-color:#1565C0; color:white; margin-right:6px;'>T1</span>Tier 1 - variants of strong clinical significance</div>",
  "<div style='margin-bottom:4px;'><span style='display:inline-block; padding:2px 10px; border-radius:4px; font-size:0.90em; font-weight:600; background-color:#42A5F5; color:white; margin-right:6px;'>T2</span>Tier 2 - variants of potential clinical significance</div>",
  "<div style='margin-bottom:4px;'><span style='display:inline-block; padding:2px 10px; border-radius:4px; font-size:0.90em; font-weight:600; background-color:#868686; color:white; margin-right:6px;'>T3</span>Tier 3 - variants of uncertain significance</div>",
  "</div></div>\n"
))
htmltools::br()

```


```{r tier1_2_sensitivity_variants_snv}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_sensitivity$tier1$n_var > 0) || isTRUE(vstats$therapeutic_sensitivity$tier2$n_var > 0)

rctbl_recs <- pcgrr::prep_actble_display_tbl(
  rep = pcg_report, 
  tier = c(1,2,3),
  etype_for_tiering = c("predictive"),
  variant_category = 'snv_indel')

result_table <- 
  pcgrr::render_actble_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'snv_indel',
    clnsig = "therapeutic_sensitivity"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_sensitivity$tier1$n_var == 0) && isTRUE(vstats$therapeutic_sensitivity$tier2$n_var == 0) && isTRUE(vstats$therapeutic_sensitivity$tier3$n_var == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$therapeutic_sensitivity$border, "; background:", cs$therapeutic_sensitivity$bg, "; border-radius:0 4px 4px 0; color:", cs$therapeutic_sensitivity$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no therapeutic sensitivity biomarkers (Tier 1&ndash;3) were identified among SNVs/InDels in this tumor sample.</div>"))
```


<br>

```{r tier1_2_resistance_variants_snv}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_resistance$tier1$n_var > 0) || isTRUE(vstats$therapeutic_resistance$tier2$n_var > 0)

rctbl_recs <- pcgrr::prep_actble_display_tbl(
  rep = pcg_report, 
  clnsig = "therapeutic_resistance",
  variant_category = 'snv_indel')

result_table <- 
  pcgrr::render_actble_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'snv_indel',
    clnsig = "therapeutic_resistance"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$therapeutic_resistance$tier1$n_var == 0) && isTRUE(vstats$therapeutic_resistance$tier2$n_var == 0) && isTRUE(vstats$therapeutic_resistance$tier3$n_var == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$therapeutic_resistance$border, "; background:", cs$therapeutic_resistance$bg, "; border-radius:0 4px 4px 0; color:", cs$therapeutic_resistance$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no therapeutic resistance biomarkers for ", tumor_site, " cancers were identified among SNVs/InDels in this tumor sample.</div>"))
```

#### Prognostic biomarkers

```{r tier1_2_poor_prognosis_variants_snv}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_poor$tier1$n_var > 0) || isTRUE(vstats$prognostic_poor$tier2$n_var > 0)

rctbl_recs <- pcgrr::prep_progn_display_tbl(
  rep = pcg_report,
  etype_for_tiering = c("prognostic"),
  variant_category = 'snv_indel',
  clnsig = "prognostic_poor")

result_table <-
  pcgrr::render_progn_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'snv_indel',
    clnsig = "prognostic_poor"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_poor$tier1$n_var == 0) && isTRUE(vstats$prognostic_poor$tier2$n_var == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$prognostic_poor$border, "; background:", cs$prognostic_poor$bg, "; border-radius:0 4px 4px 0; color:", cs$prognostic_poor$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no poor outcome biomarkers for ", tumor_site, " cancers were identified among SNVs/InDels in this tumor sample.</div>"))
```

<br>

```{r tier1_2_better_prognosis_variants_snv}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_better$tier1$n_var > 0) || isTRUE(vstats$prognostic_better$tier2$n_var > 0)

rctbl_recs <- pcgrr::prep_progn_display_tbl(
  rep = pcg_report,
  etype_for_tiering = c("prognostic"),
  variant_category = 'snv_indel',
  clnsig = "prognostic_better")

result_table <-
  pcgrr::render_progn_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'snv_indel',
    clnsig = "prognostic_better"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$prognostic_better$tier1$n_var == 0) && isTRUE(vstats$prognostic_better$tier2$n_var == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$prognostic_better$border, "; background:", cs$prognostic_better$bg, "; border-radius:0 4px 4px 0; color:", cs$prognostic_better$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no better outcome biomarkers for ", tumor_site, " cancers were identified among SNVs/InDels in this tumor sample.</div>"))
```

#### Diagnostic biomarkers


```{r tier1_2_positive_diagnosis_variants_snv}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$diagnostic_positive$tier1$n_var > 0) || isTRUE(vstats$diagnostic_positive$tier2$n_var > 0)

rctbl_recs <- pcgrr::prep_diagn_display_tbl(
  rep = pcg_report,
  etype_for_tiering = c("diagnostic"),
  variant_category = 'snv_indel',
  clnsig = "diagnostic_positive")

result_table <-
  pcgrr::render_diagn_bm_table(
    rctbl_recs = rctbl_recs,
    variant_category = 'snv_indel',
    clnsig = "diagnostic_positive"
  )

result_table

```

```{r}
#| echo: false
#| output: asis
#| eval: !expr isTRUE(vstats$diagnostic_positive$tier1$n_var == 0) && isTRUE(vstats$diagnostic_positive$tier2$n_var == 0)
cat(paste0("<div style='padding:8px 14px; border-left:4px solid ", cs$diagnostic_positive$border, "; background:", cs$diagnostic_positive$bg, "; border-radius:0 4px 4px 0; color:", cs$diagnostic_positive$text, "; font-size:0.92em;'>Based on available data in ", bm_db_label, ", no diagnostic biomarkers for ", tumor_site, " cancers were identified among SNVs/InDels in this tumor sample.</div>"))
```


:::

<br>
