### Variant classification - actionability

Acquired SNVs and InDels in the tumor sample have been assessed for
_clinical actionability_ through an implementation of ACMG/AMP guidelines [@Li2017-ew]. We use CIViC and CGI [@Griffith2017-do;@Tamborero2018-aj] as the underlying sources for actionable variants (versions indicated [below](#documentation)). 

The tabset below indicates which variants have been classified into three distinct _tiers of clinical significance_ according to ACMG/AMP criteria.

::: {.callout-note}
## Note - biomarker matching

The same somatic aberration may frequently match with multiple evidence items listed in biomarker databases , pending on e.g. the _disease context_, _therapeutic context_, _biomarker resolution_, or underlying _evidence source_ (i.e. publication). Duplicated entries in the table below may thus also occur, as the same type of evidence from different sources (publications) have not
been merged.
:::

```{r}
#| echo: false
#| output: asis
#| include: !expr as.logical(pcg_report$content$snv_indel$vstats$n_actionable_tier1 != 0) == TRUE | as.logical(pcg_report$content$snv_indel$vstats$n_actionable_tier2 != 0) == TRUE | as.logical(pcg_report$content$snv_indel$vstats$n_actionable_tier3 != 0) == TRUE

cat("\n::: {.callout-note}\n## Note - biomarker resolution\n\n Biomarkers registered in CIViC/CGI are provided at different _resolutions_ (i.e. filter <b>Biomarker resolution</b>). The accuracy of a match between a variant in the sample and a clinical evidence item (biomarker) will vary accordingly (highlighted by sample alterations with different color backgrounds):\n\n")

cat('<ul><li>Biomarker match at the <mark style="background-color:#2c313c; font-weight:bold; color:white">&nbsp; genomic position/amino acid/codon level &nbsp;</mark></li>')
cat(paste0(
  '<br><li>Biomarker match at the <mark style="background-color:', 
  pcgrr::color_palette$warning,
  '; color:white; font-weight:bold">&nbsp; exon/gene level&nbsp; </mark></li></ul>\n\n:::'))
htmltools::br()

```

::: {.panel-tabset .nav-pills}

#### <b>TIER I</b> - strong clinical significance

<br>

```{r tier1_variants, echo=FALSE}
#| echo: false
#| output: asis
#| eval: !expr as.logical(pcg_report$content$snv_indel$vstats$n_actionable_tier1 != 0) == TRUE

dt_snv_content_tier1 <- 
  pcgrr:::get_dt_tables(
    report = pcg_report, tier = 1, variant_class = "snv_indel")

# if(expression_present_snv == FALSE){
#   dt_snv_content_tier1[['by_gene']]$consTPM <- NULL
# }

tier1_gene_dt <- DT::datatable(
  dt_snv_content_tier1[['by_gene']],
  escape = F, 
  options = list(dom = 't'),
  extensions = c("Responsive")) |>
  DT::formatStyle(
    color="white", 
    "MOLECULAR_ALTERATION", 
    "BM_TOP_RESOLUTION", 
    fontWeight = 'bold', 
    `text-align` = 'center',
    backgroundColor = DT::styleEqual(
      c('high','low'), 
      c(pcgrr::color_palette$bg_dark,
        pcgrr::color_palette[['warning']])
      )
  ) |>
  DT::formatStyle(
    'BIOMARKER_EVIDENCE',
    'BM_TOP_RESOLUTION',
    color = "white",
     backgroundColor = DT::styleEqual(
      c('high','low'), 
      c(pcgrr::color_palette$bg_dark,
        pcgrr::color_palette[['warning']])
      )
  )

# if(expression_present_snv == TRUE){
#   tier1_gene_dt <- tier1_gene_dt |>
#     DT::formatStyle(
#      "consTPM",
#      "consTPM", 
#      color = "white", 
#      backgroundColor = 
#        DT::styleInterval(
#          pcgrr::color_palette$gene_expression$breaks,
#          pcgrr::color_palette$gene_expression$values
#        )
#    )
# }

tier1_gene_dt

htmltools::br()
htmltools::hr()
htmltools::br()

cat(paste0(
  "<font style='font-size: 110%'>Details with respect to the clinical",
  " evidence items associated with <b>",
  paste(dt_snv_content_tier1[['by_gene']]$MOLECULAR_ALTERATION, collapse=", "),
  " </b>can be explored in the data table listed below:</font>\n"))
htmltools::br()

snv_eitem_tier1_shared <- crosstalk::SharedData$new(
  dt_snv_content_tier1[['by_eitem']] |>
    dplyr::rename(SAMPLE_ALTERATION = MOLECULAR_ALTERATION))

crosstalk::bscols(
  list(
     crosstalk::filter_select(
      "SAMPLE_ALTERATION", "Sample alteration", 
      snv_eitem_tier1_shared, ~SAMPLE_ALTERATION),
    crosstalk::filter_select(
      "BM_CANCER_TYPE", "Cancer type", 
      snv_eitem_tier1_shared, ~BM_CANCER_TYPE),
    crosstalk::filter_select(
      "BM_CLINICAL_SIGNIFICANCE", "Clinical significance", 
      snv_eitem_tier1_shared, ~BM_CLINICAL_SIGNIFICANCE),
    crosstalk::filter_slider(
      "BM_RATING", "Rating", snv_eitem_tier1_shared, 
      ~BM_RATING, min = 0, max = 5, step = 1, ticks = T),
    crosstalk::filter_select(
      "BM_RESOLUTION", "Biomarker resolution", 
      snv_eitem_tier1_shared, ~BM_RESOLUTION)
  ),
  list(
    crosstalk::filter_select(
      "BM_EVIDENCE_TYPE", "Evidence type", 
      snv_eitem_tier1_shared, ~BM_EVIDENCE_TYPE),
    crosstalk::filter_select(
      "BM_EVIDENCE_LEVEL", "Evidence level", 
      snv_eitem_tier1_shared, ~BM_EVIDENCE_LEVEL),
    crosstalk::filter_select(
      "BM_EVIDENCE_DIRECTION", "Evidence direction", 
      snv_eitem_tier1_shared, ~BM_EVIDENCE_DIRECTION),
    crosstalk::filter_select(
      "BM_THERAPEUTIC_CONTEXT", "Therapeutic context", 
      snv_eitem_tier1_shared, ~BM_THERAPEUTIC_CONTEXT)
  )
)

htmltools::br()

DT::datatable(
  snv_eitem_tier1_shared,
  escape = F, 
  extensions = c("Buttons","Responsive"), 
  options = list(
    buttons = c('csv','excel'), 
    dom = 'Bfrtip')) |>
  DT::formatStyle(
    color = "white",
    'BM_EVIDENCE_LEVEL', 
    backgroundColor = DT::styleEqual(
      pcgrr::color_palette$clinical_evidence$levels,
      pcgrr::color_palette$clinical_evidence$values)) |>
  DT::formatStyle(
    color = "white", 
    c("SAMPLE_ALTERATION","BM_CONTEXT"), 
    "BM_RESOLUTION", 
    fontWeight = 'bold', 
    `text-align` = 'center',
    backgroundColor = DT::styleEqual(
      c('genomic','hgvsp','codon',
        'exon','gene_region_mut',
        'gene_lof','gene_mut'), 
      c(pcgrr::color_palette$bg_dark,
        pcgrr::color_palette$bg_dark,
        pcgrr::color_palette$bg_dark,
        pcgrr::color_palette[['warning']],
        pcgrr::color_palette[['warning']],
        pcgrr::color_palette[['warning']],
        pcgrr::color_palette[['warning']])
    )
  )

```


```{r missing_tier1_variants}
#| output: asis
#| echo: false
#| eval: !expr as.logical(pcg_report$content$snv_indel$vstats$n_actionable_tier1 == 0) == TRUE

htmltools::br()
cat('\n*  <i> <font style="font-size: 110%"><b>No</b> somatic SNVs/InDels of strong clinical significance were found.</font></i>', sep = '\n')
cat('\n')
htmltools::br()

```

#### <b>TIER II</b>- potential clinical significance


```{r tier2_variants}
#| echo: false
#| output: asis
#| eval: !expr as.logical(pcg_report$content$snv_indel$vstats$n_actionable_tier2 != 0) == TRUE

dt_snv_content_tier2 <- 
  pcgrr:::get_dt_tables(
    report = pcg_report, 
    tier = 2, 
    variant_class = "snv_indel")

# if(expression_present_snv == FALSE){
#   dt_snv_content_tier2[['by_gene']]$consTPM <- NULL
# }

tier2_gene_dt <- DT::datatable(
  dt_snv_content_tier2[['by_gene']],
  escape = F, 
  options = list(dom = 't'),
  extensions = c("Responsive")) |>
  DT::formatStyle(
    color="white", 
    "MOLECULAR_ALTERATION", 
    "BM_TOP_RESOLUTION", 
    fontWeight = 'bold', 
    `text-align` = 'center',
    backgroundColor = DT::styleEqual(
      c('high','low'), 
      c(pcgrr::color_palette$bg_dark,
        pcgrr::color_palette[['warning']])
      )
  ) |>
  DT::formatStyle(
    'BIOMARKER_EVIDENCE',
    'BM_TOP_RESOLUTION',
    color = "white",
     backgroundColor = DT::styleEqual(
      c('high','low'), 
      c(pcgrr::color_palette$bg_dark,
        pcgrr::color_palette[['warning']])
      )
  ) 

# if(expression_present_snv == TRUE){
#   tier2_gene_dt <- tier2_gene_dt |>
#     DT::formatStyle(
#      "consTPM",
#      "consTPM", 
#      color = "white", 
#      backgroundColor = 
#        DT::styleInterval(
#          pcgrr::color_palette$gene_expression$breaks,
#          pcgrr::color_palette$gene_expression$values
#        )
#    )
# }

tier2_gene_dt

htmltools::br()
htmltools::hr()
htmltools::br()
cat(paste0(
  "<font style='font-size: 110%'>Details with respect to the clinical",
  " evidence items associated with <b>",
  paste(dt_snv_content_tier2[['by_gene']]$MOLECULAR_ALTERATION, collapse=", "),
  " </b>can be explored in the data table listed below:</font>\n"))
htmltools::br()
htmltools::br()

snv_eitem_tier2_shared <- crosstalk::SharedData$new(
  dt_snv_content_tier2[['by_eitem']] |>
    dplyr::rename(SAMPLE_ALTERATION = MOLECULAR_ALTERATION))

crosstalk::bscols(
  list(
     crosstalk::filter_select(
      "SAMPE_ALTERATION", "Sample alteration", 
      snv_eitem_tier2_shared, ~SAMPLE_ALTERATION),
    crosstalk::filter_select(
      "BM_CANCER_TYPE", "Cancer type", 
      snv_eitem_tier2_shared, ~BM_CANCER_TYPE),
    crosstalk::filter_select(
      "BM_CLINICAL_SIGNIFICANCE", "Clinical significance", 
      snv_eitem_tier2_shared, ~BM_CLINICAL_SIGNIFICANCE),
    crosstalk::filter_slider(
      "BM_RATING", "Rating", snv_eitem_tier2_shared, 
      ~BM_RATING, min = 0, max = 5, step = 1, ticks = T),
    crosstalk::filter_select(
      "BM_RESOLUTION", "Biomarker resolution", 
      snv_eitem_tier2_shared, ~BM_RESOLUTION)

  ),
  list(
    crosstalk::filter_select(
      "BM_EVIDENCE_TYPE", "Evidence type", 
      snv_eitem_tier2_shared, ~BM_EVIDENCE_TYPE),
    crosstalk::filter_select(
      "BM_EVIDENCE_LEVEL", "Evidence level", 
      snv_eitem_tier2_shared, ~BM_EVIDENCE_LEVEL),
    crosstalk::filter_select(
      "BM_EVIDENCE_DIRECTION", "Evidence direction", 
      snv_eitem_tier2_shared, ~BM_EVIDENCE_DIRECTION),
    crosstalk::filter_select(
      "BM_THERAPEUTIC_CONTEXT", "Therapeutic context", 
      snv_eitem_tier2_shared, ~BM_THERAPEUTIC_CONTEXT)
  )
)

htmltools::br()

DT::datatable(
  snv_eitem_tier2_shared,
  escape = F, 
  extensions = c("Buttons","Responsive"), 
  options = list(
    buttons = c('csv','excel'), 
    dom = 'Bfrtip')) |>
  DT::formatStyle(
    color = "white",
    'BM_EVIDENCE_LEVEL', 
    backgroundColor = DT::styleEqual(
      pcgrr::color_palette$clinical_evidence$levels,
      pcgrr::color_palette$clinical_evidence$values)) |>
  DT::formatStyle(
    color = "white", 
    c("SAMPLE_ALTERATION",
      "BM_CONTEXT"), 
    "BM_RESOLUTION", 
    fontWeight = 'bold', 
    `text-align` = 'center',
    backgroundColor = DT::styleEqual(
      c('hgvsp','genomic','codon',
        'exon','gene_region_mut',
        'gene_lof','gene_mut'), 
      c(pcgrr::color_palette$bg_dark,
        pcgrr::color_palette$bg_dark,
        pcgrr::color_palette$bg_dark,
        pcgrr::color_palette[['warning']],
        pcgrr::color_palette[['warning']],
        pcgrr::color_palette[['warning']],
        pcgrr::color_palette[['warning']])
    )
  )

```


```{r }
#| output: asis
#| echo: false
#| eval: !expr as.logical(pcg_report$content$snv_indel$vstats$n_actionable_tier2 == 0) == TRUE

htmltools::br()
cat('\n*  <i> <font style="font-size: 110%"><b>No</b> somatic SNVs/InDels of potential clinical significance were found.</font></i>', sep = '\n')
cat('\n')
htmltools::br()

```

#### <b>TIER III</b> - uncertain clinical significance

Here, we list protein-altering somatic variants located in proto-oncogenes and tumor suppressors that are currently _not_ linked to any clinical evidence items (as specified in [@Li2017-ew]). These variants are considered to have uncertain clinical significance and are thus not recommended for clinical action.

Variants have been ranked based on their oncogenicity score, and their strength of association to cancer at the gene level (here, specifically __`r pcg_report$settings$conf$sample_properties$site`__ cancers), from <span style="background-color:`r pcgrr::color_palette$cancer_assoc$values[1]`; padding:2x; color:white"><b>&nbsp; weak associations &nbsp; </b></span>&nbsp; to <span style="background-color:`r pcgrr::color_palette$cancer_assoc$values[length(pcgrr::color_palette$cancer_assoc$values)]`; padding:2x; color:white"><b>&nbsp; strong associations &nbsp; </b></span> [@Ochoa2021-jc;@Nakken2023-ol]. 
<br>

```{r }
#| output: asis
#| echo: false
#| eval: !expr as.logical(pcg_report$content$snv_indel$vstats$n_actionable_tier3 != 0) == TRUE

tier3_variants <- 
  pcg_report$content[['snv_indel']]$callset$variant_display |>
  dplyr::filter(.data$ACTIONABILITY_TIER == 3) |>
  dplyr::select(
    dplyr::any_of(
      pcgrr::dt_display$snv_indel_tier3
    )
  ) |>
  dplyr::arrange(
    dplyr::desc(.data$ONCOGENICITY_SCORE),
    dplyr::desc(.data$TISSUE_ASSOC_RANK),
    dplyr::desc(.data$GLOBAL_ASSOC_RANK),
  )

# if(expression_present_snv == FALSE){
#   tier3_variants$consTPM <- NULL
# }


tier3_dt <- DT::datatable(
  tier3_variants,
  escape = F, 
  extensions = c("Buttons","Responsive"), 
  options = list(
    buttons = c('csv','excel'), 
    dom = 'Bfrtip')) |>
   DT::formatStyle(
     "SYMBOL",
     "TISSUE_ASSOC_RANK", 
     color = "white", 
     backgroundColor = 
       DT::styleInterval(
         pcgrr::color_palette$cancer_assoc$breaks,
         pcgrr::color_palette$cancer_assoc$values
       )
   )

# if(expression_present_snv == TRUE){
#   tier3_dt <- tier3_dt |>
#     DT::formatStyle(
#      "consTPM",
#      "consTPM", 
#      color = "white", 
#      backgroundColor = 
#        DT::styleInterval(
#          pcgrr::color_palette$gene_expression$breaks,
#          pcgrr::color_palette$gene_expression$values
#        )
#    )
# }

tier3_dt



```

```{r }
#| output: asis
#| echo: false
#| eval: !expr as.logical(pcg_report$content$snv_indel$vstats$n_actionable_tier3 == 0) == TRUE

htmltools::br()
cat('\n*  <i> <font style="font-size: 110%"><b>No</b> somatic SNVs/InDels of uncertain clinical significance were found.</font></i>', sep = '\n')
cat('\n')
htmltools::br()

```

:::
