```{r variant_filtering_prep}
#| eval: true
#| output: asis

to_settings <- pcg_report$settings$conf$somatic_snv$tumor_only

if(NROW(pcg_report$content$snv_indel$callset$variant_unfiltered) > 0 &
   "SOMATIC_CLASSIFICATION" %in% colnames(pcg_report$content$snv_indel$callset$variant_unfiltered)){
 
  filtering_stats <- 
    plyr::count(
      pcg_report$content$snv_indel$callset$variant_unfiltered$SOMATIC_CLASSIFICATION
    ) |>
    dplyr::arrange(dplyr::desc(freq)) |>
    dplyr::rename(FILTER = x, FREQUENCY = freq) |>
    dplyr::mutate(FILTER = stringr::str_replace_all(FILTER, "\\|", ", ")) |>
    dplyr::mutate(FILTER = dplyr::if_else(
      !(FILTER %in% pcgrr::germline_filter_levels) & FILTER != "SOMATIC",
      "MULTIPLE FILTERS",
      FILTER
    )) |>
    dplyr::group_by(FILTER) |>
    dplyr::reframe(FREQUENCY = sum(FREQUENCY)) |>
    dplyr::arrange(dplyr::desc(FREQUENCY)) |>
    dplyr::mutate(FILTER = factor(
      FILTER, levels = pcgrr::germline_filter_levels)) |>
    dplyr::mutate(PERCENT = scales::percent(
      FREQUENCY / sum(FREQUENCY), accuracy = 0.1))
}
  
```

### Variant filtering

In an effort to minimize the presence/impact of germline events from input variants (SNVs/InDels) called through a **tumor-only assay**, we here show the results of germline variant filters applied on the raw input set. The filters provide a classification of variants as somatic or likely germline (catched by various filters), the latter set not included for downstream analysis in this report.

::: {.panel-tabset}

#### Filtering results

```{r variant_filtering_pie}
#| eval: true
#| output: asis

t <- list(
  family = "Helvetica",
  size = 15)

variant_filter_pie <- 
  plotly::plot_ly(
    filtering_stats, 
    marker = list(
      colors = pcgrr::color_palette$tier$values,
      line = list(color = '#FFFFFF', width = 1)
      )
    ) |>
  plotly::add_pie(
    filtering_stats, 
    labels =~ FILTER,
    values = ~FREQUENCY,
    textinfo = "PERCENT",
    type = 'pie', hole = 0.5) |>
  plotly::layout(
    legend = list(orientation = "h", font = t))

bslib::card(
    full_screen = TRUE,
    height = "400px",
    bslib::card_header(
      class = "bg-dark", 
      paste0("Variant filtering statistics - ",
             pcg_report$settings$sample_id)),
    variant_filter_pie
  )


```

#### Settings

The variant filtering has been performed based on the following criteria:

:::: {.columns}

::: {.column width="47.5%"}

* Variant filtering aginst gnomAD (filter *GERMLINE_GNOMAD*): __TRUE__ (not configurable)
* maximum allowed population-specific gnomAD MAF's for somatic events (configurable):
    * AFR: __`r to_settings$maf_gnomad_afr`__
    * AMR: __`r to_settings$maf_gnomad_amr`__
    * ASJ: __`r to_settings$maf_gnomad_asj`__
    * EAS: __`r to_settings$maf_gnomad_eas`__
    * FIN: __`r to_settings$maf_gnomad_fin`__
    * NFE: __`r to_settings$maf_gnomad_nfe`__
    * OTH: __`r to_settings$maf_gnomad_oth`__
    * SAS: __`r to_settings$maf_gnomad_sas`__
* maximum allowed global gnomAD MAF for somatic events: __`r to_settings$maf_gnomad_global`__

:::

::: {.column width="5%"}
<!-- empty column to create gap -->
:::

::: {.column width="47.5%"}
* Variant filtering against ClinVar (filter *GERMLINE_CLINVAR*): __`r as.logical(to_settings$exclude_clinvar_germline)`__
* Variant filtering against dbSNP (filter *GERMLINE_DBSNP*): __`r as.logical(to_settings$exclude_dbsnp_nonsomatic)`__
* Variant filtering based on tumor VAF value (likely heterozygous germline event, VAF in 0.4 - 0.6, filter *GERMLINE_HET*): __`r as.logical(to_settings$exclude_likely_het_germline)`__
* Variant filtering based on tumor VAF value (likely homozygous germline event, VAF = 1, filter *GERMLINE_HOM*): __`r as.logical(to_settings$exclude_likely_hom_germline)`__
* Variant filtering against panel-of-normals (filter *GERMLINE_PON*: __`r as.logical(to_settings$exclude_pon)`__
* Variant filtering based on exonic regions: __`r as.logical(to_settings$exclude_nonexonic)`__

:::

::::

:::

<br><br>
