---
title: "Primary Tumor Sites"
output: rmarkdown::html_document
---

```{r load_pkgs, echo=FALSE, message=FALSE, warning=FALSE}
require(DT)
require(readr)
```

PCGR integrates a variety of different datasets, that typically use different notations or ontologies when
it comes to diseases or phenotypes. We have cross-referenced a variety of ontologies (DiseaseOntology, EFO, OncoTree, UMLS) in use in the oncology domain, and organized all the various tumor subtypes into 30 different main tumor sites (see also https://github.com/sigven/oncoPhenoMap). 

Below you can browse all the various phenotype terms that are associated with each of the primary sites. This as a guide to choose the primary tumor type when running PCGR for a given sample (option `--tumor_site`).

<br><br>

```{r read_table, echo=FALSE, message=FALSE, warning=FALSE}
tab <- system.file("extdata/oncotree_phenotypes.tsv.gz", package = "pcgrr")
readr::read_tsv(tab, col_types = readr::cols(.default = "character")) |>
  dplyr::arrange(primary_site) |>
  DT::datatable(filter = list(position = "top", clear = FALSE, plain = TRUE),
                class = "cell-border display compact",
                rownames = FALSE, extensions = c("Buttons", "KeyTable"),
                options = list(
                  pageLength = 300, autoWidth = FALSE, keys = TRUE,
                  buttons = c("csv", "excel"), dom = "Blfrtip"))
```
