---
title: "Variant classification"
output: rmarkdown::html_document
bibliography: references.bib
---

## Oncogenicity

```{r oncogenicity_note, echo=F, results='asis'}
cat(pcgrr::oncogenicity_doc_note())
```

The following criteria/codes are currently used for variant oncogenicity classification in PCGR (key resources/tools used for implementation indicated in parentheses):

```{r oncogenicity_criteria, echo=F, results='asis'}

i <- 1
while(i <= NROW(pcgrr::oncogenicity_criteria)){
  crit <- pcgrr::oncogenicity_criteria[i,]
  cat(sprintf("  * _%s_ - %s (%s)\n", crit$code, crit$description, crit$resource))
  i <- i + 1
}

```

## Actionability

```{r actionability_note, echo=F, results='asis'}
cat(pcgrr::actionability_doc_note())
```

### Tier I/II — alteration-type agnostic biomarker matching

Tier I and II assignment applies identically to all somatic alteration types
(SNVs/InDels, copy number aberrations, and RNA fusions). Assignment depends on
whether the user has specified a primary tumor site (tumor-type specific mode)
or the generic `Any` site (tumor-type agnostic mode):

**Tumor-type specific** (a primary tumor site is specified, e.g. _Lung_,
_Breast_):

| Biomarker site | Evidence strength | Tier |
|---|---|---|
| Matching site **or** pan-cancer (`Any`) | Strong (CIViC A/B; [OncoKB therapeutic 1/2/3A/3B](https://www.oncokb.org/therapeutic-levels)) | **I** |
| Non-matching tumor-specific site | Strong (CIViC A/B; [OncoKB therapeutic 1/2/3A/3B](https://www.oncokb.org/therapeutic-levels)) | **II** |
| Matching site | Weak (CIViC C/D/E; OncoKB therapeutic 4) | **II** |
| Non-matching tumor-specific site **or** pan-cancer (`Any`) | Weak (CIViC C/D/E; OncoKB therapeutic 4) | **III** |

**Tumor-type agnostic** (primary site set to `Any`):

| Biomarker site | Evidence strength | Tier |
|---|---|---|
| Pan-cancer (`Any`) | Strong (CIViC A/B; [OncoKB therapeutic 1/2/3A/3B](https://www.oncokb.org/therapeutic-levels)) | **I** |
| Non-pan-cancer (any tumor-specific site) | Strong (CIViC A/B; [OncoKB therapeutic 1/2/3A/3B](https://www.oncokb.org/therapeutic-levels)) | **II** |
| Pan-cancer (`Any`) **or** any tumor-specific site | Weak (CIViC C/D/E; OncoKB therapeutic 4) | **III** |

Pan-cancer strong evidence is treated as equivalent to a matching-site strong
item and always promotes a variant to Tier I, regardless of query mode.
Pan-cancer weak evidence contributes to Tier III in both modes.

### Tier III — alteration-type specific routes

Unlike Tier I/II, Tier III assignment is alteration-type specific. Two distinct
routes lead to Tier III and it is important to distinguish between them:

**Route 1 — Biomarker-linked, weak evidence only** (all alteration types)

Variants that _are_ linked to known biomarkers but only through weak clinical
evidence (CIViC C/D/E; [OncoKB therapeutic level 4](https://www.oncokb.org/therapeutic-levels)) in non-matching tumor types
or pan-cancer contexts. These variants have some clinical signal but insufficient
evidence to reach Tier II. This route applies equally to SNVs/InDels, copy
number aberrations, and RNA fusions.

**Route 2 — No biomarker link** (alteration-type specific criteria)

Variants not linked to any known biomarker in the underlying databases, but
flagged as potentially significant based on gene/alteration properties:

| Alteration type | Tier III criterion |
|---|---|
| SNVs/InDels | Coding variant in an oncogene or tumor suppressor gene and population allele frequency < 0.001 (gnomAD) |
| Copy number aberrations | Amplified oncogene, or deleted (homozygous, hemizygous, or heterozygous) tumor suppressor gene |
| RNA fusions | At least one fusion partner gene (5' or 3') is an oncogene and at least one record
for the fusion event in the Mitelman database|

### Evidence items: tier-defining vs. additional

Each biomarker evidence item associated with a variant is classified as either
_tier-defining_ or _additional_ support:

- **Tier-defining** — the item directly justifies the variant's assigned tier
  (e.g. a matching-site strong item on a Tier I variant, or a weak item on a
  Tier III variant)
- **Additional** — the item provides supporting context but did not drive the
  tier assignment (e.g. a non-matching-site strong item on a Tier I variant,
  or pan-cancer weak evidence on a Tier I or II variant)

## References
