
### Tier 2 - Variants of potential clinical significance {.tabset}

- Tier 2 considers evidence items of _i)_ strong evidence levels (A & B) in other tumor types, and _ii)_ weak evidence levels (C, D & E) in the query tumor type (_`r pcg_report[['metadata']][['tumor_class']]`_). Using the [database for clinical interpretations of variants in cancer (CIViC) ](https://civic.genome.wustl.edu/#/home) and [Cancer Biomarkers database](https://www.cancergenomeinterpreter.org/biomarkers), a total of __`r NROW(pcg_report[['content']][['snv_indel']][['variant_display']][['tier2']])`__ unique, somatic variants were found in the tumor sample:
    * Tier 2 - Predictive/Therapeutic: __`r NROW(pcg_report[['content']][['snv_indel']][['clinical_evidence_item']][['other_tumortype']][['predictive']][['A_B']]) + NROW(pcg_report[['content']][['snv_indel']][['clinical_evidence_item']][['specific_tumortype']][['predictive']][['C_D_E']])`__ evidence items
    * Tier 2 - Prognostic: __`r NROW(pcg_report[['content']][['snv_indel']][['clinical_evidence_item']][['other_tumortype']][['prognostic']][['A_B']]) + NROW(pcg_report[['content']][['snv_indel']][['clinical_evidence_item']][['specific_tumortype']][['prognostic']][['C_D_E']])`__ evidence items
    * Tier 2 - Diagnostic: __`r NROW(pcg_report[['content']][['snv_indel']][['clinical_evidence_item']][['other_tumortype']][['diagnostic']][['A_B']]) + NROW(pcg_report[['content']][['snv_indel']][['clinical_evidence_item']][['specific_tumortype']][['diagnostic']][['C_D_E']])`__ evidence items

<br><br>


```{r prepare_tier2_data, echo=F, results='asis'}

show_tier2_filters <- list()
missing_tier2_items <- list()
eitems <- list()
for(type in c('diagnostic','prognostic','predictive')){
  show_tier2_filters[[type]] <- F
  missing_tier2_items[[type]] <- T
  eitems[[type]] <- data.frame()
  if(NROW(pcg_report[['content']][['snv_indel']][['clinical_evidence_item']][['other_tumortype']][[type]][['A_B']]) > 0 | NROW(pcg_report[['content']][['snv_indel']][['clinical_evidence_item']][['specific_tumortype']][[type]][['C_D_E']]) > 0){
    eitems[[type]] <- rbind(pcg_report[['content']][['snv_indel']][['clinical_evidence_item']][['other_tumortype']][[type]][['A_B']], pcg_report[['content']][['snv_indel']][['clinical_evidence_item']][['specific_tumortype']][[type]][['C_D_E']]) %>%
      dplyr::arrange(desc(RATING))
    show_tier2_filters[[type]] <- T
    missing_tier2_items[[type]] <- F
  }
}

```


#### Predictive biomarkers

<br>

```{r tier2_predictive, echo=F, results = 'asis', eval = show_tier2_filters[['predictive']]}
library(summarywidget)
library(crosstalk)

variants_tier2_predictive_shared <- crosstalk::SharedData$new(eitems[['predictive']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", variants_tier2_predictive_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical significance", variants_tier2_predictive_shared, ~CLINICAL_SIGNIFICANCE),
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", variants_tier2_predictive_shared, ~EVIDENCE_LEVEL),
    crosstalk::filter_slider("RATING", "Rating", variants_tier2_predictive_shared, ~RATING, min = 0, max = 5, step = 1, ticks = T)
  ),
  list(
    crosstalk::filter_select("CONSEQUENCE", "Consequence", variants_tier2_predictive_shared, ~CONSEQUENCE),
    crosstalk::filter_select("BIOMARKER_MAPPING", "Biomarker mapping", variants_tier2_predictive_shared, ~BIOMARKER_MAPPING),
    crosstalk::filter_select("THERAPEUTIC_CONTEXT", "Therapeutic context", variants_tier2_predictive_shared, ~THERAPEUTIC_CONTEXT)
  )
)

cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(variants_tier2_predictive_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip')) %>%
 DT::formatStyle('EVIDENCE_LEVEL', backgroundColor = DT::styleEqual(c('A: Validated','A: FDA/NCCN/ELN guidelines','B: Clinical evidence','B1: Clinical evidence: late trials','B2: Clinical evidence: early trials','C: Case study','D: Preclinical evidence','E: Indirect evidence'), c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))
)

#htmltools::br()

```

```{r tier2_predictive_missing, echo=F, results = 'asis', eval = missing_tier2_items[['predictive']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>

#### Prognostic biomarkers

<br>

```{r tier2_prognostic, echo=F, results='asis', eval = show_tier2_filters[['prognostic']]}

library(summarywidget)
library(crosstalk)

variants_tier2_prognostic_shared <- crosstalk::SharedData$new(eitems[['prognostic']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CONSEQUENCE", "Consequence", variants_tier2_prognostic_shared, ~CONSEQUENCE),
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", variants_tier2_prognostic_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical significance", variants_tier2_prognostic_shared, ~CLINICAL_SIGNIFICANCE)
  ),
  list(
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", variants_tier2_prognostic_shared, ~EVIDENCE_LEVEL),
    crosstalk::filter_select("BIOMARKER_MAPPING", "Biomarker mapping", variants_tier2_prognostic_shared, ~BIOMARKER_MAPPING),
    crosstalk::filter_slider("RATING", "Rating", variants_tier2_prognostic_shared, ~RATING, min = 0, max = 5, step = 1, ticks = T)
  )
)
cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(variants_tier2_prognostic_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip')) %>%
  DT::formatStyle('EVIDENCE_LEVEL', backgroundColor = DT::styleEqual(c('A: Validated','A: FDA/NCCN/ELN guidelines','B: Clinical evidence','B1: Clinical evidence: late trials','B2: Clinical evidence: early trials','C: Case study','D: Preclinical evidence','E: Indirect evidence'), c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))
)
#htmltools::br()

```

```{r tier2_prognostic_missing, echo=F, results = 'asis', eval = missing_tier2_items[['prognostic']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>

#### Diagnostic biomarkers

<br>

```{r tier2_diagnostic, echo=F, results='asis', eval = show_tier2_filters[['diagnostic']]}

library(summarywidget)
library(crosstalk)

variants_tier2_diagnostic_shared <- crosstalk::SharedData$new(eitems[['diagnostic']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CONSEQUENCE", "Consequence", variants_tier2_diagnostic_shared, ~CONSEQUENCE),
    crosstalk::filter_select("CANCER_TYPE", "Cancer type", variants_tier2_diagnostic_shared, ~CANCER_TYPE),
    crosstalk::filter_select("CLINICAL_SIGNIFICANCE", "Clinical association", variants_tier2_diagnostic_shared, ~CLINICAL_SIGNIFICANCE)

  ),
  list(
    crosstalk::filter_select("EVIDENCE_LEVEL", "Evidence level", variants_tier2_diagnostic_shared, ~EVIDENCE_LEVEL),
    crosstalk::filter_select("BIOMARKER_MAPPING", "Biomarker mapping", variants_tier2_diagnostic_shared, ~BIOMARKER_MAPPING),
    crosstalk::filter_slider("RATING", "Rating", variants_tier2_diagnostic_shared, ~RATING, min = 0, max = 5, step = 1, ticks = T),
  )
)
cat('The table below lists all variant-evidence item associations:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(variants_tier2_diagnostic_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip')) %>%
  DT::formatStyle('EVIDENCE_LEVEL', backgroundColor = DT::styleEqual(c('A: Validated','A: FDA/NCCN/ELN guidelines','B: Clinical evidence','B1: Clinical evidence: late trials','B2: Clinical evidence: early trials','C: Case study','D: Preclinical evidence','E: Indirect evidence'), c("#009E73","#009E73","#56B4E9", "#56B4E9","#56B4E9","#0072B2","#E69F00", "#F0E442")))
)
#htmltools::br()

```

```{r tier2_diagnostic_missing, echo=F, results = 'asis', eval = missing_tier2_items[['diagnostic']]}
cat('<i>No variant-evidence item associations found.</i>',sep='\n')
cat('\n')
```

<br><br>


