
### Tier 3 - Variants of unknown clinical significance {.tabset}

- A total of __`r NROW(pcg_report[['snv_indel']][['variant_display']][['tier3']][['proto_oncogene']]) + NROW(pcg_report[['snv_indel']][['variant_display']][['tier3']][['tumor_suppressor']])`__ unique, somatic variant(s) in the tumor sample are of unknown clinical significance, as found within known proto-oncogenes or tumor suppressor genes.
<br><br>


```{r prepare_tier3_data, echo=F, results='asis'}

show_tier3_filters <- list()
missing_tier3_items <- list()
for(type in c('proto_oncogene','tumor_suppressor')){
  show_tier3_filters[[type]] <- F
  missing_tier3_items[[type]] <- T
  if(NROW(pcg_report[['snv_indel']][['variant_display']][['tier3']][[type]]) > 0){
    show_tier3_filters[[type]] <- T
    missing_tier3_items[[type]] <- F
  }
}

```


#### Tumor suppressor gene mutations

<br>

```{r tsgene_mutations, echo=F, results = 'asis', eval = show_tier3_filters[['tumor_suppressor']]}
library(summarywidget)
library(crosstalk)

variants_tsgene_shared <- crosstalk::SharedData$new(pcg_report[['snv_indel']][['variant_display']][['tier3']][['tumor_suppressor']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CONSEQUENCE", "Consequence", variants_tsgene_shared, ~CONSEQUENCE),
    crosstalk::filter_select("MUTATION_HOTSPOT", "Mutation hotspot", variants_tsgene_shared, ~MUTATION_HOTSPOT)
  ),
  list(
    crosstalk::filter_select("INTOGEN_DRIVER_MUT", "Predicted driver mutation", variants_tsgene_shared, ~INTOGEN_DRIVER_MUT)
    
  )
)

cat('The table below lists all variants:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(variants_tsgene_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip'))
)

#htmltools::br()

```

```{r tsgene_missing, echo=F, results = 'asis', eval = missing_tier3_items[['tumor_suppressor']]}
cat('<i>No variants found.</i>',sep='\n')
cat('\n')
```

<br><br>

#### Proto-oncogene mutations

<br>

```{r oncogene_mutations, echo=F, results = 'asis', eval = show_tier3_filters[['proto_oncogene']]}
library(summarywidget)
library(crosstalk)

variants_oncogene_shared <- crosstalk::SharedData$new(pcg_report[['snv_indel']][['variant_display']][['tier3']][['proto_oncogene']])
crosstalk::bscols(
  list(
    crosstalk::filter_select("CONSEQUENCE", "Consequence", variants_oncogene_shared, ~CONSEQUENCE),
    crosstalk::filter_select("MUTATION_HOTSPOT", "Mutation hotspot", variants_oncogene_shared, ~MUTATION_HOTSPOT)
  ),
  list(
    crosstalk::filter_select("INTOGEN_DRIVER_MUT", "Predicted driver mutation", variants_oncogene_shared, ~INTOGEN_DRIVER_MUT)
    
  )
)

cat('The table below lists all variants:',sep='\n')
htmltools::br()
htmltools::br()

crosstalk::bscols(
  DT::datatable(variants_oncogene_shared, escape=F,extensions=c("Buttons","Responsive"), width = "100%",options=list(buttons = c('csv','excel'),dom = 'Bfrtip'))
)

#htmltools::br()

```

```{r oncogene_missing, echo=F, results = 'asis', eval = missing_tier3_items[['proto_oncogene']]}
cat('<i>No variants found.</i>',sep='\n')
cat('\n')
```

<br><br>
